---
title: "Bayesian UPL"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian-UPL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  error=TRUE,
  warning=FALSE,
  message=FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(forcats)
library(EPA.MACT.floor.UPL)
library(matrixStats)
library(EnvStats)
library(np)
library(sfsmisc)
library(parallel)
library(rjags)
library(runjags)
library(sn)
```


## Introduction

This vignette describes the use of the `Bayesian_UPL()` function. The goal of 
`Bayesian_UPL()` is to provide a flexible and consistent method of calculating 
Upper Predictive Limits (UPL) for any type of emissions data to determine 
National Emissions Standards for Hazardous Air Pollutants (NESHAP) for existing 
and new sources. 

Current UPL methods are only applicable to emissions data that are distributed 
Normally, Skewed, or Lognormally. The UPL for Normal data can be calculated 
analytically with `Normal_UPL()` using the mean, standard deviation, and t-score. 
However, there are no equivalent analytical solutions for Skewed or Lognormal 
UPL calculations. Instead, `Lognormal_UPL()` uses a Gram-Charlier Type A Series 
expansion to approximate the density distribution and determine the arithmetic 
mean of future samples, and `Skewed_UPL()` uses the skewness and kurtosis 
moments to iteratively adjust the t-score until the desired significance is 
achieved. All three UPL methods assume the distribution explicitly and without
uncertainty, and assume the emissions are observed perfectly and independently.

By operating in a Bayesian framework, we can use the same method of deriving the 
UPL while agnostic to the type of distribution, but also dramatically increase the 
choices for the distribution. Bayesian inference also benefits from fewer 
assumptions and better uncertainty quantification. This allows for more accurate 
representation of emissions data and better transparency and metrics of fitness 
to support the UPL calculation. 

The `Bayesian_UPL()` uses a Monte Carlo Markov Chain (MCMC) sampler (a Gibbs 
sampler implemented via JAGS) to fit probability distributions to emissions data. 
For example, when using a Normal distribution, this entails estimating the mean 
and standard deviation which define the distribution. Unlike the `Normal_UPL()` 
analytical solution, we don't merely get a point-estimate of the mean and 
standard deviation, but instead derive a posterior distribution that describes 
the full probability for the mean and standard deviation. This allows us to 
quantify the uncertainty in the predicted probability distribution for any 
future emission observation. 

Numerous probability distributions are already implemented in R `stats` and JAGS, 
and any distribution that isn't, or any combination of distributions, can be 
coded manually in JAGS if needed. For any probability distribution, the UPL is 
calculated by taking the 99th percentile of the average of new draws (usually 3 
new runs to be comparative to a typical compliance test).

## Example: HCl Emissions from Integrated Iron and Steel

### Step 1: Load and plot the emissions observations

Let's look at an example data set of HCl emissions from Integrated Iron and Steel.
This data set as provided is already the set of top performers in the industry, 
so there is no need to run `MACT_EG()` to select the data. The first step is to 
load and plot the emissions data. 

```{r load, echo=TRUE}
dat_emiss=read.csv('IIS_HCl.csv')

emission_mean=mean(dat_emiss$emissions)
dat_means=summarize(dat_emiss,means=mean(emissions),counts=n(),.by='sources')

dat_means$sources=as.factor(dat_means$sources)
dat_means$sources=fct_reorder(dat_means$sources,
                              dat_means$means,.desc = FALSE)
dat_emiss$sources=factor(dat_emiss$sources,levels=levels(dat_means$sources))
dat_means=arrange(dat_means,means)
```

Take a quick look at the average emission for each source in the data set:

```{r table1}
knitr::kable(mutate(dat_means,across(c(2),~signif(.,digits=3))),
             caption="Top sources for existing source calculation",
             col.names=c('Source','Average emission','No. of Tests'))
```

Next we want to look at density plots of the emission observations. Our goal is 
to fit a probability density that best describes the data, from which we can 
determine the UPL. In order to know the probability density is a good match for 
our emissions data, we want to look at the empirical density of the observations. 
This is easy to do with `ggplot2::geom_density()`, which uses `stats::density()` 
and can take additional arguments related to how the density is determined such 
as the bandwidth, `bw`, kernel `kernel`, and known upper and lower limits in 
`bounds`. For this example we will plot the density by `source` using the default 
bandwidth and kernel (gaussian) with a lower limit of `c(0,Inf)` since negative 
emissions are impossible. The `UPLforOAR` package includes several `ggplot2` 
themes designed for density figures. The figure below uses `multi_source_theme()`.

```{r figure1, fig.dim=c(5.5,4), fig.align = 'center', fig.cap="Observation density of IIS HCl emissions by top performing source."}

ggplot(aes(emissions,fill=sources),data=dat_emiss)+
  geom_density(alpha=0.5,trim=FALSE,bounds=c(0,Inf))+
  multi_source_theme()+theme(legend.text = element_text(size=8))+
  guides(fill=guide_legend(nrow=2,byrow=TRUE),
         color=guide_legend(nrow=2,byrow=TRUE))+
  scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
  scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
  geom_rug(sides='b',aes(x=emissions,color=sources),
           alpha=0.5,outside=TRUE)+coord_cartesian(clip='off')+
  ggtitle("Source observed populations")+
  ylab("Density")+xlab("IIS HCl emissions")+
  labs(fill='Top sources',color='Top sources')

```

In most situations, we will want more control over the density than `ggplot2::geom_density()` provides, and we will want to use the observation densities 
outside of the plot later on. For example, all emissions data will be bounded by 
0 and will often have an upper bound as well. Due to the strictly positive 
nature of emissions data, a gaussian kernel is not often a good choice. 

Instead we will use the `obs_density()` function in `UPLforOAR`, which builds on 
the `np::npuniden.boundary()` function, a more accurate density function for 
bounded data. By default, `obs_density()` will assume a lower bound of 0, no 
upper bound, and the `kernel` type is gamma. Other kernel types accepted by `np::npuniden.boundary()` can be used, for example if data are percent removals instead of emissions then "beta1" or "beta2" with an upper bound of 1 are more appropriate. The default bandwidth `bw` is $\sigma*n^{-2/5}$ (Renault & Scaillet 
2004), though manual values or a least-squares or likelihood search can be used 
instead by setting `bw = "cv.ls"` or `bw = "cv.ml"`.

The results of `obs_density()` include `ObsonPoint`, with the observation densities 
paired with each emission value in the provided data set, and `obs_den_df`, a 
sequence of densities across a range of possible emission values (`xvals`) based on the 
observed emissions. The range of emissions `xvals` can be supplied manually, and 
by default is `seq(0,3*max(data$emissions,length.out=1024))`. 

```{r obs_density, echo=TRUE}
obs_dens_results=obs_density(data=dat_emiss)
Obs_onPoint=obs_dens_results$Obs_onPoint
obs_den_df=obs_dens_results$obs_den_df
```

The figure below shows the overall observation density of the top performing source 
population, of which it is our goal to determine an appropriately representative 
likelihood distribution from which to calculate the UPL. This figures uses 
`pop_distr_theme()` and plots the `obs_density()` results above using `ggplot2::geom_line` and `ggplot2::geom_line`. The emission observations themselves 
are indicated as points and as a rug underneath the plot.

```{r figure2, fig.dim=c(5,3), fig.align = 'center', fig.cap="Observation density of IIS HCl emissions top performers overall population."}

ggplot(data=Obs_onPoint)+
  geom_line(data=obs_den_df,aes(y=ydens,x=(x_hat)),size=0.75,color='black')+
  geom_area(data=obs_den_df,aes(y=ydens,x=(x_hat)),alpha=0.25,fill='black')+
  geom_point(aes(y=ydens,x=(emissions)),
             size=3,alpha=0.5,shape=19,color='black')+
  ylab("Density")+xlab("IIS HCl emissions")+
  ggtitle("Overall observed population")+
  pop_distr_theme()+
  scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
  scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
  coord_cartesian(clip='off')+
  geom_rug(sides='b',aes(x=(emissions)),data=dat_emiss,
           alpha=0.5,outside=TRUE,color='black')

```
### Step 2: Explore possible likelihood distributions
You can use an informed decision about which likelihood distributions to investigate. 
For example, a Beta distribution is only appropriate if the observations are strictly 
between 0 and 1. In this case, you might see that a Normal distribution is a poor 
choice given asymmetric shape of the observations clustered above 0. However, 
for the purposes of the vignette, we will look at all distributions currently supported.

We will fit likelihood models for the following distributions: Beta, Gamma, 
Lognormal, Normal, and Skewed. Using the default settings of `Bayesian_UPL()` the 
only other information we need to provide is the data set with `emissions`. The 
`Bayesian_UPL()` function will fit the likelihood distributions using uninformative priors with a wide range of initial values, test for convergence, test for goodness of fit, organize the MCMC results for later plotting, and return the UPL assuming a 
significance of 0.99 for 3 future emission runs. Many of these options can be 
adjusted manually or evaluated step wise, but `Bayesian_UPL()` is designed to run 
with minimal user input needed.

```{r bayesian_UPL, echo=TRUE, results=FALSE}
distributions=c('Beta','Gamma','Lognormal','Normal','Skewed')
results=Bayesian_UPL(data=dat_emiss, distr_list=distributions)
```

First, let's look at plots of the resulting distributions. 

```{r figure3, fig.dim=c(5,3), fig.align = 'center', fig.cap="Fitted likelihood distributions for IIS HCl emissions."}
palette_line=c("#377EB8", "#4DAF4A","#FF7F00", "#984EA3", "#E41A1C")
ggplot()+
  geom_line(data=results$pred_pdf_dat,
            aes(y=ydens,x=(x_hat)),size=0.75,color='black')+
  geom_area(data=results$pred_pdf_dat,
            aes(y=ydens,x=(x_hat)),fill='grey')+
  geom_point(data=results$obs_pdf_dat,
            aes(y=ydens,x=(emissions)),
            size=3,alpha=0.5,shape=19,color='black')+
  geom_line(data=results$pred_pdf_dat,
            aes(x=(x_hat),y=pdf_hat,color=distr),size=0.75)+
  geom_area(data=results$pred_pdf_dat,
            aes(x=(x_hat),y=pdf_hat,fill=distr),alpha=0.5)+
  geom_errorbar(data=results$obs_pdf_dat,
                aes(ymin=low,ymax=up,x=(emissions),color=distr),
                alpha=0.5)+
  geom_point(data=results$obs_pdf_dat,
             aes(x=(emissions),y=med,fill=distr),
             size=3,alpha=0.5,shape=21,color='black')+
  ylab("Density")+xlab("emissions")+
  pop_distr_theme()+labs(color='Distribution',fill="Distribution")+
  scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
  scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
  geom_rug(data=results$obs_pdf_dat,
           sides='b',aes(x=(emissions)),
           alpha=0.5,outside=TRUE,color='black')+
  coord_cartesian(clip='off')+scale_color_manual(values=palette_line)+
  scale_fill_manual(values=palette_line)+
  facet_wrap(~distr,scales='free_y',nrow=3)+theme(legend.position = 'none')+
  theme(strip.text = element_text(color='black',size=12))
```

## Fit



## Priors

Bayesian inference uses prior information to derive posterior distributions of 
model parameters. The default behavior in `Bayesian_UPL()` is to use uninformative 
priors which is calculated based on the range and variance of the emissions data. 
The prior information can be manually supplied to either increase the parameter 
space to search wider, or to provide more specific information, which might be 
particularly useful in circumstances where there are very few observations (i.e. 
`n = 3`). 

## Convergence

Convergence of distribution parameters is assessed quantitatively using the 
Gelman-Rubin diagnostic and qualitatively by producing figures in a convergence 
report pdf document. 

## Long- and Heavy-Tailed Distributions 

For example some Lognormal. The 99th percentile can be unreasonably large 
emission number when long or heavy tail. This is often an issue when the 
observations themselves are no where near the tail, and thus the tail is largely 
unsupported by any data. In such circumstances, the UPL can be a value larger 
than any in the observation data set, or larger than is physically possible. In 
such cases it is necessary to truncate the probability distribution to physically 
reasonable bounds. This is done by default as 3 times the maximum emissions value 
in the training data, but can be adjusted by setting `maxY`. 








