---
title: "Convergence and Priors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Convergence-and-Priors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  error = TRUE,
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
library(htmltools)
knitr::knit_hooks$set(imgcenter = function(before, options, envir){
  if (before) {
    HTML("<p align='center'>")
  } else {
    HTML("</p>")
  }
})
```

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(forcats)
library(UPLforOAR)
library(matrixStats)
library(np)
library(sfsmisc)
library(parallel)
library(rjags)
library(runjags)
library(sn)
library(patchwork)
```

## Introduction

This vignette provides further details on convergence and prior distributions 
for `Bayesian_UPL()`. We recommend already having some familiarity with using 
the default settings of `Bayesian_UPL()` and working through the examples in 
`vignette("Bayesian-UPL")` before reading further. 

## Convergence

Convergence of distribution parameters is assessed quantitatively using the 
Gelman-Rubin diagnostic and qualitatively by producing figures in a convergence 
report pdf document. 

The diagnostics are available in the `$conv_output` table from `Bayesian_UPL()` 
or from using `converge_likelihood()` directly on the results of `run_likelihood()`. 
There are several variables in the `conv_output` table: the distribution as `distr`, 
its corresponding parameters in `params`, the diagnostic in `gelman_diag`, and 
if it converged in `convYN`. For example, a Normal distribution will have both 
mean and standard deviation parameters in the table, where as a Gamma distribution 
will have shape and rate. For `gelman_diag` less than 1.1, the outcome is `"Yes"` 
it converged in `convYN`. For diagnostic values between 1.1 and 1.2 it is 
`"Weak convergence"`, and `"No"` convergence for values greater than 1.2. 

### Problems: Failing Gelman-Rubin diagnostic

Things have to go really wrong for the Gelman-Rubin diagnostics to fail. You can 
fail these checks if you do not have a long enough number of iterations in the 
burn-in or a long enough number of kept iterations. However `Bayesian_UPL()` is 
set up so that the 10,000 iterations for burn-in and sampling should be sufficient 
as long as the likelihood model is somewhat reasonable. The diagnostic also relies 
on comparing multiple MCMC chains with substantially different initial values, 
which again is handled automatically in `Bayesian_UPL()`. Below is an example 
where we generated 30 random emissions between 0 and 100 and then tried to fit 
a Lognormal likelihood distribution. Unsurprisingly, it fails to converge.

```{r noise, echo = TRUE, results = FALSE}
set.seed(1)
emissions = runif(30, min = 0, max = 100)
results = Bayesian_UPL(distr_list = c("Lognormal"),
                       data = tibble(emissions = emissions))
conv_metrics = results$conv_output
```
``` {r table1, fig.align = "center"}
conv_metrics = mutate(conv_metrics, across(c(3), ~signif(., digits = 4)))
knitr::kable(conv_metrics,
             caption = "Gelman-Rubin convergence tests for likelihood parameters",
             col.names = c("Distribution",'Parameter',"Diagnostic","Converged"))
```

We can take a look at the figures generated in the convergence report as well. 
Note that for both parameters, the left figures show one of the chains is off on 
a tangent and they are clearly not well-mixed or without trend. This is the kind 
of behavior that will result in the Gelman-Rubin diagnostic failing.
```{r figure1, fig.height = 4, out.width = "100%", fig.align = "center", results = FALSE, fig.cap = "Example of bad convergence in MCMC chains."}
part1 = setup_likelihood(distribution = "Lognormal",
                         data = tibble(emissions = emissions))
part2 = run_likelihood(model_input = part1)
part3 = converge_figs(distribution = "Lognormal", jags_model_run = part2)
part3
```

It is possible for a parameter to pass the Gelman-Rubin test and indeed have chains 
well-mixed, but not arrive at a clear posterior distribution. That is why passing 
the Gelman-Rubin diagnostic is necessary but insufficient to indicate convergence. 
This can happen when there is too little data (for example extremely small data 
sets where `n = 3`), and the parameter posterior is simply returning the prior 
distribution. For example, a poorly converged Normal distribution might have an 
even probability of the mean being any emission value, in which case the histogram 
will be flat across the entire parameter-space. 

### Problems: Flat posterior

In the example below, we use only 3 observations used to fit a Normal distribution. 
This is as small a data set as one could conceivably use, and with only 3 runs 
there is not likely to be enough information to pick an appropriate distribution 
or determine its defining parameters with high confidence. Even though the 
Gelman-Rubin diagnostics passed in this case, and indeed the MCMC chains look 
well-mixed (left figures), it has not converged well. The histograms on the right 
show the posterior is too flat. Note that there is a slight peak for the 
`emission_mean` posterior around 0.01, which is indeed the average of the 3 points, 
but there is not enough information to definitely say this is much more likely 
than any other value searched.  
```{r figure2, fig.height = 4, out.width = "100%", fig.align = "center", results = FALSE, fig.cap = "Example of bad convergence where posterior is too flat."}
dat_emiss2 = read.csv("Example_data1.csv")
small_dat = MACT_new(dat_emiss2)
part1 = setup_likelihood(distribution = "Normal", data = small_dat)
part2 = run_likelihood(model_input = part1)
part3 = converge_figs(distribution = "Normal", jags_model_run = part2)
part3
```

### Problems: Bad prior limits

The visual assessment will also catch any issues with the prior distribution 
settings. Ideally, we are using uninformative priors for UPL calculations that 
have equal probability across a parameter-space that is much wider than the 
posterior result. This is determined automatically from the data, but in some 
cases that range might not wide enough. This will be apparent if the parameter's 
posterior histogram is highly skewed towards the end of the allowed range. In 
this case, the priors will need to be set manually to a wider uninformative range. 
As an example, we've fit a Normal distribution to the HCl emissions from 
Integrated Iron and Steel but manually set the prior range so the mean is equally 
likely between `~dunif(0.01, 0.05)`. Given that the average emission is much lower 
than this range at 0.001, this is likely to cause issues for convergence even 
though it does pass the Gelman-Rubin checks. Even though the MCMC chains 
are well mixed, they are clearly stacking up against the lower bound and trying 
to converge on a lower mean than we have allowed. This skew in the chains (left) 
and histogram (right) indicate improper prior distribution bounds. 
```{r figure3, fig.height = 4, out.width = "100%", fig.align = "center", results = FALSE, fig.cap = "Example of bad convergence, where the prior limits are too restrictive."}
dat_emiss = read.csv('Example_data2.csv')
part1 = setup_likelihood(distribution = "Normal", data = dat_emiss, 
                         manual_prior = TRUE, 
                         prior_list = c(0.0001, 0.001, 0.01, 0.05))
part2 = run_likelihood(model_input = part1)
part3 = converge_figs(distribution = "Normal", jags_model_run = part2)
part3[[1]]
```

## Priors

Bayesian inference uses prior information to derive posterior distributions of 
model parameters. The default behavior in `Bayesian_UPL()` is to use uninformative 
priors which are calculated based on the range and variance of the emissions data. 
The prior information can be manually supplied to either increase the parameter 
space to search wider, or to provide more specific information, which might be 
particularly useful in circumstances where there are very few observations (i.e. 
`n = 3`). 

It is inadvisable to use manual priors without good justification. A situation 
where one might use the manual prior option is after running `Bayesian_UPL()` 
with the default uninformative priors and the parameters showed issues with 
convergence. For example, if the posterior histogram is completely flat across 
the default parameter space, and information from the data are limited by a low 
sample size, then a narrower parameter range might need to be provided manually 
to promote convergence. Alternatively, if the posterior histogram appears to be 
stacked up along one side of the parameter space, then a wider prior range might 
need to be provided manually if applicable. Some parameters can only exist above 
0, or -1, or between -1 and 1, and in these cases it is not abnormal behavior 
for their posterior distributions to be heavily skewed when close to a boundary, 
and setting further limits manually will not change any outcomes.

### Manual limits on priors

In the convergence section above we tried to fit a Normal distribution to an emission data 
set with `n = 3` and fully uninformed priors, and did not achieve good convergence. 
We could then decide to follow up with another Normal distribution where we 
provide slightly narrower but still largely uninformative priors manually. The 
first pair in `prior_list` are the bounds on standard deviation and encompass several orders of 
magnitude around the actual standard deviation of the 3 observations. The second 
pair of bounds in `prior_list` encompass the actual mean of 0.01, but aren't quite as wide as 
the automated uninformative priors uses. This slight restriction on the prior 
search is enough to promote better convergence, as can be seen in the two 
resulting fitted probability density distributions below. Note that both have 
very large error bars around the probability density at the emission observations. 
This is expected given how little information we actually get from 3 observations. 
The ability to quantify this uncertainty is a nice outcome of using a Bayesian approach.
```{r manual, echo = TRUE, fig.height = 4, out.width = "100%", fig.align = "center", results = FALSE}
result_informed = Bayesian_UPL(distr_list = c("Normal"), data = small_dat,
                               manual_prior = TRUE,
                               prior_list = c(0.0001, 0.01, 0.005, 0.03))
result_uninformed = Bayesian_UPL(distr_list = c("Normal"), data = small_dat)
```
```{r figure4, echo = FALSE, fig.height = 5, fig.width = 4, fig.align = "center", fig.cap = "Example of two Normal distributions fit to the same 3 points of data, where one was more informative using slightly smaller bounds on prior distributions, and the other was completely uninformative and did not converge well."}
result_uninformed$obs_pdf_dat$distr = rep("Uninformed", 
                                          nrow(result_uninformed$obs_pdf_dat))
result_uninformed$pred_pdf_dat$distr = rep("Uninformed", 
                                          nrow(result_uninformed$pred_pdf_dat))
result_uninformed$fit_table$distr = rep("Uninformed", 
                                          nrow(result_informed$fit_table))
result_informed$obs_pdf_dat$distr = rep("Informed", 
                                          nrow(result_informed$obs_pdf_dat))
result_informed$pred_pdf_dat$distr = rep("Informed", 
                                          nrow(result_informed$pred_pdf_dat))
result_informed$fit_table$distr = rep("Informed", 
                                          nrow(result_informed$fit_table))
combined_pdf = rbind(result_uninformed$pred_pdf_dat, result_informed$pred_pdf_dat)
combined_obs = rbind(result_uninformed$obs_pdf_dat, result_informed$obs_pdf_dat)
combined_fit = rbind(result_uninformed$fit_table, result_informed$fit_table)

palette_line = c("#377EB8", "#E41A1C")
ggplot()+
  geom_line(data = combined_pdf,
            aes(y = ydens, x = x_hat), size = 0.75, color = 'black')+
  geom_area(data = combined_pdf,
            aes(y = ydens, x = x_hat), fill = 'grey')+
  geom_point(data = combined_obs,
            aes(y = ydens, x = emissions),
            size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_line(data = combined_pdf,
            aes(x = x_hat, y = pdf_hat, color = distr), size = 0.75)+
  geom_area(data = combined_pdf,
            aes(x = x_hat, y = pdf_hat, fill = distr), alpha = 0.5)+
  geom_errorbar(data = combined_obs,
                aes(ymin = low, ymax = up, x = emissions, color = distr),
                alpha = 0.5)+
  geom_point(data = combined_obs,
             aes(x = emissions, y = med, fill = distr),
             size = 3, alpha = 0.5, shape = 21, color = 'black')+
  geom_vline(aes(xintercept = UPL, color = distr), data = combined_fit,
             linewidth = 1)+
  ylab("Density")+ xlab("Emissions")+
  pop_distr_theme()+ labs(color = 'Distribution', fill = "Distribution")+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  geom_rug(data = combined_obs,
           sides = 'b', aes( x =emissions),
           alpha = 0.5, outside = TRUE, color = 'black')+
  coord_cartesian(clip = 'off')+ scale_color_manual(values = palette_line)+
  scale_fill_manual(values = palette_line)+
  facet_wrap(~distr, scales = 'fixed', nrow = 3)+ theme(legend.position = 'none')+
  theme(strip.text = element_text(color = 'black', size = 12))
```

We can look at how the convergence figures are different in the case where we 
provided slightly more prior information. Note that the histograms are much more 
converged than the flat horizontal coverage they were when using the automated 
uninformed priors. Even though we provided the prior distributions manually, they 
are still considered to be uninformative. We have restricted the parameter space 
with lower and upper bounds, but it is still a wide accommodation, and within 
that all parameter outcomes are equally likely before we consider the data.
```{r figure5, fig.height = 4, out.width = "100%", fig.align = "center", results = FALSE, echo = FALSE, fig.cap = "Improved convergence of parameters by providing slightly more restrictive bounds as priors."}
part1 = setup_likelihood(distribution = "Normal", data = small_dat,
                         manual_prior = TRUE,
                         prior_list = c(0.0001, 0.01, 0.005, 0.03))
part2 = run_likelihood(model_input = part1)
part3 = converge_figs(distribution = "Normal", jags_model_run = part2)
part3
```

### Specific priors

In some cases there might be specific prior information about the distributions 
that can be used to set informative priors. Examples could include an operational 
lower or upper limit to emissions that physically cannot be surpassed, or the 
results from a previous decade of emissions that technology has improved on since. 
Specific prior information could be in the form of an informative distribution, 
such as `~dnorm(5, 0.1)`, rather than lower and upper bounds. In this case, the 
JAGS model scripts created by `write_likelihood()` will need to be edited to 
reflect the specifically defined distributions. Caution should be exercised if 
using stricter priors, and needs to be very well justified. 
