---
title: "Density"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Density}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  error=TRUE,
  warning=FALSE,
  message=FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
library(htmltools)
knitr::knit_hooks$set(imgcenter = function(before, options, envir){
  if (before) {
    HTML("<p align='center'>")
  } else {
    HTML("</p>")
  }
})
```

```{r setup}
library(dplyr)
library(ggplot2)
library(forcats)
library(UPLforOAR)
library(matrixStats)
library(EnvStats)
library(np)
library(sfsmisc)
library(parallel)
library(rjags)
library(runjags)
library(sn)
library(patchwork)
```

## Introduction

This vignette describes the use empirical density functions that are used to 
evaluate if the probability density function of emissions data are a good 
representation to use for the Upper Predictive Limit (UPL) calculation for 
standards. Fitting a probability density function is a necessary step for 
calculating the UPL. We may only have a limited sample of emissions observations, 
but the probability density function describes how likely an emission at *any* 
value is to occur. The y-axis is density, and relates to likelihood or frequency 
of observation. The x-axis is the emission value. The density can be interpreted 
as a smoothed-curve version of a histogram. They should have similar shapes if 
appropriate bandwidths are used. The higher the density, the more likely an 
emission value is. The scale of the density on the y-axis is unimportant, since 
it is always set such that the area under the curve of the probability density 
function sums to 1. If it does not integrate to 1, it is not a proper probability 
density function. Once the probability density at all emission values is defined, 
we can calculate the UPL as the upper 99<sup>th</sup> percentile of an average 
of a number of future emissions. 

The best way to both visualize the likelihood of emission observations and 
evaluate the probability density function used for UPL calculations is to compare 
to empirical densities. 

## Example 

We will look at some synthetic data generated with a mixture of Normal and 
Lognormal distributions for 5 sources.
```{r load, echo=TRUE}
set.seed(1)
emiss_ln1 = rlnorm(18, meanlog = log(0.01), sdlog = 2)
source1 = rep("A", 18)
emiss_ln2 = rlnorm(6, meanlog = log(0.02), sdlog = 1.2)
source2 = rep("B", 6)
emiss_ln3 = rlnorm(3, meanlog = log(0.025), sdlog = 0.5)
source3 = rep("C", 3)
emiss_n4 = rnorm(18, mean = 0.02, sd = 0.01)
source4 = rep("D", 18)
emiss_n5 = rnorm(12, mean = 0.03, sd = 0.01)
source5 = rep("E", 12)
emissions = c(emiss_ln1, emiss_ln2, emiss_ln3, emiss_n4, emiss_n5)
sources = c(source1, source2, source3, source4, source5)
dat_emiss = tibble(emissions = emissions, sources = sources)
```

First we want to look at density plots of the emission observations by source.
This is easy to do with `ggplot2::geom_density()`, which uses `stats::density()` 
and can take additional arguments related to how the density is determined such 
as the bandwidth `bw`, kernel `kernel`, and known upper and lower limits in 
`bounds`. By default, `geom_density()` uses no bounds, a gaussian kernel, and 
a rule-of-thumb method (not recommended) of arriving at a bandwidth 
(Silverman B. W. 1986. Density Estimation p.48 eq 3.31). For this example we 
will plot the density using `ggplot2::geom_density()` by `source`. The 
`UPLforOAR` package includes several `ggplot2` themes designed for density 
figures. The figure below uses `multi_source_theme()`.
```{r figure1, echo=TRUE, fig.dim=c(5.5,4), fig.align = "center", fig.cap="Observation density of simulated emissions by source."}
ggplot(aes(emissions, fill = sources), data = dat_emiss)+
  geom_density(alpha = 0.5, trim = FALSE)+
  multi_source_theme()+
  scale_x_continuous(limits =c(-0.05, 0.25))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  ggtitle("Source Densities")+
  ylab("Density")+xlab("Emissions")+
  labs(fill = 'Top sources', color = 'Top sources')
```

Note that the density curves for some sources extend below zero, even though all of the emissions are positive. This is not ideal, since we know that emissions cannot be negative. 
Let's plot this again but set a lower bound with `bounds = c(0, Inf)`. Note the 
change in the shape of the density curve now that there isn't any extrapolation 
past zero. The behavior that `ggplot` uses to accommodate bounds for 
`geom_density()` involves reflecting the density back across the boundary. This 
is alright, but there are better methods that use density estimators specific 
for bounded circumstances.
```{r figure2, echo=TRUE, fig.dim=c(5.5,4), fig.align = "center", fig.cap="Observation density of simulated emissions by source, bounded strictly by 0."}
ggplot(aes(emissions, fill = sources), data = dat_emiss)+
  geom_density(alpha = 0.5, trim = FALSE, bounds = c(0, Inf))+
  multi_source_theme()+
  scale_x_continuous(limits = c(-0.05, 0.25))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  ggtitle("Source Densities")+
  ylab("Density")+xlab("Emissions")+
  labs(fill = 'Top sources', color = 'Top sources')
```

Next, we will use a different way of plotting densities that is more appropriate. The 
default settings in `obs_density()` are designed to be compatible with typical 
emissions data, such as appropriate kernel types, bandwidths, bounds, and 
estimators, for non-zero data, as well as a better default bandwidth estimator. 
Furthermore, we can store the results directly in our environment rather than 
them being inside the `ggplot` object.

```{r obs_density, echo=TRUE}
obs_dens_results = obs_density(data = dat_emiss)
Obs_onPoint = obs_dens_results$Obs_onPoint
obs_den_df = obs_dens_results$obs_den_df
```

The figure below shows the overall observation density of the combined source 
population, of which it is our goal to determine an appropriately representative 
likelihood distribution from which to calculate the UPL. This figures uses 
`pop_distr_theme()` and plots the `obs_density()` results above using 
`ggplot2::geom_line()` and `ggplot2::geom_line()`. The emission observations 
themselves are indicated as points and as a rug underneath the plot.
```{r figure3, echo=TRUE, fig.dim=c(5,3), fig.align = "center", fig.cap="Observation density of IIS HCl emissions top performers overall population."}
ggplot(data = Obs_onPoint)+
  geom_line(data = obs_den_df, aes(y = ydens, x = x_hat), 
            size = 0.75, color = 'black')+
  geom_area(data = obs_den_df, aes(y = ydens, x = x_hat), 
            alpha = 0.25, fill = 'black')+
  geom_point(aes(y = ydens,x = emissions),
             size = 3, alpha = 0.5, shape = 19, color = 'black')+
  ylab("Density")+xlab("Emissions")+
  ggtitle("Overall observed population")+
  pop_distr_theme()+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  coord_cartesian(clip = 'off')+
  geom_rug(sides = 'b', aes(x = emissions), data = dat_emiss,
           alpha = 0.5, outside = TRUE, color = 'black')
```

## Details 

In most situations, we will want more control over the density than `ggplot2::geom_density()` 
provides, and we will want to use the observation densities 
outside of the plot later on. For example, all emissions data will be bounded by 
0 and will often have an upper bound as well. Due to the strictly positive 
nature of emissions data, a gaussian kernel is not often a good choice. There 
are numerous types of empirical density functions in R and density estimation is 
an extensive body of research in its own right. Given an appropriate choice of 
kernel and bandwidth, the results using different methods are very similar. 

We recommend the `obs_density()` function in `UPLforOAR`, which builds 
on the `np::npuniden.boundary()` function, a more accurate density function for 
bounded data. By default, `obs_density()` will assume a lower bound of 0, no 
upper bound, and the `kernel` type is gamma. Other kernel types accepted by 
`np::npuniden.boundary()` can be used, for example if data are percent removals 
instead of emissions then `"beta1"` or `"beta2"` with an upper bound of 1 are 
more appropriate. The default bandwidth `bw` is $\sigma*n^{-2/5}$ (Renault & Scaillet 2004), 
though manual values or a least-squares or likelihood search can be used 
instead by setting `bw = "cv.ls"` or `bw = "cv.ml"`.

The results of `obs_density()` include `ObsonPoint`, with the observation densities 
paired with each emission value in the provided data set, and `obs_den_df`, a 
sequence of densities across a range of possible emission values (`xvals`) based on the 
observed emissions. The range of emissions `xvals` can be supplied manually, and 
by default is `seq(0,3*max(data$emissions,length.out=1024))`. 

## Bandwidth and kernel options 

Below are some examples of `obs_denisty()` with different bandwidths and 
different kernels. For these comparisons, we are using a known gamma distribution 
to generate the sample of emissions data, then comparing the empirical densities 
to the actual probability density function. There is little difference between 
the default calculated bandwidth and the two options to search for the correct 
bandwidth using least squares or maximum likelihood. The manual option we 
supplied is also similar, though slightly over fit. 

Run `?obs_density` in the console for detailed information in the help pages.

```{r obs_example, echo=TRUE}
# Generate some data to use:
set.seed(1)
xval = seq(0, 10, length.out = 1000)
pval = dgamma(xval, shape = 2, rate = 0.65)
yval = rgamma(100, shape = 2, rate = 0.65)
exmpl_dat = tibble(xval = xval, pval = pval)
exmpl_emiss = tibble(emissions = yval)
# Empirical densities using different bandwidths:
# default settings
dens1 = obs_density(data = exmpl_emiss, xvals = exmpl_dat$xval) 
dens_pts1 = dens1$Obs_onPoint
dens_vals1 = dens1$obs_den_df
# manual bandwidth
dens2 = obs_density(data = exmpl_emiss, xvals = exmpl_dat$xval, bw = 0.1)
dens_pts2 = dens2$Obs_onPoint
dens_vals2 = dens2$obs_den_df
# least squares search
dens3 = obs_density(data = exmpl_emiss, xvals = exmpl_dat$xval, bw = "cv.ls") 
dens_pts3 = dens3$Obs_onPoint
dens_vals3 = dens3$obs_den_df
# maximum likelihood search
dens4 = obs_density(data = exmpl_emiss, xvals = exmpl_dat$xval, bw = "cv.ml")
dens_pts4 = dens4$Obs_onPoint
dens_vals4 = dens4$obs_den_df
```
```{r figure4, echo=FALSE, out.width="100%", fig.height=4, fig.align = "center", fig.cap="Examples of empirical densities using different bandwidth options. The actual density distribution used to simulate the data is in black, and various empirical densities overlaid in color."}
dens_vals1$type = rep('Default', nrow(dens_vals1))
dens_vals2$type = rep('Manual bw = 0.1', nrow(dens_vals2))
dens_vals3$type = rep('Least squares', nrow(dens_vals3))
dens_vals4$type = rep('Maximum likelihood', nrow(dens_vals3))
dens_vals = rbind(dens_vals1, dens_vals2, dens_vals3, dens_vals4)

dens_pts1$type = rep('Default', nrow(dens_pts1))
dens_pts2$type = rep('Manual bw = 0.1', nrow(dens_pts2))
dens_pts3$type = rep('Least squares', nrow(dens_pts3))
dens_pts4$type = rep('Maximum likelihood', nrow(dens_pts4))
dens_pts = rbind(dens_pts1, dens_pts2, dens_pts3, dens_pts4)

palette_line=c('black',"#377EB8", "#4DAF4A", "#984EA3", "#E41A1C")

ggplot(data = dens_vals)+
  geom_line(data = exmpl_dat, aes(y = pval, x = xval, color = "black"), 
            size = 0.75)+
  geom_area(data = exmpl_dat, aes(y = pval, x = xval, fill = "black"), 
            alpha = 0.25)+  
  geom_line(data = dens_vals, aes(y = ydens, x = x_hat, color = type), 
            size = 0.75)+
  geom_area(data = dens_vals, aes(y = ydens, x = x_hat, fill = type), 
            alpha = 0.25)+
  geom_point(aes(y = ydens, x = emissions, color = type), data = dens_pts,
             size = 3, alpha = 0.5, shape = 19)+
  ylab("Density")+xlab("Emissions")+
  ggtitle("Bandwidth Comparisons")+labs(color='Bandwidth',fill="Bandwidth")+
  multi_source_theme()+theme(panel.spacing.x = unit(4, "mm"))+
  scale_color_manual(values=palette_line, labels = c("Actual", "Default", "LS",
                                                     "Manual", "ML"))+
  scale_fill_manual(values=palette_line, labels = c("Actual", "Default", "LS",
                                                     "Manual", "ML"))+
  scale_x_continuous(expand = expansion(mult = c(0,0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0,0.05)))+
  facet_wrap(~type)+theme(strip.text = element_text(color='black',size=12))
```

Next we can change the type of kernel used: 

```{r kernel, echo = TRUE}
# Empirical densities using different kernels:
# default settings (Gamma kernel)
dens1 = obs_density(data = exmpl_emiss, xvals = exmpl_dat$xval) 
dens_pts1 = dens1$Obs_onPoint
dens_vals1 = dens1$obs_den_df
# Gaussian kernel
dens2 = obs_density(data = exmpl_emiss, xvals = exmpl_dat$xval, 
                    kernel = "gaussian1")
dens_pts2 = dens2$Obs_onPoint
dens_vals2 = dens2$obs_den_df
# Floating boundary kernel
dens3 = obs_density(data = exmpl_emiss, xvals = exmpl_dat$xval,
                    kernel = "fbl") 
dens_pts3 = dens3$Obs_onPoint
dens_vals3 = dens3$obs_den_df
# Reciprocal inverse Gaussian
dens4 = obs_density(data = exmpl_emiss, xvals = exmpl_dat$xval,
                    kernel = "rigaussian")
dens_pts4 = dens4$Obs_onPoint
dens_vals4 = dens4$obs_den_df
```

Note that the Gaussian kernel is likely a poor choice since we used a strictly 
positive distribution to simulate the data. The floating boundary kernel option 
is warning us that the density curve doesn't integrate to 1, meaning we should 
try a different bandwidth or don't use this kernel type. Consequently, in the 
figure below, it is clear that both the Gaussian and floating boundary look poor 
recreations of the actual gamma distribution used to simulate the data.

```{r figure6, echo=FALSE, out.width="100%", fig.height=4, fig.align = "center", fig.cap="Examples of empirical densities using different kernel options. The actual density distribution used to simulate the data is in black, and various empirical densities overlaid in color."}
dens_vals1$type = rep('Default (Gamma)', nrow(dens_vals1))
dens_vals2$type = rep('Gaussian', nrow(dens_vals2))
dens_vals3$type = rep('Floating boundary', nrow(dens_vals3))
dens_vals4$type = rep('Reciprocal inverse Gaussian', nrow(dens_vals3))
dens_vals = rbind(dens_vals1, dens_vals2, dens_vals3, dens_vals4)

dens_pts1$type = rep('Default (Gamma)', nrow(dens_pts1))
dens_pts2$type = rep('Gaussian', nrow(dens_pts2))
dens_pts3$type = rep('Floating boundary', nrow(dens_pts3))
dens_pts4$type = rep('Reciprocal inverse Gaussian', nrow(dens_pts4))
dens_pts = rbind(dens_pts1, dens_pts2, dens_pts3, dens_pts4)

ggplot(data = dens_vals)+
  geom_line(data = exmpl_dat, aes(y = pval, x = xval, color = "black"), 
            size = 0.75)+
  geom_area(data = exmpl_dat, aes(y = pval, x = xval, fill = "black"), 
            alpha = 0.25)+   
  geom_line(data = dens_vals, aes(y = ydens, x = x_hat, color = type), 
            size = 0.75)+
  geom_area(data = dens_vals, aes(y = ydens, x = x_hat, fill = type), 
            alpha = 0.25)+
  geom_point(aes(y = ydens, x = emissions, color = type), data = dens_pts,
             size = 3, alpha = 0.5, shape = 19)+
  ylab("Density")+xlab("Emissions")+
  ggtitle("Kernel comparisons")+labs(color='Kernel',fill="Kernel")+
  multi_source_theme()+theme(panel.spacing.x = unit(4, "mm"))+
  scale_color_manual(values=palette_line, labels = c("Actual", "Default", 
                                                     "Gaussian",
                                                     "FB", "RIG"))+
  scale_fill_manual(values=palette_line, labels = c("Actual", "Default", 
                                                    "Gaussian",
                                                      "FB", "RIG"))+
  multi_source_theme()+theme(panel.spacing.x = unit(4, "mm"))+
  scale_x_continuous(expand = expansion(mult = c(0,0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0,0.05)))+
  coord_cartesian(clip = 'off')+facet_wrap(~type)+
  theme(strip.text = element_text(color='black',size=12))
```
