---
title: "Distributions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Distributions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  error=TRUE,
  warning=FALSE,
  message=FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
library(htmltools)
knitr::knit_hooks$set(imgcenter = function(before, options, envir){
  if (before) {
    HTML("<p align='center'>")
  } else {
    HTML("</p>")
  }
})
```

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(forcats)
library(UPLforOAR)
library(matrixStats)
library(EnvStats)
library(np)
library(sfsmisc)
library(parallel)
library(rjags)
library(runjags)
library(sn)
library(patchwork)
```

## Introduction

Vignette about distributions and things you should know to help pick the best one.

Include some useful links to JAGS and BUGS distribution sites, as well as how to code a distribution manually.

## Long- and Heavy-Tailed Distributions 

For example some Lognormal. The 99<sup>th</sup> percentile can be unreasonably large 
emission number when long or heavy tail. This is often an issue when the 
observations themselves are no where near the tail, and thus the tail is largely 
unsupported by any data. In such circumstances, the UPL can be a value larger 
than any in the observation data set, or larger than is physically possible. In 
such cases it is necessary to truncate the probability distribution to physically 
reasonable bounds. This is done by default as 3 times the maximum emissions value 
in the training data, but can be adjusted by setting `maxY`. 

*show example of two distributions, one long-tailed one not, where maxY is varied*





