---
title: "Distributions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Distributions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  error=TRUE,
  warning=FALSE,
  message=FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
library(htmltools)
knitr::knit_hooks$set(imgcenter = function(before, options, envir){
  if (before) {
    HTML("<p align='center'>")
  } else {
    HTML("</p>")
  }
})
```

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(forcats)
library(UPLforOAR)
library(matrixStats)
library(np)
library(sfsmisc)
library(parallel)
library(rjags)
library(runjags)
library(sn)
library(patchwork)
```

## Introduction

Picking a probability distribution to represent your emissions data is a 
necessary step to calculating a UPL, regardless of the method you use or any 
other parameters you change. If you are using the frequentist methods behind 
`Normal_UPL()`, `Lognormal_UPL()`, and `Skewed_UPL()`, than you are picking the 
distribution based on ratios of skewness and kurtosis and it is explicitly assumed 
to be correct in the UPL calculations. If you are using the `Bayesian_UPL()` 
method, then you can compare as many distributions as you like and choose the best 
based on quantitative and qualitative comparisons to the actual observation 
density. At the moment, the `UPLforOAR` package supports Beta, Gamma, Lognormal, 
Normal, and Skewed for `Bayesian_UPL()`. However, there are many more 
distributions already supported in R and JAGS (also known as BUGS, a slightly 
different version of the same software). This vignette describes the supported 
distributions and links to resources for more that can be included in future 
development as needed. Much of the content on distributions comes from either [BUGS documentation](https://www.multibugs.org/documentation/latest/Distributions.html) 
or the reference book "Bayesian Models: A Statistical Primer for Ecologists" by N. 
Thompson Hobbs and Mevin B. Hooten, Princeton University Press 2015.

### Beta

The Beta distribution models random variables that can only take on values 
between 0 and 1. It is often used for probabilities and proportions, and is a 
good choice to consider for emissions variables that are represented as 
destruction or removal efficiencies. The probability density, $p(x)$, of Beta 
distributed emissions, $x$, is dependent on parameters $\alpha$ and $\beta$:

$$
p(x) = \frac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}x^{\alpha -1}(1-x)^{\beta-1}
$$

Note that $\Gamma$ in this case is the gamma *function* not the Gamma probability 
distribution. Using moment-matching, you can calculate the mean and variance of 
the Beta distribution as:

$$
\mu = \frac{\alpha}{\alpha + \beta},~~ \sigma^{2} = \frac{\alpha \beta}{(\alpha + \beta)^{2}(\alpha + \beta +1)}
$$

The parameters $\alpha$ and $\beta$ can be positive and negative, though shouldn't 
be able to take on values below -1. In JAGS the distribution is defined as 
$\sim dbeta(\alpha, ~\beta)$.

### Gamma

The Gamma distribution is very useful for emissions data because it represents 
strictly positive skewed continuous data. The probability density, $p(x)$, of 
Gamma distributed emissions, $x$, is dependent on parameters $\alpha$, the shape, 
and $\beta$, the rate:

$$
p(x) = \frac{\beta^{\alpha}}{\Gamma(\alpha)}x^{\alpha-1}e^{-\beta x}
$$

Note that $\Gamma$ in this case is the gamma *function* not the Gamma probability 
distribution. Both the *shape* and *rate* must be positive real numbers. Using 
moment-matching, you can calculate the mean and variance of the Gamma distribution as:

$$
\mu = \frac{\alpha}{\beta},~~\sigma^{2} = \frac{\alpha}{\beta^{2}}
$$

In JAGS the distribution is defined as $\sim dgamma(\alpha, ~\beta)$.

### Lognormal

If a variable $z$ is Normally distributed, then $e^z$ is Lognormally distributed. 
If a variable $x$ is Lognormally distributed, than a variable $log(x)$ is 
Normally distributed. The product of large numbers of random variables is 
Lognormally distributed regardless of the distributions of the individual variables. 
It is often used to model growth or products of states or parameters. It is 
continuous and strictly positive, and can be asymmetric, which can be well suited 
to describe emissions data. The probability for the Lognormal distribution is 
dependent on parameters $\mu$, the mean of $log(x)$ and $\sigma$, the standard 
deviation of $log(x)$:

$$
p(x) = \frac{1}{x\sqrt{2\pi \sigma^{2}}}e^{-\frac{(log(x)-\mu)^{2}}{2\sigma^2}}
$$

Using moment-matching, you can calculate the mean and variance of the Lognormal 
distribution as:
$$
mean = e^{\mu+\sigma^2/2}, ~~variance = (e^{\sigma^2}-1)e^{2\mu+\sigma^2}
$$

Note that the variance and mean are interdependent, which makes the Lognormal 
particularly useful in cases where the variance of the data increases with the 
mean. In JAGS the distribution is defined as $\sim dlnorm(\mu, ~\frac{1}{\sigma^{2}})$. 
Note that rather than variance like R uses, JAGS takes the precision as the 
second input. There is no analytical solution for the arithmetic mean of future 
samples from the Lognormal distribution, so the UPL has to be calculated using an 
approximation in `Lognormal_UPL()` based on Gram-Charlier Series-A expansion 
(for more details see "An upper prediction limit for the arithmetic mean of a 
lognormal random variable" by Dulal Kumar Bhaumik and Robert David Gibbons, 2004.). 
Using the Bayesian methods in `Bayesian_UPL()`, the UPL can be calculated the 
same as it would be for any distribution; by drawing the future samples from the 
posterior distributions. Given that the Lognormal distribution is often 
heavy-tailed, and that the 99<sup>th</sup> UPL is by definition going to be in 
that tail, the distribution will often need to be truncated by some upper bound 
such that emissions beyond it unreasonable to consider. See the section below on 
heavy and long tailed distributions for more information.

### Normal

The Normal distribution applies to continuous random variables across the entire 
number line, and has been widely used in statistics because it has properties 
allowing many results to be derived analytically. In addition, because the sum of 
a large number of samples from any distribution will be normally distributed, 
there are many situations it is suited to. However, the Normal distribution is not 
often a good choice for emissions data, which tend not to be symmetric, are 
strictly positive, and the variance is often related to the mean. The probability 
density, $p(x)$, of Normally distributed emissions, $x$, is dependent on 
parameters $\mu$, the mean, and $\sigma$, the standard deviation:

$$
p(x) = \frac{1}{\sigma \sqrt{2\pi}}e^{-\frac{(x-\mu)^{2}}{2\sigma^{2}}}
$$

In JAGS the distribution is defined as $\sim dnorm(\mu, ~\frac{1}{\sigma^{2}})$. 
Note that rather than variance like R uses, JAGS takes the precision as the second 
input. Since the mean and variance can be solved for analytically, `Normal_UPL()` 
uses these fixed-point estimates to calculate the UPL based on the desired number 
of future samples and significance. The Bayesian method employed in `Bayesian_UPL()` 
will arrive at the same estimates of mean, variance, and UPL, but will incorporate 
uncertainty in the mean and variance and offers more ways to evaluate if the 
Normal distribution is the best choice to represent the data.

### Skewed

The Skewed or skew-normal distribution is a generalization of the Normal 
distribution that allows for skewness. As such, it also exists the entire number 
line, but it is not symmetric. The Skew-normal distribution is defined by three 
parameters; location $\xi$, scale $\omega$, and skewness $\alpha$. This is 
equivalent to a Normal distribution when $\alpha = 0$, it is right-skewed if 
$\alpha > 0$, and left-skewed if $\alpha < 0$. The probability 
density, $p(x)$, of Skewed distributed emissions is:

$$
p(x) = \frac{2}{\omega}\phi \left(\frac{x-\xi}{\omega}\right) \Phi \left(\alpha\left(\frac{x-\xi}{\omega}\right)\right)
$$

Where $\phi(x)$ is the standard normal probability distribution, $\Phi(x)$ 
is its cumulative probability distribution function, and $x$ is transformed to 
$\frac{x-\xi}{\omega}$. Using the error function, this can be expanded to:

$$
p(x) = \frac{2}{\omega}\frac{1}{\sqrt{2\pi}}e^{-\frac{\left(\frac{x-\xi}{\omega}\right)^{2}}{2}}\frac{1}{2}\left[1+\frac{2}{\sqrt{\pi}}\int_0^{\alpha\frac{x-\xi}{\omega\sqrt{2}}}e^{-t^2}dt\right]
$$

Unlike the Normal distribution, there is no closed-form expression for $\alpha$, 
$\xi$, and $\omega$. The method of moments can be used for approximations when 
sample skewness is not too large, which is useful for determining initial values 
for MCMC chains for these parameters, but not as a solution for parameterization. 
Since there is no analytical solution, the approach in `Skewed_UPL()` uses 
iterative shifting of the *t*-statistic until the desired significance is achieved.

The Skew-normal distribution is not included in JAGS, but it is in R via the `sn` 
package. However, any distribution can be coded manually in JAGS by using the 
'zero poisson' [method](https://www.multibugs.org/documentation/latest/Tricks.html#GenericDistribution). 
This approach is equivalent to any standard distribution in how the parameters 
are solved for using MCMC and posterior distributions are returned. The main 
difference in implementation is that for the standard distributions already 
defined in JAGS, we can draw the 'future' samples and probability density at any 
emission observation value from within JAGS, whereas with a manually defined 
distribution like Skew-normal we need to use the posterior distributions of $\xi$, 
$\omega$, and $\alpha$ to define the probability density in R using `sn::dsn()` 
and take those future draws using `sample()`. There is no difference in results 
either way as long as we are controlling for stochasticity. 

## Long- and Heavy-Tailed Distributions 

Some probability distributions, for example some Lognormal, can have long or 
heavy tails. This means that there is some tiny probability of extreme values 
that can occur, and the range of these extreme values, i.e. the tail, can be 
very long. Where exactly the tail ends is quite uncertain because we usually 
don't have any data supporting the probability estimates along these tails. The 
vast majority of the probability and emission values are in the body of the 
distribution, which is well supported by the parameter estimates and data. To 
put it another way, you can have very similar mean and standard deviations 
defining two Lognormal distributions, and the bulk of their probability density 
curves will be very similar, but the extent of the tails might be quite different. 
Furthermore, any method that involves a stochastic component in solving for 
parameters (for example MCMC in JAGS) can end up with very different extents of 
the tails due to randomness, rather than as a result of process or data. While 
we can control the randomness to have exactly reproducible results, we still want 
the stochastic component of any Bayesian method to be minuscule in its 
contribution to the final result.

This matters quite a lot when the UPL is the 99<sup>th</sup> percentile of a 
long-tailed distribution, because it is going to be within that tail. The UPL of 
a long-tailed distribution, whether it's solved for using `Bayesian_UPL()` or 
`Lognormal_UPL()`, can end up being unrealistically large. Sometimes the 
99<sup>th</sup> percentile of a long-tailed distribution will be past the range 
of all emissions data, not just the top performers used to calculate the UPL, and 
sometimes it might be beyond what is physically possible. These are obvious 
situations where the UPL of such a long-tailed distribution should not be used, 
and either a different distribution should be used or measures should be taken to 
truncate the tail to physically reasonable bounds. This is done by default in 
`Bayesian_UPL()` with an upper limit of 3 times the maximum emissions value in 
the training data, but can be adjusted by setting `maxY`. Any distribution that 
does not have an unsupported long-tail will be unaffected by this setting.

### Example

Let's look at an example using HCl emissions from Lime processing. First we will 
load the data and use the Excel workbook methods to pick a distribution and 
calculate the UPL.
```{r lime, echo = TRUE}
dat_emiss = read.csv("Example_data1.csv")
distr_choice = distribution_type(data = dat_emiss, emissions = 'emissions')
UPL_excel = Lognormal_UPL(data = dat_emiss, emissions = 'emissions')
```
The emissions are a `r distr_choice` distribution with a corresponding UPL of 
`r round(UPL_excel,2)` (Lb/ton of stone fed Lime). Next we will plot the density of the 
emission observations along with the corresponding `r distr_choice` probability 
density distribution to see how it looks.
```{r density, echo = FALSE}
x_hat = seq(0, 2.6, length.out = 1024)
obs_dens_results = obs_density(dat_emiss, xvals = x_hat, 
                               emissions = 'emissions')
Obs_onPoint = obs_dens_results$Obs_onPoint
obs_den_df = obs_dens_results$obs_den_df
pdf_ln = dlnorm(x_hat,mean = log(mean(dat_emiss$emissions, na.rm = T)),
              sd = sd(log(dat_emiss$emissions), na.rm = T))
pred_dat = tibble(x_hat, pdf_ln)
```
```{r figure1, echo = TRUE, fig.dim = c(5, 4), fig.align = "center", fig.cap = "Observation density of HCl emissions. The obseration data are indicated in black as points and a rug along the axis, with the observation density distribution as a black line. The fitted lognormal distribution that is the basis of the UPL estimate is colored orange. The UPL result is the vertical orange line."}
ggplot()+
  geom_line(data = obs_den_df, size=0.75,
            aes(y = ydens, x = x_hat, color = 'a'))+
  geom_area(data = obs_den_df, alpha=0.25,
            aes(y = ydens, x = x_hat, fill = 'a'))+
  geom_point(aes(y = ydens, x = emissions), data = Obs_onPoint,
             size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_line(aes(y = pdf_ln, x = x_hat, color = 'c'),
               data = pred_dat, size = 0.75, linetype = 3)+
  geom_area(aes(y = pdf_ln, x = x_hat, fill = 'c'), alpha = 0.25,
            data = pred_dat)+
  ylab("Density")+xlab("HCl emissions (lb/ton stone fed lime)")+
  ggtitle("Excel-based distribution and UPL")+
  pop_distr_theme()+
  geom_vline(aes(xintercept = UPL_excel, color = 'c'), 
             linewidth = 1, linetype = 1)+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  coord_cartesian(clip = 'off')+
  labs(color = 'Density:', fill = 'Density:')+
  geom_rug(sides = 'b', aes(x = emissions), data = dat_emiss,
           alpha = 0.5, outside = TRUE, color = 'black')+
  scale_color_manual(values = c('black', '#FF7F00'),
                     labels = c('Observations', 'Lognormal'))+
  scale_fill_manual(values = c('black', '#FF7F00'),
                    labels = c('Observations', 'Lognormal'))
```
It is clear why the distribution is Lognormal over Normal or Skewed, since is a 
high density of very low emissions near zero that needs to be accommodated by 
the probability density distribution. However, it has a very long tail that 
results in an unreasonably large UPL that is `r round(UPL_excel/max(dat_emiss$emissions),1)` 
times larger than the largest emission. Instead, we will use a Bayesian approach 
that can do a better job representing the emissions.

### Again with Bayes

Normally we would investigate multiple distributions because it is possible that 
a Gamma or a Beta might be better than the Lognormal and not have such a long 
tail. But for the sake of demonstrating how the tail affects the UPL, we will 
stick with the Lognormal. The methods in `Bayesian_UPL()` allow you to set a value at 
which to truncate the upper limit of emission probabilities, called `maxY`. By 
default, this is `3*max(data$emissions)`. We will also look setting `maxY` to `2*max(data$emissions)` and `25*max(data$emissions)` to see how the UPL estimate 
changes.
```{r maxy, echo = TRUE, results = FALSE}
results_3X = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                          emissions = 'emissions',
                          xvals = seq(0, 3, length.out = 1000))
results_2X = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                          emissions = 'emissions',
                          maxY = 2*max(dat_emiss$emissions),
                          xvals = seq(0, 3, length.out = 1000))
results_25X = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                           emissions = 'emissions',
                           maxY = 25*max(dat_emiss$emissions),
                           xvals = seq(0, 3, length.out = 1000))
```
```{r figure2, echo = FALSE, fig.height = 7, fig.width = 5, fig.align = "center", fig.cap = "Example of three Lognormal distributions fit to the same data, but with different maximum emission thresholds."}
results_3X$obs_pdf_dat$distr = rep("3X Max", 
                                          nrow(results_3X$obs_pdf_dat))
results_3X$pred_pdf_dat$distr = rep("3X Max", 
                                          nrow(results_3X$pred_pdf_dat))
results_3X$fit_table$distr = rep("3X Max", 
                                          nrow(results_3X$fit_table))
results_2X$obs_pdf_dat$distr = rep("2X Max", 
                                          nrow(results_2X$obs_pdf_dat))
results_2X$pred_pdf_dat$distr = rep("2X Max", 
                                          nrow(results_2X$pred_pdf_dat))
results_2X$fit_table$distr = rep("2X Max", 
                                          nrow(results_2X$fit_table))
results_25X$obs_pdf_dat$distr = rep("25X Max", 
                                          nrow(results_25X$obs_pdf_dat))
results_25X$pred_pdf_dat$distr = rep("25X Max", 
                                          nrow(results_25X$pred_pdf_dat))
results_25X$fit_table$distr = rep("25X Max", 
                                          nrow(results_25X$fit_table))
combined_pdf = rbind(results_2X$pred_pdf_dat, 
                     results_3X$pred_pdf_dat, 
                     results_25X$pred_pdf_dat)
combined_obs = rbind(results_2X$obs_pdf_dat, 
                     results_3X$obs_pdf_dat, 
                     results_25X$obs_pdf_dat)
combined_fit = rbind(results_2X$fit_table, 
                     results_3X$fit_table, 
                     results_25X$fit_table)
combined_pdf$distr = factor(combined_pdf$distr, 
                            levels = c("2X Max", "3X Max", "25X Max"))
combined_obs$distr = factor(combined_obs$distr, 
                            levels = c("2X Max", "3X Max", "25X Max"))
combined_fit$distr = factor(combined_fit$distr, 
                            levels = c("2X Max", "3X Max", "25X Max"))

palette_line = c("#377EB8", "#E41A1C", "#4DAF4A")
ggplot()+
  geom_line(data = combined_pdf,
            aes(y = ydens, x = x_hat), size = 0.75, color = 'black')+
  geom_area(data = combined_pdf,
            aes(y = ydens, x = x_hat), fill = 'grey')+
  geom_point(data = combined_obs,
            aes(y = ydens, x = emissions),
            size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_line(data = combined_pdf,
            aes(x = x_hat, y = pdf_hat, color = distr), size = 0.75)+
  geom_area(data=combined_pdf,
            aes(x=(x_hat),y=pdf_hat,fill=distr),alpha=0.5)+
  geom_errorbar(data = combined_obs,
                aes(ymin = low, ymax = up, x = emissions, color = distr),
                alpha = 0.5)+
  geom_point(data = combined_obs,
             aes(x = emissions, y = med, fill = distr),
             size = 3, alpha = 0.5, shape = 21, color = 'black')+
  geom_vline(aes(xintercept = UPL, color = distr), data = combined_fit,
             linewidth = 1)+
  ylab("Density")+ xlab("Emissions")+
  pop_distr_theme()+ labs(color = 'Distribution', fill = "Distribution")+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 3.1))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  coord_cartesian(clip = 'off')+ scale_color_manual(values = palette_line)+
  scale_fill_manual(values = palette_line)+
  facet_wrap(~distr, scales = 'fixed', nrow = 3)+
  theme(legend.position = 'none')+
  theme(strip.text = element_text(color = 'black', size = 12))
```
Clearly, the longer the tail the larger the UPL: when the tail limit increases 
from `r round(2*max(dat_emiss$emissions),2)`, to `r round(3*max(dat_emiss$emissions),2)`, 
and then `r round(10*max(dat_emiss$emissions),2)` we get UPL estimates of `r round(results_2X$fit_table$UPL,2)`, `r round(results_3X$fit_table$UPL,2)`, and `r round(results_25X$fit_table$UPL,2)`. However, not all of these results are equal 
in terms of performance. 

Let's look at the results from `fit_likelihood()` to evaluate which distribution 
did a better job representing the actual observations. The distribution with the 
lowest upper limit of `r round(2*max(dat_emiss$emissions),2)` did the best at 
actually fitting the data: it has the lowest SSE and the most observations within 
the 95\% CI around predicted probability densities. The next higher limit had as 
many observations in agreement but a higher SSE. Lastly, the distribution with an 
extremely high upper limit that is unlikely to cause any truncation of 
probabilities, has the worst fit in terms of substantially higher SSE and more 
than half of the observations are outside the 95\% CI on predicted density. Note 
that in all cases the area under the curve sums to less than 1, but the more the 
probability density function is truncated, the farther below 1.
```{r fit1, echo = FALSE}
fit_metrics = arrange(combined_fit, SSE)
fit_metrics = mutate(fit_metrics, across(c(2, 3, 5), ~signif(., digits = 2)))
knitr::kable(fit_metrics,
             caption = "Performance of Lognormal distributions with different MaxY limits",
             col.names = c("Upper Limit", 'UPL', 'SSE', 'No. Obs. in 95% CI', 'pdf integral'))
```

Next we ought to check the convergence of all of the parameters. The model with 
`maxY = 3*max(data$emissions)` has one parameter with weak convergence according 
to the Gelman-Rubin diagnostics, so we will want to check all of the convergence 
figures. Issues with convergence can indicate that a distribution is not a good 
choice to represent the emissions data. The parameters technically passed the 
Gelman-Rubin diagnostic checks for all of them. However, for the 2X and 3X limits 
they do not look as well mixed as we would like, and you can see the histograms 
aren't as Gaussian as can expect with good convergence. It is not so bad as to 
invalidate the results, but you can imagine that if we tried limits that were 
too small the model would fail.
```{r figure3a, fig.height = 4, out.width = "100%", fig.align = "center", results = FALSE, fig.cap = "Convergence figures from maxY = 2X."}
part1 = setup_likelihood(distribution = "Lognormal", data = dat_emiss,
                         emissions = 'emissions')
part2 = run_likelihood(model_input = part1, maxY = 2 * max(dat_emiss$emissions),
                          xvals = seq(0, 3, length.out = 1000))
part3 = converge_figs(distribution = "Lognormal", jags_model_run = part2)
part3
```

```{r figure3b, fig.height = 4, out.width = "100%", fig.align = "center", results = FALSE, fig.cap = "Convergence figures from maxY = 3X."}
part1 = setup_likelihood(distribution = "Lognormal", data = dat_emiss,
                         emissions = 'emissions')
part2 = run_likelihood(model_input = part1, maxY = 3 * max(dat_emiss$emissions),
                          xvals = seq(0, 3, length.out = 1000))
part3 = converge_figs(distribution = "Lognormal", jags_model_run = part2)
part3
```
```{r figure3c, fig.height = 4, out.width = "100%", fig.align = "center", results = FALSE, fig.cap = "Convergence figures from maxY = 25X."}
part1 = setup_likelihood(distribution = "Lognormal", data = dat_emiss,
                         emissions = 'emissions')
part2 = run_likelihood(model_input = part1, maxY = 25 * max(dat_emiss$emissions),
                          xvals = seq(0, 10, length.out = 3000))
part3 = converge_figs(distribution = "Lognormal", jags_model_run = part2)
part3
```
Picking an appropriate upper limit for `maxY` might not be an easy choice, and 
you should always vary it to check if you think your UPL results are in an 
unsupported long-tail. The alternative to using `maxY` to truncate the probability distribution is to use a different significance level. One that is lower than 
0.99 might not be sensitive to the tail, and will be much less affected by u
ncertainty and stochasticity in the tail of the distribution.

### Alternative significance levels

Let's investigate the sensitivity of different UPL significance levels to `maxY`. Ideally we will find a level that is insensitive to `maxY` and converges well. 
```{r sig, echo = TRUE, results = FALSE}
results_3X95 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                            significance = 0.95, emissions = 'emissions',
                            xvals = seq(0, 3, length.out = 1000))
results_2X95 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                            significance = 0.95, emissions = 'emissions',
                            maxY = 2*max(dat_emiss$emissions),
                            xvals = seq(0, 3, length.out = 1000))
results_25X95 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                            maxY = 25*max(dat_emiss$emissions),
                            significance = 0.95, emissions = 'emissions',
                            xvals = seq(0, 3, length.out = 1000))
results_3X90 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                            significance = 0.90, emissions = 'emissions',
                            xvals = seq(0, 3, length.out = 1000))
results_2X90 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                            significance = 0.90, emissions = 'emissions',
                            maxY = 2*max(dat_emiss$emissions),
                            xvals = seq(0, 3, length.out = 1000))
results_25X90 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                             significance = 0.90, emissions = 'emissions',
                             maxY = 25*max(dat_emiss$emissions),
                             xvals = seq(0, 3, length.out = 1000))
results_3X85 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                            significance = 0.85, emissions = 'emissions',
                            xvals = seq(0, 3, length.out = 1000))
results_2X85 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                            significance = 0.85, emissions = 'emissions',
                            maxY = 2*max(dat_emiss$emissions),
                            xvals = seq(0, 3, length.out = 1000))
results_25X85 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                             significance = 0.85, emissions = 'emissions',
                             maxY = 25*max(dat_emiss$emissions),
                             xvals = seq(0, 3, length.out = 1000))
results_3X80 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                            significance = 0.80, emissions = 'emissions',
                            xvals = seq(0, 3, length.out = 1000))
results_2X80 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                            significance = 0.80, emissions = 'emissions',
                            maxY = 2*max(dat_emiss$emissions),
                            xvals = seq(0, 3, length.out = 1000))
results_25X80 = Bayesian_UPL(distr_list = c("Lognormal"), data = dat_emiss,
                             significance = 0.80, emissions = 'emissions',
                             maxY = 25*max(dat_emiss$emissions),
                             xvals = seq(0, 3, length.out = 1000))
```
The fitted probability density distribution won't change with significance level, just the resulting UPL. The figure below shows the same distributions as before with the different `maxY` limits, but now we've added the UPL when calculated using 80\%, 85\%, 90\%, 95\% and 99\% significance as vertical lines. As expected, lower significance has a lower UPL. The important outcome here is that at 80\% and 85\% significance the UPL doesn't change no matter how we vary `maxY`. The UPL changes slightly at the highest `maxY` for 90\%, and quite a lot as we increase `maxY` for 95\% and 99\% significance. 
```{r figure4, echo = FALSE, fig.height = 7, fig.width = 5, fig.align = "center", fig.cap = "Example of three Lognormal distributions fit to the same data, but with different maximum emission thresholds and levels of significance. THe significnace only changes the UPL outcome, and each level has a different line type as a vertical line at the UPL."}
results_3X95$fit_table$distr = rep("3X Max", 
                                          nrow(results_3X95$fit_table))
results_2X95$fit_table$distr = rep("2X Max", 
                                          nrow(results_2X95$fit_table))
results_25X95$fit_table$distr = rep("25X Max", 
                                          nrow(results_25X95$fit_table))
results_3X90$fit_table$distr = rep("3X Max", 
                                          nrow(results_3X90$fit_table))
results_2X90$fit_table$distr = rep("2X Max", 
                                          nrow(results_2X90$fit_table))
results_25X90$fit_table$distr = rep("25X Max", 
                                          nrow(results_25X90$fit_table))
results_3X85$fit_table$distr = rep("3X Max", 
                                          nrow(results_3X85$fit_table))
results_2X85$fit_table$distr = rep("2X Max", 
                                          nrow(results_2X85$fit_table))
results_25X85$fit_table$distr = rep("25X Max", 
                                          nrow(results_25X85$fit_table))
results_3X80$fit_table$distr = rep("3X Max", 
                                          nrow(results_3X80$fit_table))
results_2X80$fit_table$distr = rep("2X Max", 
                                          nrow(results_2X80$fit_table))
results_25X80$fit_table$distr = rep("25X Max", 
                                          nrow(results_25X80$fit_table))
sig95 = rbind(results_2X95$fit_table, results_3X95$fit_table, 
              results_25X95$fit_table)
sig95$Significance = rep("95th", nrow(sig95))
sig90 = rbind(results_2X90$fit_table, results_3X90$fit_table, 
              results_25X90$fit_table)
sig90$Significance = rep("90th", nrow(sig90))
sig85 = rbind(results_2X85$fit_table, results_3X85$fit_table, 
              results_25X85$fit_table)
sig85$Significance = rep("85th", nrow(sig85))
sig80 = rbind(results_2X80$fit_table, results_3X80$fit_table, 
              results_25X80$fit_table)
sig80$Significance = rep("80th", nrow(sig80))
combined_fit2 = rbind(sig95, sig90, sig85, sig80)

combined_fit2$distr = factor(combined_fit2$distr,
                             levels = c("2X Max", "3X Max", "25X Max"))

combined_fit$Significance = rep("99th", nrow(combined_fit))

combined_fit_all = rbind(combined_fit, combined_fit2)

combined_fit_all$Significance = factor(combined_fit_all$Significance,
                                     levels = c("99th", "95th", "90th",
                                                "85th",'80th'))

palette_line = c("#377EB8", "#E41A1C", "#4DAF4A")
ggplot()+
  geom_line(data = combined_pdf,
            aes(y = ydens, x = x_hat), size = 0.75, color = 'black')+
  geom_area(data = combined_pdf,
            aes(y = ydens, x = x_hat), fill = 'grey')+
  geom_point(data = combined_obs,
            aes(y = ydens, x = emissions),
            size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_line(data = combined_pdf,
            aes(x = x_hat, y = pdf_hat, color = distr), size = 0.75)+
  geom_area(data = combined_pdf,
            aes(x = x_hat, y = pdf_hat, fill = distr), alpha = 0.5)+
  geom_errorbar(data = combined_obs,
                aes(ymin = low, ymax = up, x = emissions, color = distr),
                alpha = 0.5)+
  geom_point(data = combined_obs,
             aes(x = emissions, y = med, fill = distr),
             size = 3, alpha = 0.5, shape = 21, color = 'black')+
  geom_vline(aes(xintercept = UPL, color = distr, linetype = Significance), 
             data = combined_fit_all,
             linewidth = 1)+
  ylab("Density")+ xlab("Emissions")+
  pop_distr_theme()+ labs(linetype = 'Significance')+
  guides(color = "none", fill = "none")+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 3.1))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  coord_cartesian(clip = 'off')+ scale_color_manual(values = palette_line)+
  scale_fill_manual(values = palette_line)+
  facet_wrap(~distr, scales = 'fixed', nrow = 3)+
  theme(legend.position = 'top')+
  theme(strip.text = element_text(color = 'black', size = 12))
```
Based on these results, we would conclude that at 85\% significance the UPL is 
stable with regards to long-tail effects in the distribution. It is also worth 
pointing out that this area of the distribution also has data underlying it, 
increasing our confidence. Adjusting the significance to a stable level or 
changing `maxY` distribution limits are both ways to handle the unstable nature 
of long-tailed distributions, but both have draw backs. The best path forward is 
to avoid using UPL's from long-tailed distributions all together. If you have a 
similar or almost as good fit to another distribution without the tail that will 
be a better choice for the UPL. You can always test the sensitivity to find out 
if your fitted distribution has a problematic tail by varying `maxY` across a 
range of high values outside your data as we did in this demo. Fundamentally, the
99<sup>th</sup> percentile of a long tail is an unrealistic emission value. If no 
other distributions can be fit without a long tail, than the issue is likely a 
lack of representative data. More data will always go a long ways towards 
improving the distribution fit, and a lack of data can lead to one weird point 
forcing bad behavior. 
```{r fit2, echo = FALSE}
fit_metrics = group_by(combined_fit2, Significance)
fit_metrics = arrange(fit_metrics, SSE, .by_group = TRUE)
fit_metrics = mutate(fit_metrics, across(c(2, 3 ,5), ~signif(., digits = 2)))
knitr::kable(fit_metrics,
             caption = "Performance of Lognormal distributions with different MaxY limits and different significance levels.",
             col.names = c("Upper Limit", 'UPL', 'SSE', 'No. Obs. in 95% CI', 
                         'pdf integral', 'Significance'))
```
### Alternative distributions

Logically the next step is to try other distributions and see if we can find something better. We won't bother with the Normal distribution since the data are so heavily skewed. We also won't look at the Beta distribution since it has an upper limit of 1, and while the emissions observations are all less than that in our set of top performing sources, theoretically there could emissions higher than 1.
```{r bayesian_UPL, echo = TRUE, results = FALSE}
distributions = c('Gamma', 'Lognormal', 'Skewed')
results = Bayesian_UPL(data = dat_emiss, emissions = 'emissions', 
                       distr_list = distributions)
```
```{r figure5, echo = FALSE, out.width = "100%", fig.height = 7, fig.align = "center", fig.cap = "Fitted likelihood distributions for Lime HCl emissions."}
palette_line = c("#377EB8", "#4DAF4A", "#FF7F00")
ggplot()+
  geom_line(data = results$pred_pdf_dat,
            aes(y = ydens, x = x_hat), size = 0.75, color = 'black')+
  geom_area(data = results$pred_pdf_dat,
            aes(y = ydens, x = x_hat), fill = 'grey')+
  geom_point(data = results$obs_pdf_dat,
            aes(y = ydens, x = emissions),
            size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_line(data = results$pred_pdf_dat,
            aes(x = x_hat, y = pdf_hat, color = distr), size = 0.75)+
  geom_area(data = results$pred_pdf_dat,
            aes(x = x_hat, y = pdf_hat, fill = distr), alpha = 0.5)+
  geom_errorbar(data = results$obs_pdf_dat,
                aes(ymin = low, ymax = up, x = emissions, color = distr),
                alpha = 0.5)+
  geom_point(data = results$obs_pdf_dat,
             aes(x = emissions, y = med, fill = distr),
             size = 3, alpha = 0.5, shape = 21, color = 'black')+
    geom_vline(aes(xintercept = UPL, color = distr), data = results$fit_table,
             linewidth = 1)+
  ylab("Density")+ xlab("Lime HCl emissions")+
  pop_distr_theme()+ labs(color = 'Distribution', fill = "Distribution")+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  geom_rug(data = results$obs_pdf_dat,
           sides = 'b', aes(x = emissions),
           alpha = 0.5, outside = TRUE, color = 'black')+
  coord_cartesian(clip = 'off')+ scale_color_manual(values = palette_line)+
  scale_fill_manual(values = palette_line)+
  facet_wrap(~distr, scales = 'free_y', nrow = 3)+
  theme(legend.position = 'none')+
  theme(strip.text = element_text(color = 'black', size = 12))
```
```{r fit3, echo = TRUE}
fit_metrics = results$fit_table
```
``` {r table2, fig.align = "center"}
fit_metrics = arrange(fit_metrics, SSE)
knitr::kable(mutate(fit_metrics, across(c(2, 3), ~signif(., digits = 3))),
             caption = "Comparing goodness of fit results for other distributions",
             col.names = c('Distribution', 'UPL', 'SSE', 'No. Obs. in 95% CI', 
                           'pdf integral'))
```
The Gamma distribution looks the best and also has the lowest SSE. Furthermore, 
the Gamma distribution converges well. It still has a bit of a tail so we might want to repeat our experiments with increasingly larger values of `maxY`:
```{r maxy2, echo = TRUE, results = FALSE}
results_3X = Bayesian_UPL(distr_list = c("Gamma"), data = dat_emiss,
                          emissions = 'emissions',
                          xvals = seq(0, 3, length.out = 1000))
results_2X = Bayesian_UPL(distr_list = c("Gamma"), data = dat_emiss,
                          emissions = 'emissions',
                          maxY = 2 * max(dat_emiss$emissions),
                          xvals = seq(0, 3, length.out = 1000))
results_25X = Bayesian_UPL(distr_list = c("Gamma"), data = dat_emiss,
                           emissions = 'emissions',
                           maxY = 25 * max(dat_emiss$emissions),
                           xvals = seq(0, 3, length.out = 1000))
```
The outcome shows the tail is way more stable at the 99<sup>th</sup> percentile for Gamma than it was for Lognormal. Furthermore, the fit is just as good (SSE barely changes) and the UPL hardly changes. This is all great news, and leads us to finally recommend a UPL of `r round(results_25X$fit_table$UPL,2)`. 
```{r fit4, echo = FALSE}
results_3X$fit_table$distr = rep("3X Max", 
                                          nrow(results_3X$fit_table))
results_2X$fit_table$distr = rep("2X Max", 
                                          nrow(results_2X$fit_table))
results_25X$fit_table$distr = rep("25X Max", 
                                          nrow(results_25X$fit_table))
combined_fit = rbind(results_2X$fit_table, results_3X$fit_table, 
                     results_25X$fit_table)
combined_fit$distr = factor(combined_fit$distr, 
                            levels = c("2X Max", "3X Max", "25X Max"))

fit_metrics = mutate(combined_fit, across(c(2, 3, 5), ~signif(., digits = 2)))
knitr::kable(fit_metrics,
             caption = "Performance of Gamma distributions with different MaxY limits",
             col.names = c("Upper Limit", 'UPL', 'SSE', 'No. Obs. in 95% CI', 
                           'pdf integral'))
```
