---
title: "Hierarchical UPL"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hierarchical-UPL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  error = TRUE,
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
library(htmltools)
knitr::knit_hooks$set(imgcenter = function(before, options, envir){
  if (before) {
    HTML("<p align='center'>")
  } else {
    HTML("</p>")
  }
})
```

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(forcats)
library(UPLforOAR)
library(matrixStats)
library(np)
library(sfsmisc)
library(parallel)
library(rjags)
library(runjags)
library(sn)
library(patchwork)
```

## Introduction

One of the benefits of using the Bayesian method for UPL calculations, is that there are many ways the basic application in `Bayesian_UPL()` can be expanded upon for complex situations that need more nuance. Hierarchical Bayesian modeling offers solutions to many situations that cannot be handled when using `Normal_UPL()`, `Lognormal_UPL()`, or `Skewed_UPL()` due to their restrictive assumptions. For example, the assumption that emission values were perfectly measured is often not valid, and we would like a method that can take into account measurement uncertainty. The hierarchical structure we can use with Bayesian methods can allow us to say that the *true* emission value is a latent, unobserved, state, but we can use a prior distribution to inform it with a mean and variance based on the observed measurement and its associated uncertainty. The *true* emission values then inform the likelihood distribution for the population of top performing source emissions. 

Another assumption that is not always valid is that the emission test run values are all independent of one another. Typically, there are three runs within a test with some spacing of time between them to encourage independence, which might be minutes or hours depending on the type of measurement. Some sources might only have one test a year, or many within a year, or a much longer or shorter record of testing. 
