---
title: "Methods Comparison"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Methods-Comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  error = TRUE,
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
library(htmltools)
knitr::knit_hooks$set(imgcenter = function(before, options, envir){
  if (before) {
    HTML("<p align='center'>")
  } else {
    HTML("</p>")
  }
})
```

```{r setup, message=FALSE}
library(dplyr)
library(ggplot2)
library(forcats)
library(UPLforOAR)
library(matrixStats)
library(np)
library(sfsmisc)
library(parallel)
library(rjags)
library(runjags)
library(sn)
library(patchwork)
library(grid)
library(readr)
```

## Introduction

There are two families of UPL methods included in the `UPLforOAR` package: those 
replicating the functionality of the Excel workbook in the frequentist framework, 
(`distribution_type()`, `Normal_UPL()`, `Lognormal_UPL()`, and `Skewed_UPL()`)
and those using fitted likelihood distributions in the Bayesian framework 
(`Bayesian_UPL()`). The UPL is the same MACT floor limit in principle, with 
these two families of methods simply being different ways of arriving at the 
same answer. However, there are differences worth explaining, pros and cons, and 
circumstances where the frequentist and Bayesian methods will get different UPL 
estimates.

### Frequentist

The UPL for Normal data is well-suited to the frequentist framework because it 
can be calculated analytically using the mean, standard deviation, and 
*t*-statistic of the emissions data. However, there are no equivalent analytical 
solutions for Skewed or Lognormal UPL calculations. Instead, `Lognormal_UPL()` 
uses a Gram-Charlier Type A Series expansion to approximate the density 
distribution and determine the arithmetic mean of future samples, and 
`Skewed_UPL()` uses the skewness and kurtosis moments to iteratively adjust the 
*t*-statistic until the desired significance is achieved. All three UPL methods 
assume the distribution explicitly and without uncertainty, and assume the 
emissions are observed perfectly and independently. This means that the 
parameters defining the probability distributions from which the UPL is derived 
are fixed point estimates. For example, any uncertainty around the mean of 
emissions in normally distributed data does *not* carry through to the UPL estimate. 
Furthermore, if there are circumstances where we don't want to assume all test 
runs are independent, we want to incorporate additional sources of variance, or 
perhaps use individual test run uncertainty, then it will be very difficult if 
not impossible to accommodate in these methods. However, these formulas are very 
quick to use, have no stochastic components, and have been in use a long time. 

### Bayesian

In contrast, the UPL is estimated the same way regardless of the distribution in 
the Bayesian approach. In this method we use Bayesian MCMC to fit a likelihood 
distribution to the emissions data. The results of this include posterior 
distributions, which contain a full probability around your result rather than a 
fixed point estimate. That means the UPL for a normal distribution is based on 
full posterior distribution of the mean and standard deviation, rather than 
fixed point estimates of the mean and standard deviation. This gives us much 
better quantification of uncertainty in how well the distribution represents the 
data. While we could still use the ratios of skewness and kurtosis implemented 
in `distribution_type()` to pick the distribution to use, we can do better in 
the Bayesian framework. Those ratios only tell us if the data are approximately 
Normal or Lognormal, and are assigned as Skewed for any other outcome. But can 
use many more types of distributions with Bayes that we will want to pick between. 
With the `Bayesian_UPL()` function, we have several ways to quantitatively and 
qualitatively asses the best fit distribution. Rather than assuming the 
distribution is correct implicitly, we can evaluate how well the parameters 
converged, which is another method by which we can reject bad choices for 
distributions.

The biggest advantage in using a Bayesian approach are the expanded options for 
distributions, such as Beta and Gamma which are already implemented in 
`Bayesian_UPL()`. Most emissions data are bounded by zero, and in some 
circumstances have upper bounds as well. Bounded data can be accommodated both 
by picking an appropriate distribution and by explicitly truncating the 
likelihood distribution. More nuanced situations, such as non-independent test 
run data, additional sources of variance, or individual test run uncertainty, 
can easily be incorporated into the Bayesian approach without changing the 
underlying methods of calculating the UPL. The main draw back is in the 
difficulty of implementing the method. The Markov Chain Monte Carlo (MCMC) Gibbs 
sampler takes longer than a simple analytical solution. Since there are fewer 
assumptions, there are more steps to ensure the likelihood model worked well and 
is appropriate to the data. There is a stochastic component due to the way the 
MCMC sampler searches for the parameter solutions. While this shouldn't have a 
meaningful effect on the outcome and can be controlled to create reproducible 
results, it is a factor influencing different outcomes that might be undesirable. 
Lastly, the Bayesian method uses prior information. While in many situations 
this can be a strength in the method, here we generally want completely 
uninformative priors for UPL calculations. We very much want to avoid a
situation where we think the prior is uninformative, but in fact it is not.

For more details on `Bayesian_UPL()` see the example and details at [Bayesian UPL](https://luddaludwig.github.io/UPLforOAR/articles/Bayesian-UPL.html).

## Example Comparison

We are going to calculate the UPL from emissions data using multiple methods and 
compare the outcomes. This example uses Hg emissions data from the recent [EPA rule-making](https://www.regulations.gov/document/EPA-HQ-OAR-2009-0234-20132) 
NESHAP for Coal- and Oil-fired Electric Utility Steam Generating Units. First we 
load and organize the data into `emissions` and `sources`, then select the top 
performers to use for the UPL calculations. We are only going to be considering 
the existing source data set since it does not include multiple runs per source.

```{r load, echo=TRUE}
dat_emiss = read_csv("./../man/data_example/MATS_Hg.csv", col_names = TRUE)
dat_emiss$sources = paste0(dat_emiss$`Plant Name`, "_", dat_emiss$`Unit Number`,
                         "_", dat_emiss$boiler_id)
dat_emiss$emissions = dat_emiss$Mercury_min_lb_MMBtu
dat_emiss = subset(dat_emiss, select = c(sources, emissions))

dat_exist = MACT_existing(CAA_section=112, dat_emiss)
dat_exist_avg = dat_exist %>% group_by(sources) %>% 
  summarize(avg = mean(emissions), counts = n())
dat_exist_avg = arrange(dat_exist_avg, avg)
distribution_result_exist = distribution_type(dat_exist)
```

The skewness and kurtosis ratio tests tell us the distribution is 
`r distribution_result_exist`. However, looking at plot of the data it is clear 
that the Normal (in orange) doesn't actually fit well (Fig. 1). This is because 
those ratio tests do not account for zero as a lower boundary. 

```{r plot1, echo=FALSE, fig.dim = c(5,4), fig.align = "center", fig.cap = "Observation density of Hg for the overall population. The obseration data are indicated in black as points and a rug along the axis, with the observation density distribution as a black line. The fitted normal distribution that is the basis of the UPL estimate is colored orange, with a zero lower boundary. The average of the Hg emissions is the vertical black line."}
x_hat = seq(0, 3 * max(dat_exist$emissions), length.out = 1024)
obs_dens_results = obs_density(dat_exist, xvals = x_hat)
Obs_onPoint = obs_dens_results$Obs_onPoint
obs_den_df = obs_dens_results$obs_den_df
pdf_n = dnorm(x_hat, mean = mean(dat_exist$emissions, na.rm = T),
              sd = sd(dat_exist$emissions, na.rm = T))
pred_dat = tibble(x_hat, pdf_n)
ggplot()+
  geom_line(data = obs_den_df, size=0.75,
            aes(y = ydens, x = x_hat, color = 'a'))+
  geom_area(data = obs_den_df, alpha = 0.25,
            aes(y = ydens, x = x_hat, fill = 'a'))+
  geom_point(aes(y = ydens, x = emissions), data = Obs_onPoint,
             size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_line(aes(y = pdf_n, x = x_hat, color = 'b'),
               data = pred_dat, size = 0.75, linetype = 2)+
  geom_area(aes(y = pdf_n, x = x_hat, fill = 'b'), alpha = 0.25,
            data = pred_dat)+
  ylab("Density")+xlab("Hg emissions (lb/MMBtu)")+
  ggtitle("Overall observed population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept = (mean(dat_exist$emissions)),
                 color = 'a'), size = 1, linetype = 1)+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  coord_cartesian(clip = 'off')+
  labs(color = 'Distribution:', fill = 'Distribution:')+
  geom_rug(sides = 'b', aes(x = emissions), data = dat_exist,
           alpha = 0.5, outside = TRUE, color = 'black')+
  scale_color_manual(values = c('black', '#FF7F00'),
                     labels = c('Observations', 'Normal'))+
  scale_fill_manual(values = c('black', '#FF7F00'),
                    labels = c('Observations', 'Normal'))
```
We can recreate this plot without the zero boundary for the density. Notice that 
the Normal distribution a lot looks more reasonable when we allow the left tails 
to extend a bit below zero, and things become symmetric (Fig. 2). This is 
unrealistic, since we know we cannot have negative emissions. However, we cannot 
account for boundaries in the analytical normal UPL calculation. In the Bayesian 
framework, we can explore both bounded and unbounded normal distributions.

```{r plot2, echo=FALSE, fig.dim = c(5,4), fig.align = "center", fig.cap = "Observation density of Hg for the overall population. The obseration data are indicated in black as points and a rug along the axis, with the observation density distribution as a black line. The fitted normal distribution that is the basis of the UPL estimate is colored orange, without boundaries. The average of the Hg emissions is the vertical black line."}
x_hat = seq(-mean(dat_exist$emissions), 3 * max(dat_exist$emissions), length.out = 1024)
pdf_n = dnorm(x_hat, mean = mean(dat_exist$emissions, na.rm = T),
              sd = sd(dat_exist$emissions, na.rm = T))
pred_dat = tibble(x_hat, pdf_n)
ggplot()+
  geom_density(aes(emissions, color = 'a', fill = 'a'), 
               data = dat_exist, alpha = 0.25, size = 0.75)+
  geom_line(aes(y = pdf_n, x = x_hat, color = 'b'),
               data = pred_dat, size = 0.75, linetype = 2)+
  geom_area(aes(y = pdf_n, x = x_hat, fill = 'b'), alpha = 0.25,
            data = pred_dat)+
  ylab("Density")+xlab("Hg emissions (lb/MMBtu)")+
  ggtitle("Overall observed population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept = (mean(dat_exist$emissions)),
                 color = 'a'), size = 1, linetype = 1)+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), 
                     limits = c(-5e-8, 8e-8))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  coord_cartesian(clip = 'off')+
  labs(color = 'Distribution:', fill = 'Distribution:')+
  geom_rug(sides = 'b', aes(x = emissions), data = dat_exist,
           alpha = 0.5, outside = TRUE, color = 'black')+
  scale_color_manual(values = c('black', '#FF7F00'),
                     labels = c('Observations', 'Normal'))+
  scale_fill_manual(values = c('black', '#FF7F00'),
                    labels = c('Observations', 'Normal'))
```
Ignoring the best distribution for the moment, we are going to calculate the UPL 
using the frequentist methods, the Bayesian methods unbounded, and the Bayesian 
methods bounded, for the Normal, Lognormal, and Skewed distributions. We are 
using 1 for the number of future runs, since the data here are already averages 
of 3 runs.

```{r UPL, echo = TRUE, results = FALSE}
UPL_freq_N = Normal_UPL(data = dat_exist,
                     future_runs = 1,
                     significance = 0.99)
UPL_freq_LN = Lognormal_UPL(data = dat_exist,
                        future_runs = 1,
                        significance = 0.99)
UPL_freq_SN = Skewed_UPL(data = dat_exist,
                        future_runs = 1,
                        significance = 0.99)
Bayes_0bounded = Bayesian_UPL(distr_list = c("Normal", "Lognormal", "Skewed"),
                             data = dat_exist, minY = 0, 
                             future_runs = 1, significance = 0.99)
Bayes_unbounded = Bayesian_UPL(distr_list = c("Normal", "Lognormal", "Skewed"),
                             data = dat_exist, 
                             minY = -mean(dat_exist$emissions), 
                             future_runs = 1, significance = 0.99)
```
While the zero bounded and unbounded Bayesian results look very similar, there 
are slight differences caused by the zero hard boundary. As a result the bounded 
distributions look slightly better, and are more realistic.
```{r plot3, echo = FALSE, fig.dim = c(4,8), fig.align = "center", fig.cap = "Fitted likelihood distributions for zero bounded (right) and unbounded (left) Bayes distributions. The fitted distributions are colored blue for lognormal, orange for normal, and green for skewed."}
Bayes_0bounded$obs_pdf_dat$type = rep("Zero bounded", 
                                          nrow(Bayes_0bounded$obs_pdf_dat))
Bayes_0bounded$pred_pdf_dat$type = rep("Zero bounded", 
                                          nrow(Bayes_0bounded$pred_pdf_dat))
Bayes_0bounded$fit_table$type = rep("Zero bounded", 
                                          nrow(Bayes_unbounded$fit_table))
Bayes_unbounded$obs_pdf_dat$type = rep("Unbounded", 
                                          nrow(Bayes_unbounded$obs_pdf_dat))
Bayes_unbounded$pred_pdf_dat$type = rep("Unbounded", 
                                          nrow(Bayes_unbounded$pred_pdf_dat))
Bayes_unbounded$fit_table$type = rep("Unbounded", 
                                          nrow(Bayes_unbounded$fit_table))
combined_pdf = rbind(Bayes_0bounded$pred_pdf_dat, 
                     Bayes_unbounded$pred_pdf_dat)
combined_obs = rbind(Bayes_0bounded$obs_pdf_dat, 
                     Bayes_unbounded$obs_pdf_dat)
combined_fit = rbind(Bayes_0bounded$fit_table, 
                     Bayes_unbounded$fit_table)

palette_line = c("#377EB8", "#FF7F00", "#4DAF4A")
ggplot()+
  geom_line(data = combined_pdf,
            aes(y = ydens, x = x_hat), size = 0.75, color = 'black')+
  geom_area(data = combined_pdf,
            aes(y = ydens, x = x_hat), fill = 'grey')+
  geom_point(data = combined_obs,
            aes(y = ydens, x = emissions),
            size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_line(data = combined_pdf,
            aes(x = x_hat, y = pdf_hat, color = distr), size = 0.75)+
  geom_area(data = combined_pdf,
            aes(x = x_hat, y = pdf_hat, fill = distr), alpha = 0.5)+
  geom_errorbar(data = combined_obs, width = 1e-9,
                aes(ymin = low, ymax = up, x = emissions, color = distr),
                alpha = 0.5)+
  geom_point(data = combined_obs,
             aes(x = emissions, y = med, fill = distr),
             size = 3, alpha = 0.5, shape = 21, color = 'black')+
  ylab("Density")+ xlab("Hg emissions")+
  pop_distr_theme()+ labs(color = 'Distribution', fill = "Distribution")+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  geom_rug(data = combined_obs,
           sides = 'b', aes(x = emissions),
           alpha = 0.5, outside = TRUE, color = 'black')+
  coord_cartesian(clip = 'off')+ scale_color_manual(values = palette_line)+
  scale_fill_manual(values = palette_line)+
  facet_wrap(~distr+type, scales = 'fixed', nrow = 3)+
  theme(legend.position = 'none')+theme(panel.spacing.x = unit(5, "mm"))+
  theme(axis.text.y = element_text(size = 10), 
        axis.text.x = element_text(size = 10))+
  theme(strip.text = element_text(color = 'black', size = 12))
```
All six of the Bayesian distributions converged. When it comes to the 
goodness-of-fit metrics, the best distribution was Skewed for both Bayes bounded 
and unbounded, with the zero bounded being the best overall. The Normal 
distribution did have the most points within 95\% CI, but had the worst fit when 
zero-bounded and did not integrate to 1. 

``` {r table1, fig.align = "center", echo = FALSE}
fit_metrics = subset(combined_fit, select = -c(UPL)) %>% group_by(type)
fit_metrics = arrange(fit_metrics, SSE, .by_group = TRUE)
knitr::kable(mutate(fit_metrics, across(c(2, 3), ~signif(., digits = 3))),
             caption = "Goodness of fit results for Hg emissions",
             col.names = c('Distribution', 'SSE', 'No. Obs. in 95% CI', 
                           'integral', 'Type'))
```

But we aren't interested in finding the best distribution in this exercise, just comparing UPL methods. The Lognormal doesn't change between bounded and unbounded Bayesian methods. That's because it already is strictly positive. The frequentist Lognormal result is similar but not the same, but it uses an approximation that is not replicated in the Bayesian method that isn't unexpected. All Lognormal results are similarly larger than all others, which is typical for the long-tailed distribution. The Skewed changes a little with the zero boundary, since it isn't strictly positive but is asymmetric. It changes a similar amount with the frequentist method, which again is an approximation, and not actually the 99$^{th}$ percentile of a skew-normal distribution. The unbounded Bayesian and frequentist UPL's for the Normal distribution are very similar to one another. Since the frequentist method has an exact analytical solution, we should be aiming to arrive at the same values with the Bayesian method, excepting the stochastic component and added accountability to uncertainty. The zero-bounded Bayesian Normal UPL is quite different, since we are essentially fitting a partial Normal.

```{r table2, echo = FALSE}
fit_metrics = subset(combined_fit, select = c(distr, type, UPL)) 
fit_metrics = rbind(fit_metrics, c("Normal", "Frequentist", UPL_freq_N))
fit_metrics = rbind(fit_metrics, c("Lognormal", "Frequentist", UPL_freq_LN))
fit_metrics = rbind(fit_metrics, c("Skewed", "Frequentist", UPL_freq_SN))
fit_metrics$UPL = as.numeric(fit_metrics$UPL)
fit_metrics = fit_metrics %>% group_by(distr)
fit_metrics = arrange(fit_metrics, type, .by_group = TRUE)
knitr::kable(fit_metrics,
             caption = "UPL results for fitting 3 distributions with 
             frequentist, unbounded Bayesian, and zero bounded Bayesian.",
             col.names = c("Distribution", "Method", "UPL"), digits = 10)
```

The next logical question is how much does the UPL change if we use different random seeds when starting the MCMC iterations? We can explore this be re-running distribution fits with `random = TRUE`. These results will *not* be exactly reproducible. We will fit three sets of unbounded Normal distributions with identical inputs, but allowing the random seeds to change. 

```{r random, echo = TRUE, results = FALSE}
Bayes_random1 = Bayesian_UPL(distr_list = c("Normal"),
                             data = dat_exist, random = TRUE, 
                             minY = -mean(dat_exist$emissions), 
                             future_runs = 1, significance = 0.99)
Bayes_random2 = Bayesian_UPL(distr_list = c("Normal"),
                             data = dat_exist, random = TRUE, 
                             minY = -mean(dat_exist$emissions), 
                             future_runs = 1, significance = 0.99)
Bayes_random3 = Bayesian_UPL(distr_list = c("Normal"),
                             data = dat_exist, random = TRUE, 
                             minY = -mean(dat_exist$emissions), 
                             future_runs = 1, significance = 0.99)
```
```{r random_UPL, echo = FALSE}
Bayes_fixed = subset(Bayes_unbounded$fit_table, 
                     Bayes_unbounded$fit_table$distr == "Normal")
Bayes_fixed$type = rep("Fixed", nrow(Bayes_fixed))
Bayes_random1$fit_table$type = rep("Random 1", nrow(Bayes_random1$fit_table))
Bayes_random2$fit_table$type = rep("Random 2", nrow(Bayes_random2$fit_table))
Bayes_random3$fit_table$type = rep("Random 3", nrow(Bayes_random3$fit_table))
combined_fit = rbind(Bayes_fixed, 
                     Bayes_random1$fit_table,
                     Bayes_random2$fit_table,
                     Bayes_random3$fit_table)
fit_metrics = subset(combined_fit, select = c(type, UPL)) 
```

Looking at the UPL results, we do see some variance in the UPL, but it is much 
smaller than the difference between approximation methods and distributions. The 
standard deviation in these 4 repeated UPL results is `r signif(sd(fit_metrics$UPL),3)` 
which is similar to the difference between the Bayesian and frequentist Normal UPL's.
```{r table3, echo = FALSE}
knitr::kable(fit_metrics,
             caption = "UPL results for fitting unbounded Normal distributions with 
             fixed and then 3 different random seeds.",
             col.names = c("Type", "UPL"), digits = 11)
```
