
#### `r type` Source `r HAP` `r subcat_text`

The selected distribution for the `r type` source data set for `r HAP` `r subcat_text` was the `r distr` distribution, with the likelihood truncated such that it was zero outside the bounds of (`r minY`, `r signif(maxY, 3)`) emissions. The parameter prior distributions, MCMC initial chain values, and Gelman-Rubin diagnostics for convergence are listed in the table below, followed by the random number generator (RNG) state that was used.

```{r, echo = FALSE, appendix = FALSE}
# This R chunk uses the initial value settings copied from setup_likelihood(). 
# If setup_likelihood() has these calculations changed, it will need to be updated here.
# Eventually, it would be ideal to have a separate function to set initial values 
# based on data and distribution, then setup_likelihood() would call it and it 
# would be called here, so updates only happen in a single place.
if (!manual_prior){
  if (distr == 'Gamma'){
    sdY =  sd(df$emissions)
    shape = mean(df$emissions)^2 / sdY^2
    rate = mean(df$emissions) / sdY^2
    parameter = c('Rate', 'Shape')
    prior = c(paste0('~dunif (0, ', signif(maxY / (sdY^2), 3), ')'), 
              paste0('~dunif (0, ', signif((maxY^2) / (sdY^2), 3), ')'))
    chain1 = c(signif(rate, 3), signif(shape, 3))
    chain2 = c(signif(1.5 * rate, 3), signif(0.5 * shape, 3))
    chain3 = c(signif(0.5 * rate, 3), signif(1.5 * shape, 3))
    diagnostics = c(signif(results$conv_output$gelman_diag[1], 3), 
                    signif(results$conv_output$gelman_diag[2], 3))
    convYN = c(results$conv_output$convYN[1], 
              results$conv_output$convYN[2])
  } else if (distr == 'Normal'){
    sdY =  sd(df$emissions)
    meanY = mean(df$emissions)
    parameter = c('Mean', 'Standard deviation')
    prior = c(paste0('~dunif (0, ', signif(maxY, 3), ')'),
              paste0('~dunif (', signif(0.001 * sdY, 3), ' ,',
                    signif(1000 * sdY, 3), ')'))
    chain1 = c(signif(meanY, 3), signif(sdY, 3))
    chain2 = c(signif(1.5 * meanY, 3), signif(0.5 * sdY, 3))
    chain3 = c(signif(0.5 * meanY, 3), signif(1.5 * sdY, 3))
    diagnostics = c(signif(results$conv_output$gelman_diag[1], 3),
                  signif(results$conv_output$gelman_diag[2], 3))
    convYN = c(results$conv_output$convYN[1],
            results$conv_output$convYN[2])
  } else if (distr == 'Skewed'){
    sdY =  sd(df$emissions)
    meanY = mean(df$emissions)
    skew1 = min(0.99, abs((1 / length(df$emissions)) *
                          sum(((df$emissions - meanY) / sdY)^3)))
    delta = sqrt((pi / 2) * ((abs(skew1)^(2 / 3)) / ((abs(skew1)^(2 / 3)) +
                                                   ((4 - pi) / 2)^(2 / 3))))
    delta = delta * abs(skew1) / skew1
    alpha = delta / sqrt(1 - delta^2)
    omega = sdY / sqrt(1 - 2 * delta^2 / pi)
    xi = meanY - omega * delta * sqrt(2 / pi)
    parameter = c('Omega', 'Xi', 'Alpha')
    prior = c(paste0('~dunif (0, ', signif(maxY, 3), ')'),
              paste0('~dnorm (0, ', signif(1 / (100 * maxY), 3), ')'),
              '~dunif (-100, 100)')
    chain1 = c(signif(xi, 3), signif(omega, 3), signif(alpha, 3))
    chain2 = c(signif(1.5 * xi, 3), signif(0.5 * omega, 3), signif(0.5 * alpha, 3))
    chain3 = c(signif(0.5 * xi, 3), signif(1.5 * omega, 3), signif(1.5 * alpha, 3))
    diagnostics = c(signif(results2$conv_output$gelman_diag[1],3),
                    signif(results2$conv_output$gelman_diag[2],3),
                    signif(results2$conv_output$gelman_diag[3],3))
    convYN = c(results2$conv_output$convYN[1],
               results2$conv_output$convYN[2],
               results2$conv_output$convYN[3])
  } else if(distr == "Lognormal"){
    sdY =  sd(df$ln_emiss)
    meanY = mean(df$ln_emiss)
    parameter = c('Mean', 'Standard deviation')
    prior = c(paste0('~dnorm (', meanY, ', ', signif(0.001 / (sdY^2), 3), ')'),
              paste0('~dunif (', signif(0.001 * sdY, 3), ' ,',
                    signif(1000 * sdY, 3), ')'))
    chain1 = c(signif(meanY, 3), signif(sdY, 3))
    chain2 = c(signif(1.5 * meanY, 3), signif(0.5 * sdY, 3))
    chain3 = c(signif(0.5 * meanY, 3), signif(1.5 * sdY, 3))
    diagnostics = c(signif(results$conv_output$gelman_diag[1], 3),
                    signif(results$conv_output$gelman_diag[2], 3))
    convYN = c(results$conv_output$convYN[1],
              results$conv_output$convYN[2])
  } else if (distr == 'Beta'){
    sdY =  sd(df$emissions)
    meanY = mean(df$emissions)
    alpha = meanY^2 / sdY^2 - meanY^3 / sdY^2 - meanY
    beta = meanY - 1 + meanY / sdY^2 + meanY^3 / sdY^2 - 2 * meanY^2 / sdY^2
    parameter = c('Lambda', 'Tau')
    prior = c(paste0('~dunif (-1, ', signif((0.75^2 / sdY^2 - 0.75^3 / sdY^2 - 0.75)
                                        * 1.25, 3), ')'),
              paste0('~dunif (-1, ', signif((0.25 - 1 + 0.25 / sdY^2 +
                                           0.25^3 / sdY^2 - 2 * 0.25^2 / sdY^2)
                                        * 1.25, 3), ')'))
    chain1 = c(signif(alpha, 3), signif(beta, 3))
    chain2 = c(signif(1.5 * alpha, 3), signif(0.5 * beta, 3))
    chain3 = c(signif(0.5 * alpha, 3), signif(1.5 * beta, 3))
    diagnostics = c(signif(results$conv_output$gelman_diag[1], 3),
                    signif(results$conv_output$gelman_diag[2], 3))
    convYN = c(results$conv_output$convYN[1],
               results$conv_output$convYN[2])
  }
} else if (manual_prior){
  if (distr == "Normal"){
    parameter = c('Mean', 'Standard deviation')
    prior = c(paste0('~dunif (', prior_list[3], ', ', prior_list[4], ')'),
              paste0('~dunif (', prior_list[1], ', ', prior_list[2], ')'))
    chain1 = c(signif(mean(c(prior_list[3], prior_list[4])), 3), 
               signif(mean(c(prior_list[1], prior_list[2])), 3))
    chain2 = c(signif(0.9 * prior_list[4], 3), signif(0.9 * prior_list[2], 3))
    chain3 = c(signif(1.1 * prior_list[3], 3), signif(1.1 * prior_list[1], 3))
    diagnostics = c(signif(results$conv_output$gelman_diag[1], 3),
                  signif(results$conv_output$gelman_diag[2], 3))
    convYN = c(results$conv_output$convYN[1],
               results$conv_output$convYN[2])  
  } else if (distr == 'Lognormal'){
          if (prior_list[4] < 0){
            initial1 = 1.1 * prior_list[4]
          } else if (prior_list[4] > 0){
            initial1 = 0.9 * prior_list[4]
          }
          if (prior_list[3] < 0){
            initial2 = 0.9 * prior_list[3]
          } else if (prior_list[3] > 0){
            initial2 = 1.1 * prior_list[3]
          }
    parameter = c('Mean', 'Standard deviation')
    prior = c(paste0('~dunif (', prior_list[3], ', ', prior_list[4], ')'),
              paste0('~dunif (', prior_list[1], ', ', prior_list[2], ')'))
    chain1 = c(signif(mean(c(prior_list[3], prior_list[4])), 3), 
               signif(mean(c(prior_list[1], prior_list[2])), 3))
    chain2 = c(signif(initial1, 3), signif(0.9 * prior_list[2], 3))
    chain3 = c(signif(initial2, 3), signif(1.1 * prior_list[1], 3))
    diagnostics = c(signif(results$conv_output$gelman_diag[1], 3),
                    signif(results$conv_output$gelman_diag[2], 3))
    convYN = c(results$conv_output$convYN[1],
               results$conv_output$convYN[2])      
  } else if (distr == 'Skewed'){
      if (prior_list[4] < 0){
        initial1 = 1.1 * prior_list[4]
      } else if (prior_list[4] > 0){
        initial1 = 0.9 * prior_list[4]
      }
      if (prior_list[3] < 0){
        initial2 = 0.9 * prior_list[3]
      } else if (prior_list[3] > 0){
        initial2 = 1.1 * prior_list[3]
      }
      if (prior_list[6] < 0){
        initial3 = 1.1 * prior_list[6]
      } else if (prior_list[6] > 0){
        initial3 = 0.9 * prior_list[6]
      }
      if (prior_list[5] < 0){
        initial4 = 0.9 * prior_list[5]
      } else if (prior_list[5] > 0){
        initial4 = 1.1 * prior_list[5]
      }
    parameter = c('Omega', 'Xi', 'Alpha')
    prior = c(paste0('~dunif (', prior_list[1], ', ', prior_list[2], ')'),
              paste0('~dunif (', prior_list[3], ', ', prior_list[4], ')'),
              paste0('~dunif (', prior_list[5], ', ', prior_list[6], ')'))
    chain1 = c(signif(mean(c(prior_list[1], prior_list[2])), 3),
               signif(mean(c(prior_list[3], prior_list[4])), 3),
               signif(mean(c(prior_list[5], prior_list[6])), 3))
    chain2 = c(signif(1.1 * prior_list[1], 3), 
               signif(initial1, 3), 
               signif(initial4, 3))
    chain3 = c(signif(0.9 * prior_list[2], 3), 
               signif(initial2, 3), 
               signif(initial3, 3))
    diagnostics = c(signif(results2$conv_output$gelman_diag[1],3),
                    signif(results2$conv_output$gelman_diag[2],3),
                    signif(results2$conv_output$gelman_diag[3],3))
    convYN = c(results2$conv_output$convYN[1],
               results2$conv_output$convYN[2],
               results2$conv_output$convYN[3])
  } else if (distr == 'Gamma'){
    parameter = c('Rate', 'Shape')
    prior = c(paste0('~dunif (', prior_list[1], ', ', prior_list[2], ')'),
              paste0('~dunif (', prior_list[3], ', ', prior_list[4], ')'))
    chain1 = c(signif(mean(c(prior_list[1], prior_list[2])), 3),
               signif(mean(c(prior_list[3], prior_list[4])), 3))
    chain2 = c(signif(1.1 * prior_list[1], 3), signif(1.1 * prior_list[3], 3))
    chain3 = c(signif(0.9 * prior_list[2], 3), signif(0.9 * prior_list[4], 3))
    diagnostics = c(signif(results$conv_output$gelman_diag[1], 3), 
                    signif(results$conv_output$gelman_diag[2], 3))
    convYN = c(results$conv_output$convYN[1], 
               results$conv_output$convYN[2])
  } else if (distr == 'Beta'){
      if (prior_list[2] < 0){
        initial1 = 1.1 * prior_list[2]
      } else if (prior_list[2] > 0){
        initial1 = 0.9 * prior_list[2]
        if (initial1 < (-1)){
          initial1 = -0.99999
        }
      }
      if (prior_list[1] < 0){
        initial2 = 0.9 * prior_list[1]
      } else if (prior_list[1] > 0){
        initial2 = 1.1 * prior_list[1]
        if (initial1 < (-1)){
          initial1 = -0.99999
        }
      }
      if (prior_list[4] < 0){
        initial3 = 1.1 * prior_list[4]
      } else if (prior_list[4] > 0){
        initial3 = 0.9 * prior_list[4]
        if (initial1 < (-1)){
          initial1 = -0.99999
        }
      }
      if (prior_list[3] < 0){
        initial4 = 0.9 * prior_list[3]
      } else if (prior_list[3] > 0){
        initial4 = 1.1 * prior_list[3]
        if (initial1 < (-1)){
          initial1 = -0.99999
        }
      }
    parameter = c('Lambda', 'Tau')
    prior = c(paste0('~dunif (', prior_list[1], ', ', prior_list[2], ')'),
              paste0('~dunif (', prior_list[3], ', ', prior_list[4], ')'))
    chain1 = c(signif(mean(c(prior_list[1], prior_list[2])), 3), 
               signif(mean(c(prior_list[3], prior_list[4])), 3))
    chain2 = c(signif(initial1, 3), signif(initial3, 3))
    chain3 = c(signif(initial2, 3), signif(initial4, 3))
    diagnostics = c(signif(results$conv_output$gelman_diag[1], 3),
                    signif(results$conv_output$gelman_diag[2], 3))
    convYN = c(results$conv_output$convYN[1],
               results$conv_output$convYN[2])
  }
}
param_tab = tibble(parameter, prior, chain1, chain2, chain3, 
                   diagnostics, convYN)
param_tab %>% knitr::kable(
  caption = paste0("Parameter information for the ", distr, " distribution 
  UPL calculation for ", type, " source", HAP, " ", subcat_text),
             col.names = c('Parameter',
                         'Prior',
                         'Chain 1 initial value',
                         'Chain 2 initial value',
                         'Chain 3 initial value',
                         'Gelman-Rubin diagnostic',
                         'Converged?'))
```

```{r, appendix = FALSE, echo = FALSE}
state_out = results$state
chains = c('Chain 1', 'Chain 2', 'Chain 3')
rng_name = c(state_out[[1]]$.RNG.name, 
               state_out[[2]]$.RNG.name, 
               state_out[[3]]$.RNG.name)
if (is.null(RNG.state)){
  rng_state = c('seed = 5', 'seed = 12', 'seed = 151')
} else {
  rng_state = c(paste(state_out[[1]]$.RNG.state, collapse=', '), 
                paste(state_out[[2]]$.RNG.state, collapse=', '), 
                paste(state_out[[3]]$.RNG.state, collapse=', '))
}
RNG1 = tibble(chains, rng_name, rng_state)
RNG1 %>% knitr::kable(
  caption="Random number generator (RNG) names and states used in MCMC chains",
             col.names=c('MCMC Chain',
                         'Name',
                         'State'))
```
