---
title: "UPL calculations"
author: "OAR-OCAP-NRD-EDAB"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
runtime: shiny
geometry: left = 2cm, right = 2cm, top = 2cm, bottom = 2cm
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(kableExtra)
library(UPLforOAR)
```

```{r inputs, echo=FALSE}
inputPanel(
        fileInput('data', 'Upload csv file with "emissions" 
                   and "sources" as columns',
                   accept = '.csv'),
        selectInput("CAA_section", label = "CAA Section:", selected = 112,
                    choices = c(112, 129)),
        checkboxInput('runMACT', 'Subset data to top performers', FALSE),
        numericInput("future_runs",label = "Number of future test runs",
                     value = 3, min = 1),
        numericInput("significance",label = "Level of significance",
                     value = 0.99, min = 0, max = 1)
)
downloadButton('Sourcedownload1', 'Download top performers data')
downloadButton('Sourcedownload2', 'Download best performer data')
downloadButton('UPLdownload', 'Download UPL results')
downloadButton("pdf_report", 'Download PDF report')

```

## Emission dataset

Check Table 1 below to ensure you have loaded the correct emission data set and 
that the values are generally what you expect.
```{r load_data}
dat_emiss = reactive({
    req(input$data)
    ext = tools::file_ext(input$data$name)
    switch(ext,
      csv = read.csv(input$data$datapath),
      validate("Invalid file; Please upload a .csv file")
    )
})
zero_warn = reactive({
  dat = dat_emiss()
  if (any(dat$emissions == 0)){
    out = TRUE
  } else {
    out = FALSE
  }
  out
})
output$dat_display = reactive({
  validate(
    need(("emissions" %in% names(dat_emiss())),
         "data must have numeric column named 'emissions'"),
    need(("sources" %in% names(dat_emiss())),
         "data must have a column named 'sources'")
  )
  summary(dat_emiss()) %>% knitr::kable('html' ,caption = 'Summary of dataset') %>%
                kable_styling("striped", full_width = T)
})
  
tableOutput('dat_display')

CAA_text=reactive(
  if (input$CAA_section == 112){
  'CAA Section 112(d)(3)'
} else if (input$CAA_section == 129){
  'CAA Section 129(a)(2)'
}
)
mact_EVAL = reactive({
  if (is.null(input$runMACT)){
    out = FALSE
  } else if (!input$runMACT){
    out = FALSE
  } else if (input$runMACT){
    out = TRUE
  }
  out
})
```

## Source calculation

```{r text0c}
output$text0c = renderText("Warning: you have uploaded emissions data containing zero values. These have been removed from the data set.")
output$uitext0c = renderUI({
  if (zero_warn()){
    htmlOutput(outputId = 'text0c')
  }
})
uiOutput(outputId = 'uitext0c')
```

```{r text0a}
output$text0a = renderText(paste0("For existing source standards calculations, ",
                               CAA_text(),
                               " is used to determine the top sources, either the best 5 performers or the best 12% performers. The number of future tests for calculating the UPL is ",
                               input$future_runs,
                               " using a ", input$significance * 100,
                               "% limit."))
output$uitext0a = renderUI({
  if (mact_EVAL()){
    htmlOutput(outputId = 'text0a')
  }
})
uiOutput(outputId = 'uitext0a')
```

```{r text0b}
output$text0b = renderText(paste0("The emissions data were pre-selected to the top performing sources to use for the UPL calculation, thus the entire uploaded data set was used to characterize the emissions distribution. The number of future tests for calculating the UPL was ",
                               input$future_runs,
                               " using a ", input$significance * 100,
                               "% limit."))
output$uitext0b = renderUI({
  if (!mact_EVAL()){
    htmlOutput(outputId = 'text0b')
  }
})
uiOutput(outputId = 'uitext0b')
```

##### The best performer for the new source calculation is `r reactive(best_source())` with an average emission of `r reactive(signif(emission_mean_new(),digits=3))` from `r reactive(nrow(dat_top_new_raw()))` test runs.

```{r source_settings}

dat_top_exist_raw = reactive({
  if (input$runMACT){
    dat_out = MACT_existing(CAA_section = input$CAA_section,
                                   data = dat_emiss())
  } else {
    dat_out = dat_emiss()
    dat_out = subset(dat_out, dat_out$emissions != 0)
  }
  dat_out
})
  
emission_mean_exist = reactive({
  dat1 = dat_top_exist()
  results1 = mean(dat1$emissions)
  results1
  })

dat_topmeans_exist_raw = reactive(dat_top_exist_raw() %>% group_by(sources) %>%
                              summarize(means = mean(emissions) ,counts = n()))
dat_topmeans_exist = reactive({
  ordered = dat_topmeans_exist_raw()
  ordered$sources = as.factor(ordered$sources)
  ordered$sources = fct_reorder(ordered$sources,
                              ordered$means, .desc = FALSE)
  ordered = arrange(ordered, means)
  ordered = ordered %>% mutate(across(where(is.numeric), ~signif(., digits = 4)))
  names(ordered) = c('sources', 'means', 'counts')
  ordered
})

dat_top_new_raw = reactive(MACT_new(data = dat_emiss()))
emission_mean_new = reactive({
  dat2 = dat_top_new_raw()
  results2 = mean(dat2$emissions)
  results2
  })
best_source = reactive({
  dat3 = dat_top_new_raw()
  result3 = as.character(dat3$sources)[1]
  result3
})

dat_top_exist = reactive({
  ordered2 = dat_top_exist_raw()
  ordered3 = dat_topmeans_exist()
  ordered2$sources = factor(ordered2$sources, levels = levels(ordered3$sources))
  ordered2$ln_emiss = log(ordered2$emissions)
  ordered2$ln_emiss = replace(ordered2$ln_emiss,
                               !is.finite(ordered2$ln_emiss), NA)
  ordered2
})

output$Sourcedownload1 = downloadHandler(
  filename = function() {
    paste0(str_remove(input$data$name, ".csv"), "_TopSources_v", Sys.Date(), ".csv")
  },
  content = function(file) {
    write.csv(dat_top_exist(), file)
  }
)

output$mean_display = reactive(dat_topmeans_exist() %>% head(5) %>% 
                                 knitr::kable('html', caption = "Top five sources for existing source calculation", digits = 12, col.names = c('Source','Average emission','No. of Runs')) %>% 
                kable_styling("striped", full_width = T))
tableOutput('mean_display')
```

```{r nsps}
dat_top_new = reactive({
  ordered2 = dat_top_new_raw()
  ordered2$ln_emiss = log(ordered2$emissions)
  ordered2$ln_emiss = replace(ordered2$ln_emiss,
                               !is.finite(ordered2$ln_emiss), NA)
  ordered2
})
output$Sourcedownload2 = downloadHandler(
  filename = function() {
    paste0(str_remove(input$data$name, ".csv"), "_BestSource_v", Sys.Date(), ".csv")
  },
  content = function(file) {
    write.csv(dat_top_new(), file)
  }
)
nsps_EVAL = reactive({
  if (nrow(dat_top_new()) < 3){
    out = FALSE
  } else {
    out = TRUE
  }
  out
})


distribution_result_new = reactive({
    # validate(
    # need((nrow(dat_top_new_raw())>=3),
    #      "data must have at least 3 observations"))
  req(nsps_EVAL())
  distribution_type(data = dat_top_new())
  })

PI99_normal_new = reactive({
  if (nsps_EVAL()){
      df4 = dat_top_new()
      out3 = Normal_UPL(data = df4, future_runs = input$future_runs,
                             significance = input$significance)
  } else {
    out3 = NA
  }
  out3
  })
PI99_lognorm_new = reactive({
  if (nsps_EVAL()){
    df5 = dat_top_new()
    out4 = Lognormal_UPL(data = df5, future_runs = input$future_runs,
                             significance = input$significance)    
  } else{
    out4 = NA
  }
  out4
  })
PI99_skew_new = reactive({
  if (nsps_EVAL()){
    df6 = dat_top_new()
    out5 = Skewed_UPL(data = df6, future_runs = input$future_runs,
                             significance = input$significance)    
  } else{
    out5 = NA
  }
  out5
  })

dat_freq1_new = reactive({
    output = tibble(distribution = c('Normal', 'Skewed', 'Lognormal'),
                UPL_new = c(PI99_normal_new(),
                          PI99_skew_new(),
                          PI99_lognorm_new()))
  output
  })
dat_freq_new = reactive({
  dat_freq2_new = dat_freq1_new()
  dat_freq3_new = dat_freq2_new %>% mutate(across(where(is.numeric), ~signif(. ,digits = 4)))
  dat_freq3_new
  })

best_UPL_NSPS = reactive({
  req(nsps_EVAL())
  dat_freq1 = dat_freq_new()
  dat_freq1$UPL_new[dat_freq1$distribution == distribution_result_new()]
})

x_hat_new = reactive({
  req(nsps_EVAL())
  df = dat_top_new()
  seq(0, 3 * max(df$emissions), length.out = 1024)
})
obs_dens_results_new = reactive({
  req(nsps_EVAL())
  obs_density(data = dat_top_new(), xvals = x_hat_new())
}

)
Obs_onPoint_new = reactive({
  req(nsps_EVAL())
  out1 = obs_dens_results_new()
  out1$Obs_onPoint
})
obs_den_df_new = reactive({
  req(nsps_EVAL())
  out1 = obs_dens_results_new()
  out1$obs_den_df
}) 
pred_dat_new = reactive({
  req(nsps_EVAL())
  df = dat_top_new()
  pdf_norm = dnorm(x_hat_new(), mean = emission_mean_new(), sd = sd(df$emissions))
  pdf_ln = dlnorm(x_hat_new(), mean = mean(df$ln_emiss, na.rm = T),
              sd = sd(df$ln_emiss, na.rm = T))
  tibble(pdf_norm = pdf_norm, pdf_ln = pdf_ln, x_hat = x_hat_new())
})
```

```{r figure1}
source_fig = reactive({
  dat_temp = dat_topmeans_exist()
  if (all(dat_temp$counts == 1)){
    output = FALSE
  } else {
    output = TRUE
  }
    output
})
output$plot1 = renderPlot({
      ggplot(data = dat_top_exist(), aes(emissions, fill = sources))+
          geom_density(alpha = 0.75, bounds = c(0, Inf))+
          ylab("Density")+ xlab("Emissions")+
          ggtitle("Top sources observed populations")+
          multi_source_theme()+ labs(fill = 'Top sources', color = 'Top sources')+
          geom_vline(aes(xintercept = means, color = sources),
             data = dat_topmeans_exist(), linewidth = 1)+
          guides(fill = guide_legend(nrow = 2, byrow = TRUE),
                color = guide_legend(nrow = 2, byrow = TRUE))+
          scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
          scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
          coord_cartesian(clip = 'off')+
          geom_rug(sides = 'b', aes(x = emissions, color = sources),
                   data = dat_top_exist(), alpha = 0.5, outside = TRUE)

})

output$uiplot1 = renderUI({
  if (source_fig()){
    plotOutput(outputId = 'plot1')
  }
})
uiOutput(outputId = 'uiplot1')
```

```{r label1}
output$text1 = renderText("Figure 1. Observation density of emissions by source type. Source averages are indicated by vertical lines.")

output$uitext1 = renderUI({
  if (source_fig()){
    htmlOutput(outputId = 'text1')
  }
})
uiOutput(outputId = 'uitext1')
```

```{r frequentist}
distribution_result_exist = reactive(distribution_type(data = dat_top_exist()))

PI99_normal_exist = reactive(
  Normal_UPL(data = dat_top_exist(), future_runs = input$future_runs,
                  significance = input$significance)
)
PI99_lognorm_exist = reactive({
  df2 = dat_top_exist()
  out1 = Lognormal_UPL(data = df2, future_runs = input$future_runs,
                             significance = input$significance)
  out1
  })
PI99_skew_exist = reactive({
  df3 = dat_top_exist()
  out2 = Skewed_UPL(data = df3, future_runs = input$future_runs,
                             significance = input$significance)
  out2
  })

dat_freq1 = reactive({
  pi1 = PI99_normal_exist()
  pi2 = PI99_skew_exist()
  pi3 = PI99_lognorm_exist()
  tibble(distribution = c('Normal', 'Skewed', 'Lognormal'),
                UPL_exist = c(pi1, pi2, pi3))
  })
dat_freq = reactive({
  dat_freq2 = dat_freq1()
  dat_freq3 = dat_freq2 %>% mutate(across(where(is.numeric), ~signif(., digits = 4)))
  dat_freq3
  })

best_UPL = reactive({
  dat_freq1 = dat_freq()
  dat_freq1$UPL_exist[dat_freq1$distribution == distribution_result_exist()]
})

x_hat = reactive({
  df = dat_top_exist()
  seq(0, 3 * max(df$emissions), length.out = 1024)
})

obs_dens_results = reactive(
  obs_density(data = dat_top_exist(), xvals = x_hat())
)
Obs_onPoint = reactive({
  out1 = obs_dens_results()
  out1$Obs_onPoint
})
obs_den_df = reactive({
  out1 = obs_dens_results()
  out1$obs_den_df
})

pred_dat = reactive({
  df = dat_top_exist()
  pdf_norm = dnorm(x_hat(), mean = emission_mean_exist(), sd = sd(df$emissions))
  pdf_ln = dlnorm(x_hat(), mean = mean(df$ln_emiss, na.rm = T),
              sd = sd(df$ln_emiss, na.rm = T))
  tibble(pdf_norm = pdf_norm, pdf_ln = pdf_ln, x_hat = x_hat())
})

```

## UPL Results

The following calculations replicate the results from the UPL excel workbook. 

The best distribution for the existing source UPL calculation is 
`r reactive(distribution_result_exist())`, and the corresponding UPL is `r reactive(best_UPL())`.

```{r text2}
output$text2 = renderText(paste0("The best distribution for the new source UPL calculation is ",
                               distribution_result_new(), 
                               " and the corresponding UPL is ", 
                               best_UPL_NSPS(), "."))

output$uitext2 = renderUI({
  if (nsps_EVAL()){
    htmlOutput(outputId = 'text2')
  }
})
uiOutput(outputId = 'uitext2')
```

```{r freqtable}
dat_UPL = reactive({
  dat1 = dat_freq()
  dat2 = dat_freq_new()
  output = full_join(dat1, dat2, by = 'distribution')
  output
})
output$UPLdownload = downloadHandler(
  filename = function() {
    paste0(str_remove(input$data$name, ".csv"),"_UPL_v", Sys.Date(), ".csv")
  },
  content = function(file) {
    write.csv(dat_UPL(), file)
  }
)
output$upl_display = reactive(dat_UPL() %>% knitr::kable('html',
                            caption = paste0("UPL results from frequentist excel 
                                           workbook using a significance of ",
                                           input$significance * 100,'% and ',
                                           input$future_runs, ' future test runs.'),
              col.names = c('Distribution', 'Existing source UPL', 'new source UPL'), digits = 10)%>%
                kable_styling("striped", full_width = T))
tableOutput('upl_display')
```

\vspace{2cm}

```{r figure2}
renderPlot({
  ggplot(data = Obs_onPoint())+
  geom_line(data = obs_den_df(), aes(y = ydens, x = x_hat, color = 'a'),
            key_glyph = 'rect', linewidth = 0.75)+
  geom_area(data = obs_den_df(), aes(y = ydens, x = x_hat), alpha = 0.25,
            fill = 'black')+
  geom_point(aes(y = ydens, x = emissions),
             size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_line(aes(y = pdf_norm, x = x_hat, color = 'c'), key_glyph = 'rect',
               data = pred_dat(), linewidth = 0.75, linetype = 3)+
  geom_area(aes(y = pdf_norm, x = x_hat), alpha = 0.25,
            data = pred_dat(), fill = '#FF7F00')+
  geom_line(aes(y = pdf_ln, x = x_hat, color = 'b'), key_glyph = 'rect',
               data = pred_dat(), linewidth = 0.75, linetype = 2)+
  geom_area(aes(y = pdf_ln, x = x_hat), alpha = 0.25,
            data = pred_dat(), fill = '#984EA3')+
  ylab("Density")+ xlab("Emissions")+
  ggtitle("Existing source test run population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept = emission_mean_exist(), color = 'a'),
             linewidth = 1, linetype = 1, key_glyph = 'rect')+
  geom_vline(aes(xintercept = PI99_normal_exist(), color = 'c'),
             linewidth = 1, linetype = 3, key_glyph = 'rect')+
  geom_vline(aes(xintercept = PI99_lognorm_exist(), color = 'b'),
             linewidth = 1, linetype = 2, key_glyph = 'rect')+
  geom_vline(aes(xintercept = PI99_skew_exist(), color = 'd'),
             linewidth = 1, linetype = 4, key_glyph = 'rect')+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  coord_cartesian(clip = 'off')+ labs(color = "Distribution:")+
  geom_rug(sides = 'b', aes(x = emissions), data = dat_top_exist(),
           alpha = 0.5, outside = TRUE, color = 'black')+
  scale_color_manual(values = c('black', '#984EA3', '#FF7F00', '#377EB8'),
                     labels = c('Observations', 'Lognormal', 'Normal', 'Skewed'))+
  guides(fill = 'none')
})
```

Figure 2. Observation density of emission test run population. Test run average 
and existing source UPL calculations are indicated by vertical lines.

\vspace{2cm}

```{r figure3}
output$plot2 = renderPlot({
  ggplot(data = Obs_onPoint_new())+
  geom_line(data = obs_den_df_new(), aes(y = ydens, x = x_hat, color = 'a'),
            key_glyph = 'rect', linewidth = 0.75)+
  geom_area(data = obs_den_df_new(), aes(y = ydens, x = x_hat), alpha = 0.25,
            fill = 'black')+
  geom_point(aes(y = ydens, x = emissions),
             size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_line(aes(y = pdf_norm, x = x_hat, color = 'c'), key_glyph = 'rect',
               data = pred_dat_new(), linewidth = 0.75, linetype = 3)+
  geom_area(aes(y = pdf_norm, x = x_hat), alpha = 0.25,
            data = pred_dat_new(), fill = '#FF7F00')+
  geom_line(aes(y = pdf_ln, x = x_hat, color = 'b'), key_glyph = 'rect',
               data = pred_dat_new(), linewidth = 0.75, linetype = 2)+
  geom_area(aes(y = pdf_ln, x = x_hat), alpha = 0.25,
            data = pred_dat_new(), fill = '#984EA3')+
  ylab("Density")+ xlab("Emissions")+
  ggtitle("New source test run population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept = emission_mean_new(), color = 'a'),
             linewidth = 1, linetype = 1, key_glyph = 'rect')+
  geom_vline(aes(xintercept = PI99_normal_new(), color = 'c'),
             linewidth = 1, linetype = 3, key_glyph = 'rect')+
  geom_vline(aes(xintercept = PI99_lognorm_new(), color = 'b'),
             linewidth = 1, linetype = 2, key_glyph = 'rect')+
  geom_vline(aes(xintercept = PI99_skew_new(), color = 'd'),
             linewidth = 1, linetype = 4, key_glyph = 'rect')+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  coord_cartesian(clip = 'off')+ labs(color = 'Distribution:')+
  geom_rug(sides = 'b', aes(x = emissions), data = dat_top_new(),
           alpha = 0.5, outside = TRUE, color = 'black')+
  scale_color_manual(values = c('black', '#984EA3', '#FF7F00', '#377EB8'),
                     labels = c('Observations', 'Lognormal', 'Normal', 'Skewed'))
})
output$uiplot2 = renderUI({
  if (nsps_EVAL()){
    plotOutput(outputId = 'plot2')
  }
})
uiOutput(outputId = 'uiplot2')
```

```{r label3}
output$text3 = renderText(paste0("Figure 3. Observation density of best performer (",
                        best_source(), 
                        ") test run population. Best performer average and new source UPL calculations are indicated by vertical lines."))

output$uitext3 = renderUI({
  if (nsps_EVAL()){
    htmlOutput(outputId = 'text3')
  }
})
uiOutput(outputId = 'uitext3')
```

\vspace{4cm}

```{r render_pdf}
output$pdf_report = downloadHandler(
  filename = function() {
    paste0(str_remove(input$data$name, ".csv"), '_UPL_report_v', Sys.Date(), '.pdf')
  },
  content =function(file){
    rmarkdown::render("templates/App_report_template.Rmd",
                      output_file=file,
                      params = list(
                          data = dat_top_exist(),
                          future_runs = input$future_runs,
                          significance = input$significance,
                          CAA_section = input$CAA_section,
                          runMACT = input$runMACT,
                          data_nsps = dat_top_new(),
                          dat_UPL = dat_UPL(),
                          zero_warn = zero_warn()
                      ))
  }
)
```
