---
title: "UPL calculations"
author: "OAR-OAQPS-SPPD-MPG"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
runtime: shiny
geometry: left = 2cm, right = 2cm, top = 2cm, bottom = 2cm
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error=TRUE,warning=FALSE, message=FALSE)

library(tidyverse)
library(kableExtra)
library(UPLforOAR)
```

```{r inputs, echo=FALSE}
inputPanel(
        selectInput("CAA_section", label="CAA Section:", selected=112,
                choices=c(112,129)),
        fileInput('data','Upload csv file with "emissions" 
                   and "sources" as columns',
                  accept='.csv'),
        checkboxInput('runMACT','Subset data to top performers', FALSE),
        numericInput("future_runs",label="Number of future test runs",
                     value=3,min=1),
        numericInput("signficance",label="Level of significance",
                     value=0.99,min=0,max=1)
)
downloadButton('Sourcedownload','Download top sources')
downloadButton('UPLdownload','Download UPL results')

```

## Emission dataset

Check Table 1 below to ensure you have loaded the correct emission data set and 
that the values are generally what you expect.
```{r load_data}
dat_emiss <- reactive({
    req(input$data)
    ext <- tools::file_ext(input$data$name)
    switch(ext,
      csv = read.csv(input$data$datapath),
      validate("Invalid file; Please upload a .csv file")
    )
})


output$dat_display=reactive({
  validate(
    need(("emissions" %in% names(dat_emiss())),
         "data must have numeric column named 'emissions'"),
    need(("sources" %in% names(dat_emiss())),
         "data must have a column named 'sources'")
  )
  summary(dat_emiss())%>%knitr::kable('html',caption='Summary of dataset')%>%
                kable_styling("striped",full_width=T)
})
  
tableOutput('dat_display')

CAA_text=reactive(
  if (input$CAA_section==112){
  'CAA Section 112(d)(3)'
} else if (input$CAA_section==129){
  'CAA Section 129(a)(2)'
}
)

```

## Source calculation

For existing source standards calculations, `r CAA_text` is used to determine 
the top sources, either the best 5 performers or the best 12\% performers. The 
number of future tests for calculating the UPL is 
`r reactive(input$future_runs)`, using a 
`r (reactive(input$signficance*100))`\% limit.

##### The best performer for the new source calculation is `r reactive(best_source())` with an average emission of `r reactive(signif(emission_mean_new(),digits=3))` from `r reactive(nrow(dat_top_new_raw()))` test runs.

```{r source_settings}

dat_top_exist_raw=reactive({
  if (input$runMACT){
    dat_out=MACT_EG(CAA_section=input$CAA_section,
                                   data=dat_emiss())
  } else {
    dat_out=dat_emiss()
  }
  dat_out
})
  
emission_mean_exist=reactive({
  dat1=dat_top_exist()
  results1=mean(dat1$emissions)
  results1
  })

dat_topmeans_exist_raw=reactive(dat_top_exist_raw()%>%group_by(sources)%>%
                              summarize(means=mean(emissions),counts=n()))
dat_topmeans_exist=reactive({
  ordered=dat_topmeans_exist_raw()
  ordered$sources=as.factor(ordered$sources)
  ordered$sources=fct_reorder(ordered$sources,
                              ordered$means,.desc = FALSE)
  ordered=arrange(ordered,means)
  ordered=ordered%>%mutate(across(where(is.numeric),~signif(.,digits=4)))
  names(ordered)=c('sources','means','counts')
  ordered
})

dat_top_new_raw=reactive(MACT_NSPS(data=dat_emiss()))
emission_mean_new=reactive({
  dat2=dat_top_new_raw()
  results2=mean(dat2$emissions)
  results2
  })
best_source=reactive({
  dat3=dat_top_new_raw()
  result3=as.character(dat3$sources)
  result3
})

dat_top_exist=reactive({
  ordered2=dat_top_exist_raw()
  ordered2$ln_emiss=log(ordered2$emissions)
  ordered2$ln_emiss=replace(ordered2$ln_emiss,
                               !is.finite(ordered2$ln_emiss),NA)
  ordered2
})

output$Sourcedownload <- downloadHandler(
  filename = function() {
    paste0(str_remove(input$data$name,".csv"),"_TopSources_v",Sys.Date(), ".csv")
  },
  content = function(file) {
    write.csv(dat_top_exist(), file)
  }
)
output$mean_display=reactive(dat_topmeans_exist()%>%knitr::kable('html',caption="Top sources for existing source calculation",digits=12,
             col.names=c('Source','Average emission','No. of Runs'))%>%
                kable_styling("striped",full_width=T))
tableOutput('mean_display')
```

```{r nsps}
dat_top_new=reactive({
  ordered2=dat_top_new_raw()
  ordered2$ln_emiss=log(ordered2$emissions)
  ordered2$ln_emiss=replace(ordered2$ln_emiss,
                               !is.finite(ordered2$ln_emiss),NA)
  ordered2
})

distribution_result_new=reactive({
    validate(
    need((nrow(dat_top_new_raw())>=3),
         "data must have at least 3 observations"))
  distribution_type(data=dat_top_new())
  })

PI99_normal_new=reactive({
  df4=dat_top_new()
  out3=Normal_UPL(data=df4,future_runs = input$future_runs,
                             significance=input$signficance)
  out3
  })
PI99_lognorm_new=reactive({
  df5=dat_top_new()
  out4=Lognormal_UPL(data=df5,future_runs = input$future_runs,
                             significance=input$signficance)
  out4
  })
PI99_skew_new=reactive({
  df6=dat_top_new()
  out5=Skewed_UPL(data=df6,future_runs = input$future_runs,
                             significance=input$signficance)
  out5
  })

dat_freq1_new=reactive({
  pi4=PI99_normal_new()
  pi5=PI99_skew_new()
  pi6=PI99_lognorm_new()
  if (is.numeric(pi4)){
    output=tibble(distribution=c('Normal','Skew','Lognormal'),
                UPL_new=c(pi4,pi5,pi6))
  } else {
    output=tibble(distribution=c('Normal','Skew','Lognormal'),
                UPL_new=c(NA,NA,NA))
  }
  output
  })
dat_freq_new=reactive({
  dat_freq2_new=dat_freq1_new()
  dat_freq3_new=dat_freq2_new%>%mutate(across(where(is.numeric),~signif(.,digits=4)))
  dat_freq3_new
  })

best_UPL_NSPS=reactive({
  validate(
  need((nrow(dat_top_new_raw())>=3),
        "data must have at least 3 observations"))
  dat_freq1=dat_freq_new()
  dat_freq1$UPL_new[dat_freq1$distribution==distribution_result_new()]
})

x_hat_new=reactive({
  validate(
  need((nrow(dat_top_new_raw())>=3),
         "data must have at least 3 observations"))
  df=dat_top_new()
  seq(0,3*max(df$emissions),length.out=1024)
})
obs_dens_results_new=reactive({
  validate(
  need((nrow(dat_top_new_raw())>=3),
        "data must have at least 3 observations"))
  obs_density(data=dat_top_new(),xvals=x_hat_new())
}

)
Obs_onPoint_new=reactive({
  validate(
  need((nrow(dat_top_new_raw())>=3),
        "data must have at least 3 observations"))
  out1=obs_dens_results_new()
  out1$Obs_onPoint
})
obs_den_df_new=reactive({
  validate(
  need((nrow(dat_top_new_raw())>=3),
        "data must have at least 3 observations"))
  out1=obs_dens_results_new()
  out1$obs_den_df
}) 
pred_dat_new=reactive({
  validate(
  need((nrow(dat_top_new_raw())>=3),
        "data must have at least 3 observations"))
  df=dat_top_new()
  pdf_norm=dnorm(x_hat_new(),mean=emission_mean_new(),sd=sd(df$emissions))
  pdf_ln=dlnorm(x_hat_new(),mean=mean(df$ln_emiss,na.rm=T),
              sd=sd(df$ln_emiss,na.rm=T))
  tibble(pdf_norm=pdf_norm,pdf_ln=pdf_ln,x_hat=x_hat_new())
})
```

```{r figure1}
renderPlot({
  ggplot(data=dat_top_exist(),aes(emissions,fill=sources))+
          geom_density(alpha=0.75,bounds=c(0,Inf))+
          ylab("Density")+xlab("Emissions")+
          ggtitle("Top sources observed populations")+
          multi_source_theme()+labs(fill='Top sources',color='Top sources')+
          geom_vline(aes(xintercept=means,color=sources),
             data=dat_topmeans_exist(),linewidth=1)+
          guides(fill=guide_legend(nrow=2,byrow=TRUE),
                color=guide_legend(nrow=2,byrow=TRUE))+
          scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
          scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
          coord_cartesian(clip='off')+
          geom_rug(sides='b',aes(x=emissions,color=sources),
                   data=dat_top_exist(),alpha=0.5,outside=TRUE)
})
```

Figure 1. Observation density of emissions by source type. Source averages are 
indicated by vertical lines.

```{r frequentist}
distribution_result_exist=reactive(distribution_type(data=dat_top_exist()))

PI99_normal_exist=reactive(
  Normal_UPL(data=dat_top_exist(),future_runs = input$future_runs,
                  significance=input$signficance)
)
PI99_lognorm_exist=reactive({
  df2=dat_top_exist()
  out1=Lognormal_UPL(data=df2,future_runs = input$future_runs,
                             significance=input$signficance)
  out1
  })
PI99_skew_exist=reactive({
  df3=dat_top_exist()
  out2=Skewed_UPL(data=df3,future_runs = input$future_runs,
                             significance=input$signficance)
  out2
  })

dat_freq1=reactive({
  pi1=PI99_normal_exist()
  pi2=PI99_skew_exist()
  pi3=PI99_lognorm_exist()
  tibble(distribution=c('Normal','Skew','Lognormal'),
                UPL_exist=c(pi1,pi2,pi3))
  })
dat_freq=reactive({
  dat_freq2=dat_freq1()
  dat_freq3=dat_freq2%>%mutate(across(where(is.numeric),~signif(.,digits=4)))
  dat_freq3
  })

best_UPL=reactive({
  dat_freq1=dat_freq()
  dat_freq1$UPL_exist[dat_freq1$distribution==distribution_result_exist()]
})

x_hat=reactive({
  df=dat_top_exist()
  seq(0,3*max(df$emissions),length.out=1024)
})

obs_dens_results=reactive(
  obs_density(data=dat_top_exist(),xvals=x_hat())
)
Obs_onPoint=reactive({
  out1=obs_dens_results()
  out1$Obs_onPoint
})
obs_den_df=reactive({
  out1=obs_dens_results()
  out1$obs_den_df
})

pred_dat=reactive({
  df=dat_top_exist()
  pdf_norm=dnorm(x_hat(),mean=emission_mean_exist(),sd=sd(df$emissions))
  pdf_ln=dlnorm(x_hat(),mean=mean(df$ln_emiss,na.rm=T),
              sd=sd(df$ln_emiss,na.rm=T))
  tibble(pdf_norm=pdf_norm,pdf_ln=pdf_ln,x_hat=x_hat())
})

```

## UPL Results

The following calculations replicate the results from the UPL excel workbook. 
The best distribution for the existing source UPL calculation is 
`r reactive(distribution_result_exist())`, and the corresponding UPL is `r reactive(best_UPL())`. 
The best distribution for the NSPS UPL calculation is 
`r reactive(distribution_result_new())`, and the corresponding UPL is `r reactive(best_UPL_NSPS())`.

```{r freqtable}
dat_UPL=reactive({
  dat1=dat_freq()
  dat2=dat_freq_new()
  output=full_join(dat1,dat2,by='distribution')
  output
})
output$UPLdownload <- downloadHandler(
  filename = function() {
    paste0(str_remove(input$data$name,".csv"),"_UPL_v",Sys.Date(), ".csv")
  },
  content = function(file) {
    write.csv(dat_UPL(), file)
  }
)
output$upl_display=reactive(dat_UPL()%>%knitr::kable('html',
                            caption=paste0("UPL results from frequentist excel 
                                           workbook using a significance of ",
                                           input$signficance*100,'% and ',
                                           input$future_runs,' future test runs.'),
              col.names=c('Distribution','Existing source UPL','NSPS UPL'), digits = 10)%>%
                kable_styling("striped",full_width=T))
tableOutput('upl_display')
```

\vspace{1cm}

```{r figure2}
renderPlot({
  # dataset=dat_top_exist()
  # Ymax=1.1*max(c(PI99_lognorm_exist(),PI99_normal_exist(),PI99_skew_exist(),
  # dataset$emissions))
  # Ymin=min(pred_dat())
  ggplot(data=Obs_onPoint())+
  geom_line(data=obs_den_df(),aes(y=ydens,x=x_hat,color='a'),
            key_glyph='rect',linewidth=0.75)+
  geom_area(data=obs_den_df(),aes(y=ydens,x=x_hat),alpha=0.25,
            fill='black')+
  geom_point(aes(y=ydens,x=emissions),
             size=3,alpha=0.5,shape=19,color='black')+
  geom_line(aes(y=pdf_norm,x=x_hat,color='c'),key_glyph='rect',
               data = pred_dat(),linewidth=0.75,linetype=3)+
  geom_area(aes(y=pdf_norm,x=x_hat),alpha=0.25,
            data = pred_dat(),fill='#FF7F00')+
  geom_line(aes(y=pdf_ln,x=x_hat,color='b'),key_glyph='rect',
               data = pred_dat(),linewidth=0.75,linetype=2)+
  geom_area(aes(y=pdf_ln,x=x_hat),alpha=0.25,
            data = pred_dat(),fill='#984EA3')+
  ylab("Density")+xlab("Emissions")+
  ggtitle("Existing source test run population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept=(emission_mean_exist()),color='a'),
             linewidth=1,linetype=1,key_glyph='rect')+
  geom_vline(aes(xintercept=(PI99_normal_exist()),color='c'),
             linewidth=1,linetype=3,key_glyph='rect')+
  geom_vline(aes(xintercept=(PI99_lognorm_exist()),color='b'),
             linewidth=1,linetype=2,key_glyph='rect')+
  geom_vline(aes(xintercept=(PI99_skew_exist()),color='d'),
             linewidth=1,linetype=4,key_glyph='rect')+
  scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
  scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
  coord_cartesian(clip='off')+labs(color="Distribution:")+
  geom_rug(sides='b',aes(x=(emissions)),data=dat_top_exist(),
           alpha=0.5,outside=TRUE,color='black')+
  scale_color_manual(values=c('black','#984EA3','#FF7F00','#377EB8'),
                     labels=c('Observations','Lognormal','Normal','Skewed'))+
  guides(fill='none')
})
```
Figure 2. Observation density of emission test run population. Test run average 
and existing source UPL calculations are indicated by vertical lines.

\vspace{1cm}
```{r figure3}
renderPlot({
  validate(
  need((nrow(dat_top_new_raw())>=3),
        "data must have at least 3 observations"))
  # dataset=dat_top_exist()
  # Ymax=1.1*max(c(PI99_lognorm_exist(),PI99_normal_exist(),PI99_skew_exist(),dataset$emissions))
  # Ymin=min(pred_dat())
  ggplot(data=Obs_onPoint_new())+
  geom_line(data=obs_den_df_new(),aes(y=ydens,x=(x_hat),color='a'),
            key_glyph='rect',linewidth=0.75)+
  geom_area(data=obs_den_df_new(),aes(y=ydens,x=(x_hat)),alpha=0.25,
            fill='black')+
  geom_point(aes(y=ydens,x=(emissions)),
             size=3,alpha=0.5,shape=19,color='black')+
  geom_line(aes(y=pdf_norm,x=(x_hat),color='c',key_glyph='rect'),
               data = pred_dat_new(),linewidth=0.75,linetype=3)+
  geom_area(aes(y=pdf_norm,x=(x_hat)),alpha=0.25,
            data = pred_dat_new(),fill='#FF7F00')+
  geom_line(aes(y=pdf_ln,x=(x_hat),color='b',key_glyph='rect'),
               data = pred_dat_new(),linewidth=0.75,linetype=2)+
  geom_area(aes(y=pdf_ln,x=(x_hat)),alpha=0.25,
            data = pred_dat_new(),fill='#984EA3')+
  ylab("Density")+xlab("Emissions")+
  ggtitle("New source test run population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept=(emission_mean_new()),color='a'),
             linewidth=1,linetype=1,key_glyph='rect')+
  geom_vline(aes(xintercept=(PI99_normal_new()),color='c'),
             linewidth=1,linetype=3,key_glyph='rect')+
  geom_vline(aes(xintercept=(PI99_lognorm_new()),color='b'),
             linewidth=1,linetype=2,key_glyph='rect')+
  geom_vline(aes(xintercept=(PI99_skew_new()),color='d'),
             linewidth=1,linetype=4,key_glyph='rect')+
  scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
  scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
  coord_cartesian(clip='off')+labs(color='Distribution:')+
  geom_rug(sides='b',aes(x=(emissions)),data=dat_top_new(),
           alpha=0.5,outside=TRUE,color='black')+
  scale_color_manual(values=c('black','#984EA3','#FF7F00','#377EB8'),
                     labels=c('Observations','Lognormal','Normal','Skewed'))
})
```
Figure 3. Observation density of best performer (`r reactive(best_source())`) 
test run population. Best performer average and new source UPL calculations are 
indicated by vertical lines.

\vspace{2cm}
