---
title: "UPL Results"
author: "Ludda Ludwig"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
params:
  data: "NA"
  future_runs: "NA"
  significance: "NA"
  CAA_section: "NA"
  runMACT: "NA"
  data_nsps: "NA"
  dat_UPL: "NA"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,error=TRUE,warning=FALSE, message=FALSE)
library(tidyverse)
library(kableExtra)
library(UPLforOAR)
```

```{r data}
dat_top_exist = params$data

if (params$CAA_section==112){
  CAA_text='CAA Section 112(d)(3)'
} else if (params$CAA_section==129){
  CAA_text='CAA Section 129(a)(2)'
}

emission_mean_exist=results1=mean(dat_top_exist$emissions)
dat_topmeans_exist=dat_top_exist%>%group_by(sources)%>%
                              summarize(means=mean(emissions),counts=n())
dat_topmeans_exist$sources=as.factor(dat_topmeans_exist$sources)
dat_topmeans_exist$sources=fct_reorder(dat_topmeans_exist$sources,
                              dat_topmeans_exist$means,.desc = FALSE)
dat_topmeans_exist=arrange(dat_topmeans_exist,means)
dat_topmeans_exist=dat_topmeans_exist%>%mutate(across(where(is.numeric),~signif(.,digits=4)))
names(dat_topmeans_exist)=c('sources','means','counts')

dat_top_new=params$data_nsps
emission_mean_new=mean(dat_top_new$emissions)
best_source=as.character(dat_top_new$sources)[1]

distribution_result_new=distribution_type(data=dat_top_new)
distribution_result_exist=distribution_type(data=dat_top_exist)

UPL_table=params$dat_UPL

best_UPL=UPL_table$UPL_exist[UPL_table$distribution==distribution_result_exist]
best_UPL_NSPS=UPL_table$UPL_new[UPL_table$distribution==distribution_result_new]


```

Upper Predictive Limit (UPL) results were provided following the frequentist 
framework and methods previously used by Excel workbooks, now implemented in 
`UPLforOAR` version `r packageVersion('UPLforOAR')`. Density figures were 
generated using `ggplot2` and `obs_density()` with a bandwidth of `r signif(sd(dat_top_exist$emissions)*nrow(dat_top_exist)^(-2/5),3)`, the `gamma` kernel, and 
an upper boundary of `r signif(3*max(dat_top_exist$emissions),3)` emissions. 

## Source calculation

For existing source standards calculations, `r CAA_text` was used to determine 
the top sources, either the best 5 performers or the best 12\% performers. The 
number of future tests for calculating the UPL is 
`r params$future_runs`, using a `r (params$significance*100)`\% limit.

##### The best performer for the new source calculation is `r best_source` with an average emission of `r signif(emission_mean_new,digits=4)` from `r nrow(dat_top_new)` test runs.

## UPL Results

Based on the kurtosis and skewness ratio test methods from the UPL excel workbook,
The best distribution for the existing source UPL calculation is 
`r distribution_result_exist`, with the corresponding UPL: `r best_UPL`, and 
the best distribution for the NSPS UPL calculation is 
`r distribution_result_new`, and the corresponding UPL is `r best_UPL_NSPS`.


```{r UPL_table}
UPL_table%>%knitr::kable(caption=paste0("UPL results from frequentist excel 
                                           workbook using a significance of ",
                                           (params$significance*100),'% and ',
                                           params$future_runs,' future test runs.'),
              col.names=c('Distribution','Existing source UPL','NSPS UPL'), digits = 10)
```

For each distribution used, a pair of plots is generated for each defining 
parameter. 

```{r plot1, results=FALSE, fig.height=4, fig.cap="Figure 1. Observation density of emissions by source type. Source averages are indicated by vertical lines."}
  ggplot(data=dat_top_exist,aes(emissions,fill=sources))+
          geom_density(alpha=0.75,bounds=c(0,Inf))+
          ylab("Density")+xlab("Emissions")+
          ggtitle("Top sources observed populations")+
          multi_source_theme()+labs(fill='Top sources',color='Top sources')+
          geom_vline(aes(xintercept=means,color=sources),
             data=dat_topmeans_exist,linewidth=1)+
          guides(fill=guide_legend(nrow=2,byrow=TRUE),
                color=guide_legend(nrow=2,byrow=TRUE))+
          scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
          scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
          coord_cartesian(clip='off')+
          geom_rug(sides='b',aes(x=emissions,color=sources),
                   data=dat_top_exist,alpha=0.5,outside=TRUE)
```

```{r plot2, fig.height=4, fig.cap="Figure 2. Observation density of emission test run population. Test run average and existing source UPL calculations are indicated by vertical lines."}

x_hat=seq(0,3*max(dat_top_exist$emissions),length.out=1024)
obs_dens_results=obs_density(data=dat_top_exist,xvals=x_hat)
Obs_onPoint=obs_dens_results$Obs_onPoint
obs_den_df=obs_dens_results$obs_den_df
pdf_norm=dnorm(x_hat,mean=emission_mean_exist,sd=sd(dat_top_exist$emissions))
pdf_ln=dlnorm(x_hat,mean=mean(dat_top_exist$ln_emiss,na.rm=T),
              sd=sd(dat_top_exist$ln_emiss,na.rm=T))
pred_dat=tibble(pdf_norm=pdf_norm,pdf_ln=pdf_ln,x_hat=x_hat)

ggplot(data=Obs_onPoint)+
  geom_line(data=obs_den_df,aes(y=ydens,x=x_hat,color='a'),
            key_glyph='rect',linewidth=0.75)+
  geom_area(data=obs_den_df,aes(y=ydens,x=x_hat),alpha=0.25,
            fill='black')+
  geom_point(aes(y=ydens,x=emissions),
             size=3,alpha=0.5,shape=19,color='black')+
  geom_line(aes(y=pdf_norm,x=x_hat,color='c'),key_glyph='rect',
               data = pred_dat,linewidth=0.75,linetype=3)+
  geom_area(aes(y=pdf_norm,x=x_hat),alpha=0.25,
            data = pred_dat,fill='#FF7F00')+
  geom_line(aes(y=pdf_ln,x=x_hat,color='b'),key_glyph='rect',
               data = pred_dat,linewidth=0.75,linetype=2)+
  geom_area(aes(y=pdf_ln,x=x_hat),alpha=0.25,
            data = pred_dat,fill='#984EA3')+
  ylab("Density")+xlab("Emissions")+
  ggtitle("Existing source test run population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept=(emission_mean_exist),color='a'),
             linewidth=1,linetype=1,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_exist[UPL_table$distribution=="Normal"],color='c'),
             linewidth=1,linetype=3,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_exist[UPL_table$distribution=="Lognormal"],color='b'),
             linewidth=1,linetype=2,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_exist[UPL_table$distribution=="Skewed"],color='d'),
             linewidth=1,linetype=4,key_glyph='rect')+
  scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
  scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
  coord_cartesian(clip='off')+labs(color="Distribution:")+
  geom_rug(sides='b',aes(x=emissions),data=dat_top_exist,
           alpha=0.5,outside=TRUE,color='black')+
  scale_color_manual(values=c('black','#984EA3','#FF7F00','#377EB8'),
                     labels=c('Observations','Lognormal','Normal','Skewed'))+
  guides(fill='none')
```

```{r plot3, fig.height=4, fig.cap="Figure 3. Observation density of best performer (`r best_source`) test run population. Best performer average and new source UPL calculations are indicated by vertical lines."}

x_hat_new=seq(0,3*max(dat_top_new$emissions),length.out=1024)
obs_dens_results_new=obs_density(data=dat_top_new,xvals=x_hat_new)
Obs_onPoint_new=obs_dens_results_new$Obs_onPoint
obs_den_df_new=obs_dens_results_new$obs_den_df
pdf_norm_new=dnorm(x_hat_new,mean=emission_mean_new,sd=sd(dat_top_new$emissions))
pdf_ln_new=dlnorm(x_hat_new,mean=mean(dat_top_new$ln_emiss,na.rm=T),
              sd=sd(dat_top_new$ln_emiss,na.rm=T))
pred_dat_new=tibble(pdf_norm=pdf_norm_new,pdf_ln=pdf_ln_new,x_hat=x_hat_new)

ggplot(data=Obs_onPoint_new)+
  geom_line(data=obs_den_df_new,aes(y=ydens,x=x_hat,color='a'),
            key_glyph='rect',linewidth=0.75)+
  geom_area(data=obs_den_df_new,aes(y=ydens,x=x_hat),alpha=0.25,
            fill='black')+
  geom_point(aes(y=ydens,x=emissions),
             size=3,alpha=0.5,shape=19,color='black')+
  geom_line(aes(y=pdf_norm,x=x_hat,color='c'),key_glyph='rect',
               data = pred_dat_new,linewidth=0.75,linetype=3)+
  geom_area(aes(y=pdf_norm,x=x_hat),alpha=0.25,
            data = pred_dat_new,fill='#FF7F00')+
  geom_line(aes(y=pdf_ln,x=x_hat,color='b'),key_glyph='rect',
               data = pred_dat_new,linewidth=0.75,linetype=2)+
  geom_area(aes(y=pdf_ln,x=x_hat),alpha=0.25,
            data = pred_dat_new,fill='#984EA3')+
  ylab("Density")+xlab("Emissions")+
  ggtitle("NSPS test run population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept=(emission_mean_new),color='a'),
             linewidth=1,linetype=1,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_new[UPL_table$distribution=="Normal"],color='c'),
             linewidth=1,linetype=3,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_new[UPL_table$distribution=="Lognormal"],color='b'),
             linewidth=1,linetype=2,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_new[UPL_table$distribution=="Skewed"],color='d'),
             linewidth=1,linetype=4,key_glyph='rect')+
  scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
  scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
  coord_cartesian(clip='off')+labs(color="Distribution:")+
  geom_rug(sides='b',aes(x=emissions),data=dat_top_new,
           alpha=0.5,outside=TRUE,color='black')+
  scale_color_manual(values=c('black','#984EA3','#FF7F00','#377EB8'),
                     labels=c('Observations','Lognormal','Normal','Skewed'))+
  guides(fill='none')
```

