---
title: "UPL Results"
author: "Ludda Ludwig"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    extra_dependencies: ["float"]
params:
  data: "NA"
  future_runs: "NA"
  significance: "NA"
  CAA_section: "NA"
  runMACT: "NA"
  data_nsps: "NA"
  dat_UPL: "NA"
  zero_warn: "NA"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error=TRUE,warning=FALSE, message=FALSE,
                      fig.pos="H",out.extra = "")
library(tidyverse)
library(kableExtra)
library(UPLforOAR)
```

```{r data, echo = FALSE}
dat_top_exist = params$data

if (params$CAA_section==112){
  CAA_text='CAA Section 112(d)(3)'
} else if (params$CAA_section==129){
  CAA_text='CAA Section 129(a)(2)'
}

emission_mean_exist=results1=mean(dat_top_exist$emissions)
dat_topmeans_exist=dat_top_exist%>%group_by(sources)%>%
                              summarize(means=mean(emissions),counts=n())
dat_topmeans_exist$sources=as.factor(dat_topmeans_exist$sources)
dat_topmeans_exist$sources=fct_reorder(dat_topmeans_exist$sources,
                              dat_topmeans_exist$means,.desc = FALSE)
dat_topmeans_exist=arrange(dat_topmeans_exist,means)
dat_topmeans_exist=dat_topmeans_exist%>%mutate(across(where(is.numeric),~signif(.,digits=4)))
names(dat_topmeans_exist)=c('sources','means','counts')
if (all(dat_topmeans_exist$counts==1)){
  source_fig = FALSE
} else {
  source_fig = TRUE
}

dat_top_new=params$data_nsps
emission_mean_new=mean(dat_top_new$emissions)
best_source=as.character(dat_top_new$sources)[1]

distribution_result_exist=distribution_type(data=dat_top_exist)

UPL_table=params$dat_UPL

best_UPL=UPL_table$UPL_exist[UPL_table$distribution==distribution_result_exist]

if (nrow(dat_top_new)<3){
  nsps_EVAL = FALSE
} else {
  distribution_result_new=distribution_type(data=dat_top_new)
  best_UPL_NSPS=UPL_table$UPL_new[UPL_table$distribution==distribution_result_new]
  nsps_EVAL = TRUE
}
if (is.null(params$runMACT)){
  mact_EVAL=FALSE
} else if (!params$runMACT){
  mact_EVAL=FALSE
} else if (params$runMACT){
  mact_EVAL=TRUE
}


```

```{r cond1a, child='App_report_children/conditional1a.Rmd', eval=nsps_EVAL}
```

```{r cond1b, child='App_report_children/conditional1b.Rmd', eval=(!nsps_EVAL)}
```

## Source Population Calculation

```{r cond5, child='App_report_children/conditional5.Rmd', eval=zero_warn}
```

```{r cond2a, child='App_report_children/conditional2a.Rmd', eval=(mact_EVAL)}
```

```{r cond2b, child='App_report_children/conditional2b.Rmd', eval=(!mact_EVAL)}
```

##### The best performer for the new source calculation was `r best_source` with an average emission of `r signif(emission_mean_new,digits=4)` from `r nrow(dat_top_new)` test runs.

```{r mean_table, echo = FALSE}
dat_topmeans_exist%>%head(5)%>%knitr::kable(caption="Top five sources for existing source calculation",digits=12,
             col.names=c('Source','Average emission','No. of Runs'))
```

## UPL Results
```{r cond3a, child='App_report_children/conditional3a.Rmd', eval=nsps_EVAL}
```

```{r cond3b, child='App_report_children/conditional3b.Rmd', eval=(!nsps_EVAL)}
```

```{r UPL_table, echo=FALSE, eval=nsps_EVAL}
UPL_table%>%knitr::kable(caption=paste0("UPL results from frequentist excel 
                                           workbook equivalent using a significance of ",
                                           (params$significance*100),'% and ',
                                           params$future_runs,' future test runs.'),
              col.names=c('Distribution','Existing source UPL','NSPS UPL'), digits = 12)
```

```{r UPL_table2, echo=FALSE, eval=(!nsps_EVAL), results=TRUE}
UPL_table=subset(UPL_table,select=-c(UPL_new))
UPL_table%>%knitr::kable(caption=paste0("UPL results from frequentist excel 
                                           workbook equivalent using a significance of ",
                                           (params$significance*100),'% and ',
                                           params$future_runs,' future test runs.'),
              col.names=c('Distribution','Existing source UPL'), digits = 12)
```

```{r plot1, echo=FALSE, eval=source_fig, results=FALSE, fig.height=4, fig.cap="Figure 1. Observation density of emissions by source type. Source averages are indicated by vertical lines."}
ggplot(data=dat_top_exist,aes(emissions,fill=sources))+
          geom_density(alpha=0.75,bounds=c(0,Inf))+
          ylab("Density")+xlab("Emissions")+
          ggtitle("Top sources observed populations")+
          multi_source_theme()+labs(fill='Top sources',color='Top sources')+
          geom_vline(aes(xintercept=means,color=sources),
             data=dat_topmeans_exist,linewidth=1)+
          guides(fill=guide_legend(nrow=2,byrow=TRUE),
                color=guide_legend(nrow=2,byrow=TRUE))+
          scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
          scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
          coord_cartesian(clip='off')+
          geom_rug(sides='b',aes(x=emissions,color=sources),
                   data=dat_top_exist,alpha=0.5,outside=TRUE)
```

```{r plot2, echo = FALSE, fig.height=4, fig.cap="Figure 2. Observation density of emission test run population. Test run average and existing source UPL calculations are indicated by vertical lines."}

x_hat=seq(0,3*max(dat_top_exist$emissions),length.out=1024)
obs_dens_results=obs_density(data=dat_top_exist,xvals=x_hat)
Obs_onPoint=obs_dens_results$Obs_onPoint
obs_den_df=obs_dens_results$obs_den_df
pdf_norm=dnorm(x_hat,mean=emission_mean_exist,sd=sd(dat_top_exist$emissions))
pdf_ln=dlnorm(x_hat,mean=mean(dat_top_exist$ln_emiss,na.rm=T),
              sd=sd(dat_top_exist$ln_emiss,na.rm=T))
pred_dat=tibble(pdf_norm=pdf_norm,pdf_ln=pdf_ln,x_hat=x_hat)

ggplot(data=Obs_onPoint)+
  geom_line(data=obs_den_df,aes(y=ydens,x=x_hat,color='a'),
            key_glyph='rect',linewidth=0.75)+
  geom_area(data=obs_den_df,aes(y=ydens,x=x_hat),alpha=0.25,
            fill='black')+
  geom_point(aes(y=ydens,x=emissions),
             size=3,alpha=0.5,shape=19,color='black')+
  geom_line(aes(y=pdf_norm,x=x_hat,color='c'),key_glyph='rect',
               data = pred_dat,linewidth=0.75,linetype=3)+
  geom_area(aes(y=pdf_norm,x=x_hat),alpha=0.25,
            data = pred_dat,fill='#FF7F00')+
  geom_line(aes(y=pdf_ln,x=x_hat,color='b'),key_glyph='rect',
               data = pred_dat,linewidth=0.75,linetype=2)+
  geom_area(aes(y=pdf_ln,x=x_hat),alpha=0.25,
            data = pred_dat,fill='#984EA3')+
  ylab("Density")+xlab("Emissions")+
  ggtitle("Existing source test run population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept=emission_mean_exist,color='a'),
             linewidth=1,linetype=1,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_exist[UPL_table$distribution=="Normal"],color='c'),
             linewidth=1,linetype=3,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_exist[UPL_table$distribution=="Lognormal"],color='b'),
             linewidth=1,linetype=2,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_exist[UPL_table$distribution=="Skewed"],color='d'),
             linewidth=1,linetype=4,key_glyph='rect')+
  scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
  scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
  coord_cartesian(clip='off')+labs(color="Distribution:")+
  geom_rug(sides='b',aes(x=emissions),data=dat_top_exist,
           alpha=0.5,outside=TRUE,color='black')+
  scale_color_manual(values=c('black','#984EA3','#FF7F00','#377EB8'),
                     labels=c('Observations','Lognormal','Normal','Skewed'))+
  guides(fill='none')
```

```{r plot3, echo=FALSE, eval=nsps_EVAL, fig.height=4, fig.cap="Figure 3. Observation density of best performer test run population. Best performer average and new source UPL calculations are indicated by vertical lines."}

x_hat_new=seq(0,3*max(dat_top_new$emissions),length.out=1024)
obs_dens_results_new=obs_density(data=dat_top_new,xvals=x_hat_new)
Obs_onPoint_new=obs_dens_results_new$Obs_onPoint
obs_den_df_new=obs_dens_results_new$obs_den_df
pdf_norm_new=dnorm(x_hat_new,mean=emission_mean_new,sd=sd(dat_top_new$emissions))
pdf_ln_new=dlnorm(x_hat_new,mean=mean(dat_top_new$ln_emiss,na.rm=T),
              sd=sd(dat_top_new$ln_emiss,na.rm=T))
pred_dat_new=tibble(pdf_norm=pdf_norm_new,pdf_ln=pdf_ln_new,x_hat=x_hat_new)

ggplot(data=Obs_onPoint_new)+
  geom_line(data=obs_den_df_new,aes(y=ydens,x=x_hat,color='a'),
            key_glyph='rect',linewidth=0.75)+
  geom_area(data=obs_den_df_new,aes(y=ydens,x=x_hat),alpha=0.25,
            fill='black')+
  geom_point(aes(y=ydens,x=emissions),
             size=3,alpha=0.5,shape=19,color='black')+
  geom_line(aes(y=pdf_norm,x=x_hat,color='c'),key_glyph='rect',
               data = pred_dat_new,linewidth=0.75,linetype=3)+
  geom_area(aes(y=pdf_norm,x=x_hat),alpha=0.25,
            data = pred_dat_new,fill='#FF7F00')+
  geom_line(aes(y=pdf_ln,x=x_hat,color='b'),key_glyph='rect',
               data = pred_dat_new,linewidth=0.75,linetype=2)+
  geom_area(aes(y=pdf_ln,x=x_hat),alpha=0.25,
            data = pred_dat_new,fill='#984EA3')+
  ylab("Density")+xlab("Emissions")+
  ggtitle("NSPS test run population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept=(emission_mean_new),color='a'),
             linewidth=1,linetype=1,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_new[UPL_table$distribution=="Normal"],color='c'),
             linewidth=1,linetype=3,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_new[UPL_table$distribution=="Lognormal"],color='b'),
             linewidth=1,linetype=2,key_glyph='rect')+
  geom_vline(aes(xintercept=UPL_table$UPL_new[UPL_table$distribution=="Skewed"],color='d'),
             linewidth=1,linetype=4,key_glyph='rect')+
  scale_x_continuous(expand=expansion(mult=c(0,0.05)))+
  scale_y_continuous(expand=expansion(mult=c(0,0.05)))+
  coord_cartesian(clip='off')+labs(color="Distribution:")+
  geom_rug(sides='b',aes(x=emissions),data=dat_top_new,
           alpha=0.5,outside=TRUE,color='black')+
  scale_color_manual(values=c('black','#984EA3','#FF7F00','#377EB8'),
                     labels=c('Observations','Lognormal','Normal','Skewed'))+
  guides(fill='none')
```

\newpage
## Appendix A: Existing Source Data Table

Below is the table containing all of emissions data used in the existing source 
UPL calculation. Note that you can download a .csv version of these data while
using the UPL_app.Rmd webpage. 

```{r dat_EG, echo = FALSE, results=TRUE}
dat_top_exist%>%knitr::kable(caption="Top performering source emissions data used in existing source UPL calculations.",digits=12,
             col.names=names(dat_top_exist))
```

```{r cond4, child='App_report_children/conditional4.Rmd', eval=nsps_EVAL}

```
```{r dat_NSPS, echo=FALSE, eval=nsps_EVAL}
dat_top_new%>%knitr::kable(caption="Best performering source emissions data used in new source UPL calculations.",digits=12,
             col.names=names(dat_top_new))
```

\newpage

## Appendix C: R Code

### MACT_EG()
```{r function_docs1, results=TRUE, message=TRUE, echo=FALSE}
MACT_EG
```

### MACT_NSPS()
```{r function_docs2, results=TRUE, message=TRUE, echo=FALSE}
MACT_NSPS
```

### distribution_type()
```{r function_docs3, results=TRUE, message=TRUE, echo=FALSE}
distribution_type
```

### Normal_UPL()
```{r function_docs4, results=TRUE, message=TRUE, echo=FALSE}
Normal_UPL
```

### Lognormal_UPL()
```{r function_docs5, results=TRUE, message=TRUE, echo=FALSE}
Lognormal_UPL
```

### Skewed_UPL()
```{r function_docs6, results=TRUE, message=TRUE, echo=FALSE}
Skewed_UPL
```

### obs_density()
```{r function_docs7, results=TRUE, message=TRUE, echo=FALSE}
obs_density
```        

\newpage
## Appendix D: System and package versions
```{r versions, results=TRUE, echo=FALSE}
sessionInfo()
```
