---
title: "National Emission Standards for Hazardous Air Pollutants (NESHAP) Maximum Achievable Control Technology (MACT) Floor Analysis for *insert stationary source category* Units for Final Rule"
author: "OAR-OCAP-NRD-EDAB or author names here"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    extra_dependencies: ["float"]
---

<!-- Change the document title and authors as needed -->

<!-- Memorandum example text was taken from past EPA rulemakings -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  error=TRUE,
  warning=FALSE,
  message=FALSE,
  fig.align = "center",
  fig.pos="H",
  out.extra = ""
)
library(htmltools)
knitr::knit_hooks$set(imgcenter = function(before, options, envir){
  if (before) {
    HTML("<p align='center'>")
  } else {
    HTML("</p>")
  }
})
library(dplyr)
library(ggplot2)
library(tidyverse)
library(EnvStats)
library(grid)
library(np)
library(sfsmisc)
library(UPLforOAR)
```

<!-- Look for *italicized content* to edit as appropriate -->

### Introduction

Section 112 of the Clean Air Act (CAA) requires the the U.S. Environmental Protection Agency (EPA) establish National Emission Standards for Hazardous Air Pollutants (NESHAP) for the control of the hazardous air pollutants (HAP) emitted from both new and existing sources in a source category. These standards must reflect the maximum degree of reduction in emissions that is achievable. The minimum level of control is referred to as the “Maximum Achievable Control Technology (MACT) floor.” The method for determining the MACT floor as the minimum control level allowed for establishing standards for a NESHAP is defined for both new and existing sources by CAA section 112(d)(3). For new sources, the MACT floor 
cannot be less stringent than the emission control that is achieved in practice by the best-controlled similar source. For existing sources, the MACT floor cannot be less stringent than the average emission limitation achieved by the best-performing 12 \% of existing sources for source categories with 30 or more sources, or the best-performing 5 sources for source categories with fewer than 30 sources. 

The purpose of this memorandum is to present the methodology and the results of the MACT floor analysis for the NESHAP source category "insert stationary source category". This MACT floor analysis uses data collected in a nationwide survey of *insert stationary source category* unit owners and operators conducted by the EPA in *2025* under the Information Collection Request for NESHAP for *insert stationary source category*. 

### MACT Floor Analysis Hazardous Air Pollutants

The specific chemicals, compounds, or groups of compounds designated to be HAP are listed in CAA section 112(b). From this HAP list, the following HAP, among others, were identified to be emitted from *insert stationary source category* units and were included in the MACT floor analysis: *chemical X, chemical Y, chemical Z...*.

### MACT Floor Analysis subcategories

Under CAA section 112(d)(1), the EPA has the discretion to “...distinguish among classes, types, and sizes of sources within a category or subcategory in establishing...” standards. When separate subcategories are established, a MACT floor is determined separately for each subcategory. To determine whether the EGU source category warrants subcategorization for the MACT floor analysis, the EPA reviewed *insert stationary source category* design, operating information, and air emissions data compiled in the ICR data set and other information collected by the Agency for 
development of the NESHAP for this source category. Based on this review, the EPA concluded that there are significant design and operational differences in *insert stationary source category* that affect the levels of 
emissions of certain HAP to justify subcategorizing *insert stationary source category* for the purpose of establishing MACT floors. 

For the MACT floor analysis, the following *insert stationary source category* subcategories were defined:

- **Subcategory 1.** *describe it here and how/why it was determined*

- **Subcategory 2.** *describe it here and how/why it was determined*

### Calculating Emission Values used in the MACT Floor Analysis

The EPA received emissions data from industry respondents under the *insert stationary source category* 
MACT ICR in a number of different units of measure and converted these units of measure 
to a common *insert measurement units*. 

*If any outlier analysis or subsequent QA/QC was done on the emissions data prior to MACT floor analysis, please include it here.*

### MACT Floor Analysis Methodology

The first step in the MACT floor analysis for each subcategory and HAP (or surrogate) 
was to rank all of the *insert stationary source category* units in the subcategory (for which there were emissions data in the ICR data set) for the pollutant type by emission level (lowest to highest). From this ranking, a MACT floor pool of sources was identified for determining the minimum control level allowed for the MACT floor, consistent with the criteria defined for new and existing sources by CAA section 112(d)(3). For the new-source MACT floors, the best-controlled similar source was identified for which there were individual source test run data in the ICR data set. For the existing-source MACT floors, selection of the MACT floor pool size (i.e., number of units to be included in the determination of the average emission limitation value) was determined on an individual subcategory basis as described below. 

- **Subcategory 1.** *do any subcategories need different handling?*

- **Subcategory 2.** *do any subcategories need different handling?*

The next step in the MACT floor analysis was to incorporate data variability into the calculations of the applicable MACT floor limits for the subcategories using a 99 \% upper prediction limit (UPL) approach. Specifically, the MACT floor limit is an UPL calculated with the Student’s t-test using the `qt()` function in R `stats` software. The Student’s t-test has also been used in other EPA rulemakings (e.g., NESHAP for Portland Cement, NSPS for Hospital/Medical/Infectious Waste Incinerators [HMIWI], NESHAP for Industrial, Commercial, and Institutional Boilers and Process Heaters) in accounting for variability and reflects the level of confidence. The level of confidence represents the level of protection afforded to facilities whose emissions are in line with the best performers, and consequently, the level of confidence is not arbitrary. For example, a 99 \% level of confidence means that a facility whose emissions are in line with the best performers has one chance in 100 of exceeding the floor limit. A prediction interval for a single future observation (or an average of several test observations) is an interval that will, with a specified degree of confidence, contain the next (or the average of some other pre-specified number) of randomly selected observation(s) from a population. In other words, the UPL estimates what the upper bound of future values will be, based upon present or past background samples taken. The UPL consequently represents the value at which we can expect the mean of future observations for the HAP or HAP surrogate emissions to fall within a specified level of confidence, based upon the measurements from an independent sample from the same population. The UPL approach encompasses all the data point-to-data point variability. 

The form of the UPL equation differs depending upon the nature of the data set to which it is applied. To this end, the data sets were evaluated for each HAP and HAP surrogate to ascertain whether the data were distributed normally, log-normally, or skew-normally. Statistical tests of the kurtosis, skewness, and goodness of fit test are then used to evaluate the normality assumption. The skewness statistic, $S$, characterizes the degree of asymmetry of a given data distribution. Normally distributed data have $S = 0$. An $S$ that is greater (less) than 0 indicates that the data are asymmetrically distributed with a right (left) tail extending towards positive (negative) values. Further, the standard error of the skewness statistic, $SES$, can be approximated by $SES = \sqrt{6/n}$, where $n$ is the sample size. According to the small sample skewness hypothesis test, if $S$ is greater than two times $SES$, the data distribution can be considered non-normal. The kurtosis statistic, $K$, characterizes the degree of peakedness or flatness of a given data distribution in comparison to a normal distribution. Normally distributed data have $K = 3$. A $K$ that is greater (less) than 3 indicates a relatively peaked (flat) distribution. Further, the standard error of the kurtosis statistic, $SEK$, can be approximated by $SEK = \sqrt{24/n}$, where $n$ is the sample size. According to the small sample kurtosis hypothesis test, if $K$ is greater than two times $SEK$ value, the data distribution is typically considered to be non-normal. 

For each data set to which the UPL was applied (i.e., the separate data sets for each HAP and HAP surrogate type applicable to the subcategory), the $S$ and $K$ statistics were calculated using both the reported test values and the log-transformed reported test values. If the $S$ and $K$ statistics of the reported data set were both less than twice $SES$ and $SEK$, respectively, the dataset was classified as normally distributed. If neither of $S$ and $K$, or only one of these statistical values were less than twice $SES$ or $SEK$, respectively, then the skewness and kurtosis hypothesis tests were conducted for the natural log-transformed data. If $S$ and $K$ of the log-transformed reported data set were both less than twice $SES$ and $SEK$, respectively, the log-transformed dataset was classified as normally distributed. In this case, the MACT floor was calculated using an UPL equation developed for log-normal data (Bhaumik and Gibbons, 2004). 

In the case of normally distributed data, the MACT floor limit is an UPL calculated as:

$$
UPL = \overline{x} + t_{df,~99}\sqrt{\sigma^2\left(\frac{1}{n}+\frac{1}{m}\right)}
$$
Where $\overline{x}$ is the mean of the top performing sources calculated as $\overline{x}=\frac{1}{n}\displaystyle\sum_{i=1}^{n}x_{i}$, $n$ is the number of test runs, $m$ is the number of future test runs, $\sigma^{2}$ is the variance calculate as $\sigma^{2} = \frac{1}{n-1}\displaystyle\sum_{i=1}^{n}(x_{i}-\overline{x})^2$, $df$ is the degrees of freedom calculated as $df = n-1$, and $t_{df,~99}$ is the 99<sup>th</sup> quantile of the *t*-distribution with $df$ degees of freedom. This UPL calculation was performed in R using the `Normal_UPL()` function from the `UPLforOAR` package version `r packageVersion('UPLforOAR')`. See Appendix C for the function details.

In the case of log-normally distributed data, the MACT floor limit is an UPL calculated as:

$$
UPL= e^{\hat{\mu}+\frac{\hat{\sigma}^2}{2}}+\frac{z_{~0.99}}{m}\sqrt{m e^{2\hat{\mu}+\hat{\sigma}^{2}}(e^{\hat{\sigma}^{2}}-1)+m^{2}e^{2\hat{\mu}+\hat{\sigma}^{2}}\left(\frac{\hat{\sigma}^{2}}{n}+\frac{\hat{\sigma}^{4}}{2(n-1)}\right)}
$$
Where $\hat{\mu}=\frac{1}{n}\displaystyle\sum_{i=1}^{n}log(x_{i})$ is the mean of the log-transformed data from the top performing sources, $\hat{\sigma}^{2}=\frac{1}{n-1}\displaystyle\sum_{i=1}^{n}(log(x_{i})-\hat{\mu})^{2}$ is the variance of the log-transformed data, and $z_{~0.99}$ is the 99<sup>th</sup> percentile of the log-normal distribution estimated using the trapezoidal rule from the following equation:

$$
0.99 = \int_{0}^{z_{~0.99}}\left(1-\frac{\sqrt{\beta_{1z}}}{6}(3z-z^3)+\frac{(\beta_{2z-3})(3-6z^2+z^4)}{24} \right)\phi(z)
$$
Which is equation *XX* in Bhaumik and Gibbons 2004. This UPL calculation was performed in R using the `Lognormal_UPL()` function from the `UPLforOAR` package version `r packageVersion('UPLforOAR')`. See Appendix C for the function details.


### Adjustment for Below Detection Level Emissions Data 

To account for the effect of measurement imprecision associated with the data in the ICR databases (that include method detection level data), a protocol specified by the EPA was used. The procedure for determining a representative detection level (RDL) begins with identifying all of the available reported pollutant-specific method detection levels for the best performing units regardless of any subcategory (e.g., existing or new, fuel type, etc.). From that combined pool of data, we calculate the arithmetic mean value. By limiting the data set to those tests used to establish the floor or emissions limit (i.e., best performers), the EPA believes that the result is representative also of the best performing testing companies and laboratories. The EPA also believes that the outcome should minimize the effect of a test(s) with an inordinately high method detection level (e.g., the sample volume was too small, the laboratory technique was insufficiently sensitive, or the procedure for determining the detection level was other than that specified). The EPA then calls the resulting mean of the method detection levels the RDL, and the Agency considers it characteristic of accepted source emissions measurement performance. 

The second step in the process is to calculate three times the RDL to compare with the calculated floor or emissions limit. We use the multiplication factor of 3 to approximate a 99 percent upper confidence interval for a data set of 7 or more values. For comparing to the floor, if three times the RDL were less than the calculated floor or emissions limit (e.g., calculated from the UPL), we would conclude that measurement variability was adequately addressed. The calculated floor or emissions limit would need no adjustment. If, on the other hand, the value equal to three times the RDL were greater than the UPL, we would conclude that the calculated floor or emissions limit does not account entirely for measurement variability. If indicated, we substituted the value equal to three times the RDL for the calculated MACT floor UPL emissions level.

*Add any RDL content specific to your source category*

```{r table1, echo=FALSE, out.width="100%"}
col1=c('Subcategory 1','Subcategory 2','Subcategory 3','Subcategory 4','Subcategory 5')
col2=c('HAP 1, HAP 2', 'HAP 1', 'HAP 1, HAP 2, HAP 3', 'None', 'None')
col3=c('HAP 1, HAP 2', 'HAP 1', 'HAP 1, HAP 2, HAP 3', 'None', 'None')
col4=c('HAP 1, HAP 2, X, Y, Z', 'HAP 1, X, Y, Z', 'HAP 1, HAP 2, HAP 3, X, Y, Z', 'None', 'X, Y, Z')
table1=tibble(col1=col1,col2=col2,col3=col3,col4=col4)
table1%>%knitr::kable(caption="Here is an example of a table in case you need one",
             col.names=c('',
                         'Existing MACT floor (old units)',
                         'Existing MACT floor (new units)',
                         'New MACT floor'))
```

### Other Data Excluded from MACT FLoor Data Set

*there is always something... put it here*

### MACT Floor Analysis Results

### References

Bhaumik, D.K. and Gibbons, R.D. 2004. An Upper Prediction Limit for the Arithmetic Mean of a Lognormal Random Variable. Technometrics, 46(2):239-248. 



```{r example, echo=TRUE}
library(UPLforOAR)
dat_emiss = read_csv("man/data_example/MATS_Hg.csv", col_names = TRUE)
dat_emiss$sources = paste0(dat_emiss$`Plant Name`, "_", dat_emiss$`Unit Number`,
                         "_", dat_emiss$boiler_id)
dat_emiss$emissions = dat_emiss$Mercury_min_lb_MMBtu
dat_emiss = subset(dat_emiss, select = c(sources, emissions))
nrow(dat_emiss) # number of tests in data set
summary(dat_emiss$emissions)

dat_EG = MACT_EG(CAA_section=112, dat_emiss)
dat_EG_avg = dat_EG%>%group_by(sources)%>%
  summarize(avg = mean(emissions), counts = n())
dat_EG_avg = arrange(dat_EG_avg, avg)
distribution_result_EG = distribution_type(dat_EG)

```
```{r EGtable, fig.align = "center"}
knitr::kable(dat_EG_avg%>%mutate(across(c(2), ~signif(., digits=4)))%>%head(),
             caption="Top 5 of 42 sources for EG standard UPL calculation",
             col.names=c('Source', 'Average emission', 'No. of Tests'), digits=12)
```

Since there were more than 30 sources in the emissions data, the top 12% were chosen to represent the top sources. This yielded `r nrow(dat_EG)` sources. The data included in this regulatory docket were test averages as opposed to individual runs. As such the number of future runs used in UPL calculations will be 1 instead of the default, an average of 3 runs. The appropriate distribution for the UPL calculation is `r distribution_result_EG`, according to ratios of skewness and kurtosis. This simple distribution test does not take into account that we cannot have negative emissions, so we will also calculate the Lognormal UPL for comparison, which is strictly positive.
```{r UPL, echo=TRUE}
UPL1_EG = Normal_UPL(data = dat_EG,
                     future_runs = 1,
                     significance = 0.99)
UPL2_EG = Lognormal_UPL(data = dat_EG,
                        future_runs = 1,
                        significance = 0.99)
```
Next we calculate the UPL using the appropriate distribution, which results in a MACT floor standard of `r UPL1_EG` lb/MMBtu in Hg emissions. Lastly, we will want to plot observation density of emissions data as well as the `r distribution_result_EG` distribution that was used to as the probability density function for the UPL calculation. From the figure below, the Lognormal appears to better represent the data, which would yield a UPL MACT floor standard of `r UPL2_EG` lb/MMBtu in Hg emissions.
```{r distr, echo=TRUE}
# make an ordered sequence of emissions 
# for which we will define the probability density
x_hat = seq(0, 3*max(dat_EG$emissions), length.out = 1024)
# next define the probability density along x_hat
# and at each emission observation.
obs_dens_results = obs_density(dat_EG, xvals = x_hat)
Obs_onPoint = obs_dens_results$Obs_onPoint
obs_den_df = obs_dens_results$obs_den_df
# create a probability density function along the same x_hat
# based on estimated distribution parameters
pdf_n = dnorm(x_hat, mean = mean(dat_EG$emissions, na.rm = T),
              sd=sd(dat_EG$emissions, na.rm = T))
pdf_ln = dlnorm(x_hat,mean = log(mean(dat_EG$emissions, na.rm = T)),
              sd = sd(log(dat_EG$emissions), na.rm = T))
pred_dat = tibble(x_hat, pdf_ln, pdf_n)
```
```{r plot1, echo=TRUE, fig.dim = c(5,4), fig.align = "center", fig.cap = "Observation density of Hg for the overall population. The obseration data are indicated in black as points and a rug along the axis, with the observation density distribution as a black line. The fitted lognormal distribution that is the basis of the UPL estimate is colored purple. The fitted normal distribution that is the basis of the UPL estimate is colored orange. The average of the Hg emissions is the vertical black line and the UPL results are the vertical purple and orange lines."}
ggplot()+
  geom_line(data = obs_den_df, size=0.75,
            aes(y = ydens, x = x_hat, color = 'a'))+
  geom_area(data = obs_den_df, alpha=0.25,
            aes(y = ydens, x = x_hat, fill = 'a'))+
  geom_point(aes(y = ydens, x = emissions), data = Obs_onPoint,
             size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_line(aes(y = pdf_n, x = x_hat, color = 'b'),
               data = pred_dat, size = 0.75, linetype = 2)+
  geom_area(aes(y = pdf_n, x = x_hat, fill = 'b'), alpha = 0.25,
            data = pred_dat)+
  geom_line(aes(y = pdf_ln, x = x_hat, color = 'c'),
               data = pred_dat, size = 0.75, linetype = 3)+
  geom_area(aes(y = pdf_ln, x = x_hat, fill = 'c'), alpha = 0.25,
            data = pred_dat)+
  ylab("Density")+xlab("Hg emissions (lb/MMBtu)")+
  ggtitle("Overall observed population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept = (mean(dat_EG$emissions)),
                 color = 'a'), size = 1, linetype = 1)+
  geom_vline(aes(xintercept = UPL1_EG, color = 'b'), 
             linewidth = 1, linetype = 2)+
  geom_vline(aes(xintercept = UPL2_EG, color = 'c'), 
             linewidth = 1, linetype = 3)+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  coord_cartesian(clip = 'off')+
  labs(color = 'Distribution:', fill = 'Distribution:')+
  geom_rug(sides = 'b', aes(x = emissions), data = dat_EG,
           alpha = 0.5, outside = TRUE, color = 'black')+
  scale_color_manual(values = c('black', '#FF7F00', '#984EA3'),
                     labels = c('Observations', 'Normal', 'Lognormal'))+
  scale_fill_manual(values = c('black', '#FF7F00', '#984EA3'),
                    labels = c('Observations', 'Normal', 'Lognormal'))

```

\newpage
## Appendix A: Existing Source data table

Below is the table containing all of emissions data used in the existing source 
UPL calculation.

```{r dat_EG, echo = FALSE}
dat_top_exist%>%knitr::kable(caption="Top performering source emissions data used in Existing source UPL calculations.",digits=12,
             col.names=names(dat_top_exist))
```

\newpage
## Appendix B: NSPS data table

Below is the table containing all of emissions data used in the NSPS 
UPL calculation.

```{r dat_NSPS, echo=FALSE}
dat_top_new%>%knitr::kable(caption="Best performering source emissions data used in NSPS UPL calculations.",digits=12,
             col.names=names(dat_top_new))
```

\newpage
## Appendix C: R code

### Disclaimer

The United States Environmental Protection Agency (EPA) GitHub project code is provided on an "as is" basis and the user assumes responsibility for its use.  EPA has relinquished control of the information and no longer has responsibility to protect the integrity , confidentiality, or availability of the information.  Any reference to specific commercial products, processes, or services by service mark, trademark, manufacturer, or otherwise, does not constitute or imply their endorsement, recommendation or favoring by EPA.  The EPA seal and logo shall not be used in any manner to imply endorsement of any commercial product or activity by EPA or the United States Government.

### MACT_EG()
```{r function_docs1, results=TRUE, message=TRUE, echo=FALSE}
MACT_EG
```

### MACT_NSPS()
```{r function_docs2, results=TRUE, message=TRUE, echo=FALSE}
MACT_NSPS
```

### distribution_type()
```{r function_docs3, results=TRUE, message=TRUE, echo=FALSE}
distribution_type
```

### Normal_UPL()
```{r function_docs4, results=TRUE, message=TRUE, echo=FALSE}
Normal_UPL
```

### Lognormal_UPL()
```{r function_docs5, results=TRUE, message=TRUE, echo=FALSE}
Lognormal_UPL
```

### Skewed_UPL()
```{r function_docs6, results=TRUE, message=TRUE, echo=FALSE}
Skewed_UPL
```

### obs_density()
```{r function_docs7, results=TRUE, message=TRUE, echo=FALSE}
obs_density
```        

\newpage
## Appendix D: System and package versions
```{r versions, results=TRUE, echo=FALSE}
sessionInfo()
```



