---
title: "Convergence Results"
author: "Ludda Ludwig"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,error=TRUE,warning=FALSE, message=FALSE)
```
Convergence results from Bayesian UPL analysis using EPA.MACT.floor.UPL version
`r packageVersion('EPA.MACT.floor.UPL')`. Convergence is assessed quantitatively
with Gelman-Rubin diagnostics, and visually with convergence plots. Gelman-Rubin 
diagnostic values of less than 1.1 indicate good convergence. Values greater 
than 1.1 but less than 1.2 indicate weak convergence. Values greater than 1.2 
indicate there is no convergence, and the results should be rejected and the model
or distribution reassessed for appropriateness to the data.
```{r gelamn}
knitr::kable(conv_output,caption="Gelman-Rubin diagnostic results",
             col.names=c('Parameter','Diagnostic','Converged?','Distribution'),
             digits = 4)
```

For each distribution used, a pair of plots is generated for each defining 
parameter. The first plot shows the MCMC iterations for each of 3 chains after 
removing the burn-in iterations. Good convergence is indicated if all chains are 
well mixed and without trends. The second plot shows the histogram of the 
parameter's posterior distribution. Ideally the histogram will look 
approximately normal centered on the converged parameter value. If the histogram 
is extremely skewed towards the edge of the parameter space, than the true value 
might be outside the limits imposed on the prior distribution. Consider 
broadening the prior information. If the histogram is flat across the parameter 
space, than it is likely there is too little information in data or too poor a 
model fit to converge, even if convergence checks have otherwise passed. In this 
case, consider narrowing prior information or removing this distribution from 
consideration.
```{r plots,results=FALSE}

figs_list

```
