---
author: "OAR-OCAP-NRD-EDAB and/or author names here"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    extra_dependencies: ["float"]
params:
  source:
    label: "Stationary source category"
    input: text
    value: "Example Industry"
  data:
    label: "Select data set, with named columns of emissions and sources"
    input: file
    accept: ".csv"
    value: "Example_data3.csv"
  CAA_section: 
    label: "CAA Section"
    value: 112
    input: select
    choices: [112, 129]
  national_N:
    label: "Number of sources nation-wide (for CAA 129 only)"
    input: numeric
    value: NA
  future_runs: 
    label: "Number of future test runs"
    value: 3
    input: numeric
  significance:
    label: "Level of significance" 
    value: 0.99
    input: numeric
---
<!-- Change the document title and authors as needed -->

---
title: "National Emission Standards for Hazardous Air Pollutants (NESHAP) Maximum Achievable Control Technology (MACT) Floor Analysis for `r params$source` Units for Final Rule"
---

<!-- Memorandum example text was taken from past EPA rulemakings -->
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  error=TRUE,
  warning=FALSE,
  message=FALSE,
  appendix = FALSE,
  fig.align = "center",
  fig.pos="H",
  out.extra = ""
)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(EnvStats)
library(grid)
library(np)
library(sfsmisc)
library(UPLforOAR)
if (params$CAA_section == 112){
  CAA_text = 'CAA Section 112(d)(3)'
} else if (params$CAA_section == 129){
  CAA_text = 'CAA Section 129(a)(2)'
}
```
<!-- Repeat the analysis block below with unique names if you have multiple HAPS or subcategories -->
```{r analysis_1, results = FALSE, appendix = TRUE}
# Below is an example of loading data, sorting and selecting to top performers, 
# and calculating the correct distribution and UPL for existing source and NSPS. 
# If there are multiple HAP or subcategories, simply repeat with code block for 
# those data, making sure to use unique names so you don't overwrite the results.

dat_emiss = read_csv(params$data)
if (any(dat_emiss$emissions == 0)){
  zero_warn = TRUE
} else {
  zero_warn = FALSE
}
dat_top_exist = MACT_existing(CAA_section = params$CAA_section, 
                              data = dat_emiss, national_N = params$national_N,
                              emissions = 'emissions', sources = 'sources')
dat_top_new = MACT_new(data = dat_emiss, emissions = 'emissions', 
                       sources = 'sources')
dat_top_exist$ln_emiss = log(dat_top_exist$emissions)
dat_top_exist$ln_emiss = replace(dat_top_exist$ln_emiss, 
                               !is.finite(dat_top_exist$ln_emiss), NA)
dat_top_new$ln_emiss = log(dat_top_new$emissions)
dat_top_new$ln_emiss = replace(dat_top_new$ln_emiss, 
                               !is.finite(dat_top_new$ln_emiss), NA)
dat_topmeans_exist = dat_top_exist %>% group_by(sources) %>% 
                              summarize(means = mean(emissions), counts = n())
if (all(dat_topmeans_exist$counts == 1)){
  source_fig = FALSE
} else {
  source_fig = TRUE
}
best_source = as.character(dat_top_new$sources)[1]
maxY = NULL # this defaults to 3*max(emissions)
xvals = NULL # this defaults to a sequence from 0 to 3*max(emissions)

# for this report, you only want one distribution per UPL that was the best choice
distribution_ex = c('Gamma')
results = Bayesian_UPL(data = dat_top_exist, distr_list = distribution_ex,
                       emissions = 'emissions',
                       future_runs = params$future_runs, 
                       significance = params$significance,
                       maxY = maxY, xvals = xvals)
maxY_new = NULL # this defaults to 3*max(emissions)
xvals_new = NULL # this defaults to a sequence from 0 to 3*max(emissions)
# for this report, you only want one distribution per UPL that was the best choice
distribution_new = c('Gamma') 
results2 = Bayesian_UPL(data = dat_top_new, distr_list = distribution_new, 
                        emissions = 'emissions',
                        future_runs = params$future_runs, 
                        significance = params$significance,
                        maxY = maxY_new, xvals = xvals_new)
```

<!-- Edit/add/remove any content as appropriate -->
<!-- Italicized comments in * * likely need to be edited for each use -->

## Introduction

Section 112 of the Clean Air Act (CAA) requires the the U.S. Environmental Protection Agency (EPA) establish National Emission Standards for Hazardous Air Pollutants (NESHAP) for the control of the hazardous air pollutants (HAP) emitted from both new and existing sources in a source category. These standards must reflect the maximum degree of reduction in emissions that is achievable. The minimum level of control is referred to as the “Maximum Achievable Control Technology (MACT) floor.” The method for determining the MACT floor as the minimum control level allowed for establishing standards for a NESHAP is defined for both new and existing sources by `r CAA_text`. For new sources, the MACT floor cannot be less stringent than the emission control that is achieved in practice by the best-controlled similar source. For existing sources, the MACT floor cannot be less stringent than the average emission limitation achieved by the best-performing 12 \% of existing sources for source categories with 30 or more sources or those governed by CAA Section 129(a)(2), or the best-performing 5 sources for source categories with fewer than 30 sources governed by CAA Section 112(d)(3). 

The purpose of this memorandum is to present the methodology and the results of the MACT floor analysis for the NESHAP source category `r params$source`. This MACT floor analysis uses data collected in a nationwide survey of `r params$source` unit owners and operators conducted by the EPA in *2025* under the Information Collection Request for NESHAP for `r params$source`. 

## MACT Floor Analysis Hazardous Air Pollutants

The specific chemicals, compounds, or groups of compounds designated to be HAP are listed in *CAA section 112(b)*. From this HAP list, the following HAP, among others, were identified to be emitted from `r params$source` units and were included in the MACT floor analysis: *chemical X, chemical Y, chemical Z...*.

## MACT Floor Analysis subcategories

Under *CAA section 112(d)(1)*, the EPA has the discretion to “...distinguish among classes, types, and sizes of sources within a category or subcategory in establishing...” standards. When separate subcategories are established, a MACT floor is determined separately for each subcategory. To determine whether the EGU source category warrants subcategorization for the MACT floor analysis, the EPA reviewed `r params$source` design, operating information, and air emissions data compiled in the ICR data set and other information collected by the Agency for development of the NESHAP for this source category. Based on this review, the EPA concluded that there are significant design and operational differences in `r params$source` that affect the levels of emissions of certain HAP to justify subcategorizing `r params$source` for the purpose of establishing MACT floors. 

For the MACT floor analysis, the following `r params$source` subcategories were defined:

- **Subcategory 1.** *describe it here and how/why it was determined*

- **Subcategory 2.** *describe it here and how/why it was determined*

## Calculating Emission Values used in the MACT Floor Analysis

The EPA received emissions data from industry respondents under the `r params$source` 
MACT ICR in a number of different units of measure and converted these units of measure 
to a common *insert measurement units*. 

*If any outlier analysis or subsequent QA/QC was done on the emissions data prior to MACT floor analysis, please include it here.*

## MACT Floor Analysis Methodology

The first step in the MACT floor analysis for each subcategory and HAP (or surrogate) was to rank all of the `r params$source` units in the subcategory (for which there were emissions data in the ICR data set) for the pollutant type by emission level (lowest to highest). From this ranking, a MACT floor pool of sources was identified for determining the minimum control level allowed for the MACT floor, consistent with the criteria defined for new and existing sources by `r CAA_text`. For the new-source MACT floors, the best-controlled similar source was identified for which there were individual source test run data in the ICR data set. The selection of top performing sources for the existing source MACT floor was performed using the  `MACT_existing()` function in the `UPLforOAR` `R` package version `r packageVersion('UPLforOAR')`, with details in Appendix C. The selection of the best performing source for the new source MACT floor was performed using the `MACT_new()` function in the `UPLforOAR` `R` package version `r packageVersion('UPLforOAR')`, with details in appendix C. 

For the existing-source MACT floors, selection of the MACT floor pool size (i.e., number of units to be included in the determination of the average emission limitation value) was determined on an individual subcategory basis as described below. 

- **Subcategory 1.** *do any subcategories need different handling?*

- **Subcategory 2.** *do any subcategories need different handling?*

### Probability distribution Selection

The next step in the MACT floor analysis was to incorporate data variability into the calculations of the applicable MACT floor limits for the subcategories using a `r (params$significance)*100`\% upper prediction limit (UPL) approach. Specifically, the UPL is calculated as the `r (params$significance)*100`$^{th}$ quantile of the average of `r params$future_runs` future emission test values. The level of confidence represents the level of protection afforded to facilities whose emissions are in line with the best performers, and consequently, the level of confidence is not arbitrary. For example, a 99\% level of confidence means that a facility whose emissions are in line with the best performers has one chance in 100 of exceeding the floor limit. A prediction interval for a single future observation (or an average of several test observations) is an interval that will, with a specified degree of confidence, contain the next (or the average of some other pre-specified number) of randomly selected observation(s) from a population. In other words, the UPL estimates what the upper bound of future values will be, based upon present or past background samples taken. The UPL consequently represents the value at which we can expect the mean of future observations for the HAP or HAP surrogate emissions to fall within a specified level of confidence, based upon the measurements from an independent sample from the same population. The UPL approach encompasses all the data point-to-data point variability. 

### Implementation

The form of the UPL equation differs depending upon the nature of the data set to which it is applied. To this end, the data sets were evaluated for each HAP and HAP surrogate to ascertain whether the data were normally, log-normally, skew-normally, beta, or gamma distributed. For each distribution, a probability density function was fit to the test run emissions using Bayesian Markov Chain Monte Carlo (MCMC). This analysis was implemented in `R` using `runjags` and JAGS (Just Another Gibbs Sampler) software, with three MCMC chains with different initial values, completely uninformed priors, a burn-in and adapt of 10,000 iterations each, and a final set of 10,000 iterations retained in posterior distributions. Random seeds were specified to control for stochasticity and ensure reproducibility. The JAGS model scripts were defined using the function `write_likelihood()`, and executed using functions `setup_likelihood()` and `run_likelihood()` in the `UPLforOAR` `R` package version `r packageVersion('UPLforOAR')`. See Appendix C for more details. 

Convergence was assessed quantitatively by checking Gelman-Rubin diagnostics and qualitatively by viewing plots of the iteration chains to ensure they were well-mixed and without trend, as well as histograms of the posterior distributions to ensure they converged. These checks were carried out using the functions `converge_likelihood()` and `converge_figs()` in the `UPLforOAR` `R` package version `r packageVersion('UPLforOAR')`. See Appendix C for more details. 

Both quantitative and qualitative evaluations were used to select the appropriate distribution. For each distribution, the Sum of Squared Errors ($SSE$) is calculated as:

$$SSE=\displaystyle\sum(y_i-\hat{y_i})^2$$

Where $y_i$ is the empirical test run probability density at emission value $x_{i}$ and $\hat{y_i}$ is the predicted probability density at emission value $x_{i}$. The lower the $SSE$, the better the probability distribution represents the probability of the emission values. The empirical test run probability densities, $y_{i}$, were also checked if they were significantly different than the predicted probability densities, $\hat{y_i}$, by comparing them to the 95\% confidence interval (CI) around the predicted densities derived from posterior distributions. The uncertainty around the predicted probability density can be variable; thus one can have a very different probability at the same $x_{i}$ but be within the 95\% CI if the uncertainty is very large, or have a significantly different probability at the same $x_{i}$ even though the $y_{i}$ and $\hat{y_i}$ are extremely similar if there is very little uncertainty. The more $y_{i}$ that are statistically equivalent to $\hat{y_i}$, the better the distribution represents the probability of $x_{i}$ emission values. A proper probability density function will integrate to $1$, so if the integral of the fitted distribution is substantially different from $1$ this can indicate an issue. Lastly, a visual assessment comparing the empirical and predicted probability densities across the range of possible emission values can identify any problematic behavior or differentiate between distributions. These fit metrics were calculated using the functions `fit_likelihood()` and `obs_density()` in the `UPLforOAR` `R` package version `r packageVersion('UPLforOAR')`. See Appendix C for more details. 

<!-- Note that if a different upper bound was used, the text below needs to be updated to `r maxY` rather than $3\times max(x_i)$. -->

To calculate the UPL, which is the `r (params$significance)*100`$^{th}$ quantile of the average of `r params$future_runs` future emission test values, the posterior distributions of `r params$future_runs` new draws from the fitted probability distribution were averaged and then the UPL is determined as the `r (params$significance)*100`$^{th}$ quantile of this resulting distribution. The fitted distribution is truncated by lower and upper bounds of $0$ and $3\times max(x_i)$ emissions, which prevents extreme values from long-tailed distributions biasing the UPL and otherwise has no effect on the results. The $2.5^{th}$ and $97.5^{th}$ quantiles of the posterior distributions of the predicted probability density $\hat{y_i}$ at each $x_i$ test run value were used to calculate 95\% CI for the goodness-of-fit metrics. The median of the posterior distributions of the predicted probability density $\hat{y_i}$ were used in comparisons to $y_i$ to calculate the $SSE$. The median of the posterior distributions for all $\hat{y_i}$ across the possible range of emission test run values was monitored for use in plotting the distributions and visual comparisons to the empirical distributions. All of these computations were done using the function `output_likelihood()` in the `UPLforOAR` `R` package version `r packageVersion('UPLforOAR')`. See Appendix C for more details. 

The parameter estimation and UPL calculations for each distribution investigated were performed using the function `Bayesian_UPL()` in the `UPLforOAR` `R` package version `r packageVersion('UPLforOAR')`, which wraps the functions described above and concatenates the results for easy comparison. See Appendix C for more details. 

### UPL Calculation for Normal Data

In the case of normally distributed data, the MACT floor limit is an UPL calculated as the `r (params$significance)*100`$^{th}$ quantile of the average of `r params$future_runs` new emission runs, where the probability density, $y_i$, of normally distributed emissions, $x_i$, is dependent on parameters $\mu$, the mean, and $\sigma$ the standard deviation:

$$
y_i = \frac{1}{\sigma \sqrt{2\pi}}e^{-\frac{(x_i-\mu)^{2}}{2\sigma^{2}}}
$$

### UPL Calculation for Log-normal Data

In the case of log-normally distributed data, the MACT floor limit is an UPL calculated as the `r (params$significance)*100`$^{th}$ quantile of the average of `r params$future_runs` new emission runs, where the probability density, $y_i$, of log-normally distributed emissions, $x_i$, is dependent on parameters $\mu$, the mean of $log(x_i)$ and $\sigma$, the standard deviation of $log(x_i)$: 

$$
y_i = \frac{1}{x_i\sqrt{2\pi \sigma^{2}}}e^{-\frac{(log(x_i)-\mu)^{2}}{2\sigma^2}}
$$

### UPL Calculation for Skew-normal Data

In the case of skew-normally distributed data, the MACT floor limit is an UPL calculated as the `r (params$significance)*100`$^{th}$ quantile of the average of `r params$future_runs` new emission runs, where the skew-normal distribution is defined by three parameters; location $\xi$, scale $\omega$, and skewness $\alpha$. This is equivalent to a normal distribution when $\alpha = 0$, it is right-skewed if $\alpha > 0$, and left-skewed if $\alpha < 0$. The probability density, $y_i$, of skewed distributed emissions is:

$$
y_i = \frac{2}{\omega}\phi \left(\frac{x_i-\xi}{\omega}\right) \Phi \left(\alpha\left(\frac{x_i-\xi}{\omega}\right)\right)
$$

Where $\phi(x)$ is the standard normal probability distribution, $\Phi(x)$ is its cumulative probability distribution function, and $x_i$ is transformed to $\frac{x_i-\xi}{\omega}$. Using the error function, this can be expanded to:

$$
y_i = \frac{2}{\omega}\frac{1}{\sqrt{2\pi}}e^{-\frac{\left(\frac{x_i-\xi}{\omega}\right)^{2}}{2}}\frac{1}{2}\left[1+\frac{2}{\sqrt{\pi}}\int_0^{\alpha\frac{x_i-\xi}{\omega\sqrt{2}}}e^{-t^2}dt\right]
$$

### UPL Calculation for Gamma Data

In the case of gamma distributed data, the MACT floor limit is an UPL calculated as the `r (params$significance)*100`$^{th}$ quantile of the average of `r params$future_runs` new emission runs, where the probability density, $y_i$, of gamma distributed emissions, $x_i$, is dependent on parameters $\alpha$, the shape, and $\beta$, the rate:

$$
y_i = \frac{\beta^{\alpha}}{\Gamma(\alpha)}x_i^{\alpha-1}e^{-\beta x_i}
$$

\*Note that $\Gamma$ in this case is the gamma *function* not the Gamma probability distribution.

### UPL Calculation for Beta Data

In the case of beta distributed data, the MACT floor limit is an UPL calculated as the `r (params$significance)*100`$^{th}$ quantile of the average of `r params$future_runs` new emission runs, where the probability density, $y_i$, of beta distributed emissions, $x_i$, is dependent on parameters $\alpha$ and $\beta$:

$$
Y_i = \frac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}x_i^{\alpha -1}(1-x_i)^{\beta-1}
$$

\*Note that $\Gamma$ in this case is the gamma *function* not the Gamma probability distribution.

### Settings used in Bayesian MCMC

The selected distribution for the *chemical X* *subcategory A* existing source MACT floor analysis was the `r distribution_ex` distribution. The distribution's parameter prior distributions, MCMC initial chain values, and Gelman-Rubin diagnostics for convergence are listed in the table below, followed by the random number generator (RNG) state that was used.
<!-- If you specified priors manually, rather than using the defaults you will need to alter the table below -->

```{r gamma_params, appendix = TRUE}
if (is.null(maxY)){
  maxY = 3 * max(dat_top_exist$emissions)
} 
sdY =  sd(dat_top_exist$emissions)
shape = mean(dat_top_exist$emissions)^2 / sdY^2
rate = mean(dat_top_exist$emissions) / sdY^2
parameter = c('Rate', 'Shape')
prior = c(paste0('~dunif (0, ', signif(maxY / (sdY^2), 3), ')'), 
          paste0('~dunif (0, ', signif((maxY^2) / (sdY^2), 3), ')'))
chain1 = c(signif(rate, 3), signif(shape, 3))
chain2 = c(signif(1.5 * rate, 3), signif(0.5 * shape, 3))
chain3 = c(signif(0.5 * rate, 3), signif(1.5 * shape, 3))
diagnostics = c(signif(results$conv_output$gelman_diag[1], 3), 
                signif(results$conv_output$gelman_diag[2], 3))
convYN = c(results$conv_output$convYN[1], 
           results$conv_output$convYN[2])
gamm_exist = tibble(parameter, prior, chain1, chain2, chain3, 
                    diagnostics, convYN)
gamm_exist %>% knitr::kable(
  caption="Parameter information for the gamma distribution 
  UPL calculation for existing source chemical X, subcategory A.",
             col.names=c('Parameter',
                         'Prior',
                         'Chain 1 initial value',
                         'Chain 2 initial value',
                         'Chain 3 initial value',
                         'Gelman-Rubin diagnostic',
                         'Converged?'))
```

```{r RNG1, appendix = TRUE}
state_out = results$state
chains = c('Chain 1', 'Chain 2', 'Chain 3')
rng_name = c(state_out[[1]]$.RNG.name, 
               state_out[[2]]$.RNG.name, 
               state_out[[3]]$.RNG.name)
rng_state = c(paste(state_out[[1]]$.RNG.state, collapse=', '), 
              paste(state_out[[2]]$.RNG.state, collapse=', '), 
              paste(state_out[[3]]$.RNG.state, collapse=', '))
RNG1 = tibble(chains, rng_name, rng_state)
RNG1 %>% knitr::kable(
  caption="Random number generator (RNG) names and states used in MCMC chains",
             col.names=c('MCMC Chain',
                         'Name',
                         'State'))
```

The selected distribution for the *chemical X* *subcategory A* new source MACT floor analysis was the `r distribution_new` distribution. The distribution's parameter prior distributions, MCMC initial chain values, and Gelman-Rubin diagnostics for convergence are listed in the table below, followed by the random number generator (RNG) state that was used.

```{r gamma_params2, appendix = TRUE}
if (is.null(maxY_new)){
  maxY_new = 3 * max(dat_top_new$emissions)
} 
sdY =  sd(dat_top_new$emissions)
shape = mean(dat_top_new$emissions)^2 / sdY^2
rate = mean(dat_top_new$emissions) / sdY^2
parameter = c('Rate', 'Shape')
prior = c(paste0('~dunif (0, ', signif(maxY_new / (sdY^2), 3), ')'), 
          paste0('~dunif (0, ', signif((maxY_new^2) / (sdY^2), 3), ')'))
chain1 = c(signif(rate, 3), signif(shape, 3))
chain2 = c(signif(1.5 * rate, 3), signif(0.5 * shape, 3))
chain3 = c(signif(0.5 * rate, 3), signif(1.5 * shape, 3))
diagnostics = c(signif(results2$conv_output$gelman_diag[1], 3), 
                signif(results$conv_output$gelman_diag[2], 3))
convYN = c(results2$conv_output$convYN[1], 
           results2$conv_output$convYN[2])
gamm_new = tibble(parameter, prior, chain1, chain2, chain3, 
                    diagnostics, convYN)
gamm_new %>% knitr::kable(
  caption="Parameter information for the gamma distribution 
  UPL calculation for new source chemical X, subcategory A.",
             col.names=c('Parameter',
                         'Prior',
                         'Chain 1 initial value',
                         'Chain 2 initial value',
                         'Chain 3 initial value',
                         'Gelman-Rubin diagnostic',
                         'Converged?'))
```

```{r RNG1, appendix = TRUE}
state_out = results2$state
chains = c('Chain 1', 'Chain 2', 'Chain 3')
rng_name = c(state_out[[1]]$.RNG.name, 
               state_out[[2]]$.RNG.name, 
               state_out[[3]]$.RNG.name)
rng_state = c(paste(state_out[[1]]$.RNG.state, collapse=', '), 
              paste(state_out[[2]]$.RNG.state, collapse=', '), 
              paste(state_out[[3]]$.RNG.state, collapse=', '))
RNG1 = tibble(chains, rng_name, rng_state)
RNG1 %>% knitr::kable(
  caption="Random number generator (RNG) names and states used in MCMC chains",
             col.names=c('MCMC Chain',
                         'Name',
                         'State'))
```

<!-- The skew-normal distribution parameter prior distributions, MCMC initial chain values, and Gelman-Rubin diagnostics for convergence that were used in the *chemical X* *subcategory A* new source UPL calculation are listed below. -->
<!-- ```{r skew_params, appendix = TRUE} -->
<!-- if (is.null(maxY)){ -->
<!--  maxY = 3 * max(dat_top_new$emissions) -->
<!-- } -->
<!-- sdY =  sd(dat_top_new$emissions) -->
<!-- meanY = mean(dat_top_new$emissions) -->
<!-- skew1 = min(0.99, abs((1 / length(dat_top_new$emissions)) *  -->
<!--                         sum(((dat_top_new$emissions - meanY) / sdY)^3))) -->
<!-- delta = sqrt((pi / 2) * ((abs(skew1)^(2 / 3)) / ((abs(skew1)^(2 / 3)) +  -->
<!--                                                    ((4 - pi) / 2)^(2 / 3)))) -->
<!-- delta = delta * abs(skew1) / skew1 -->
<!-- alpha = delta / sqrt(1 - delta^2) -->
<!-- omega = sdY / sqrt(1 - 2 * delta^2 / pi) -->
<!-- xi = meanY - omega * delta * sqrt(2 / pi) -->
<!-- parameter = c('Omega', 'Xi', 'Alpha') -->
<!-- prior = c(paste0('~dunif (0, ', signif(maxY, 3), ')'), -->
<!--          paste0('~dnorm (0, ', signif(1 / (100 * maxY), 3), ')'), -->
<!--         '~dunif (-100, 100)') -->
<!-- chain1 = c(signif(xi, 3), signif(omega, 3), signif(alpha, 3)) -->
<!-- chain2 = c(signif(1.5 * xi, 3), signif(0.5 * omega, 3), signif(0.5 * alpha, 3)) -->
<!-- chain3 = c(signif(0.5 * xi, 3), signif(1.5 * omega, 3), signif(1.5 * alpha, 3)) -->
<!-- diagnostics = c(signif(results2$conv_output$gelman_diag[1],3), -->
<!--                 signif(results2$conv_output$gelman_diag[2],3), -->
<!--                 signif(results2$conv_output$gelman_diag[3],3)) -->
<!-- convYN = c(results2$conv_output$convYN[1], -->
<!--            results2$conv_output$convYN[2], -->
<!--            results2$conv_output$convYN[3]) -->
<!-- gamm_exist = tibble(parameter, prior, chain1, chain2, chain3, -->
<!--                     diagnostics, convYN) -->
<!-- gamm_exist %>% knitr::kable( -->
<!--   caption="Parameter information for the skew-normal distribution  -->
<!--   UPL calculation for new source chemical X, subcategory A.", -->
<!--              col.names=c('Parameter', -->
<!--                          'Prior', -->
<!--                          'Chain 1 initial value', -->
<!--                          'Chain 2 initial value', -->
<!--                          'Chain 3 initial value', -->
<!--                          'Gelman-Rubin diagnostic', -->
<!--                          'Converged?')) -->
<!-- ``` -->

<!-- The parameter tables for the other distributions not used in the example are commented-out below -->

<!-- The normal distribution parameter prior distributions, MCMC initial chain values, and Gelman-Rubin diagnostics for convergence that were used in the *chemical X* *subcategory A* existing source UPL calculation are listed below. -->
<!-- ```{r normal_params, appendix = TRUE} -->
<!-- if (is.null(maxY)){ -->
<!--   maxY = 3 * max(dat_top_exist$emissions) -->
<!-- } -->
<!-- sdY =  sd(dat_top_exist$emissions) -->
<!-- meanY = mean(dat_top_exist$emissions) -->
<!-- parameter = c('Mean', 'Standard deviation') -->
<!-- prior = c(paste0('~dunif (0, ', signif(maxY, 3), ')'), -->
<!--           paste0('~dunif (', signif(0.001 * sdY, 3), ' ,',  -->
<!--                  signif(1000 * sdY, 3), ')')) -->
<!-- chain1 = c(signif(meanY, 3), signif(sdY, 3)) -->
<!-- chain2 = c(signif(1.5 * meanY, 3), signif(0.5 * sdY, 3)) -->
<!-- chain3 = c(signif(0.5 * meanY, 3), signif(1.5 * sdY, 3)) -->
<!-- diagnostics = c(signif(results$conv_output$gelman_diag[1], 3), -->
<!--                 signif(results$conv_output$gelman_diag[2], 3)) -->
<!-- convYN = c(results$conv_output$convYN[1], -->
<!--            results$conv_output$convYN[2]) -->
<!-- gamm_exist = tibble(parameter, prior, chain1, chain2, chain3, -->
<!--                     diagnostics, convYN) -->
<!-- gamm_exist %>% knitr::kable( -->
<!--   caption="Parameter information for the normal distribution  -->
<!--   UPL calculation for existing source chemical X, subcategory A.", -->
<!--              col.names=c('Parameter', -->
<!--                          'Prior', -->
<!--                          'Chain 1 initial value', -->
<!--                          'Chain 2 initial value', -->
<!--                          'Chain 3 initial value', -->
<!--                          'Gelman-Rubin diagnostic', -->
<!--                          'Converged?')) -->
<!-- ``` -->

<!-- The log-normal distribution parameter prior distributions, MCMC initial chain values, and Gelman-Rubin diagnostics for convergence that were used in the *chemical X* *subcategory A* existing source UPL calculation are listed below. -->
<!-- ```{r lognormal_params, appendix = TRUE} -->
<!-- if (is.null(maxY)){ -->
<!--  maxY = 3 * max(dat_top_exist$emissions) -->
<!-- } -->
<!-- sdY =  sd(dat_top_exist$ln_emiss) -->
<!-- meanY = mean(dat_top_exist$ln_emiss) -->
<!-- parameter = c('Mean', 'Standard deviation') -->
<!-- prior = c(paste0('~dnorm (', meanY, ', ', signif(0.001 / (sdY^2), 3), ')'), -->
<!--          paste0('~dunif (', signif(0.001 * sdY, 3), ' ,',  -->
<!--                 signif(1000 * sdY, 3), ')')) -->
<!-- chain1 = c(signif(meanY, 3), signif(sdY, 3)) -->
<!-- chain2 = c(signif(1.5 * meanY, 3), signif(0.5 * sdY, 3)) -->
<!-- chain3 = c(signif(0.5 * meanY, 3), signif(1.5 * sdY, 3)) -->
<!-- diagnostics = c(signif(results$conv_output$gelman_diag[1], 3), -->
<!--                signif(results$conv_output$gelman_diag[2], 3)) -->
<!-- convYN = c(results$conv_output$convYN[1], -->
<!--           results$conv_output$convYN[2]) -->
<!-- gamm_exist = tibble(parameter, prior, chain1, chain2, chain3, -->
<!--                    diagnostics, convYN) -->
<!-- gamm_exist %>% knitr::kable( -->
<!--   caption="Parameter information for the log-normal distribution  -->
<!--   UPL calculation for existing source chemical X, subcategory A.", -->
<!--             col.names=c('Parameter', -->
<!--                         'Prior', -->
<!--                         'Chain 1 initial value', -->
<!--                         'Chain 2 initial value', -->
<!--                         'Chain 3 initial value', -->
<!--                         'Gelman-Rubin diagnostic', -->
<!--                         'Converged?')) -->
<!-- ``` -->

<!-- The beta distribution parameter prior distributions, MCMC initial chain values, and Gelman-Rubin diagnostics for convergence that were used in the *chemical X* *subcategory A* existing source UPL calculation are listed below. -->
<!-- ```{r beta_params, appendix = TRUE} -->
<!-- if (is.null(maxY)){ -->
<!--   maxY = 3 * max(dat_top_exist$emissions) -->
<!-- } -->
<!-- sdY =  sd(dat_top_exist$emissions) -->
<!-- meanY = mean(dat_top_exist$emissions) -->
<!-- alpha = meanY^2 / sdY^2 - meanY^3 / sdY^2 - meanY -->
<!-- beta = meanY - 1 + meanY / sdY^2 + meanY^3 / sdY^2 - 2 * meanY^2 / sdY^2 -->

<!-- parameter = c('Alpha', 'Beta') -->
<!-- prior = c(paste0('~dunif (-1, ', signif((0.75^2 / sdY^2 - 0.75^3 / sdY^2 - 0.75)  -->
<!--                                         * 1.25, 3), ')'), -->
<!--           paste0('~dunif (-1, ', signif((0.25 - 1 + 0.25 / sdY^2 +  -->
<!--                                            0.25^3 / sdY^2 - 2 * 0.25^2 / sdY^2)  -->
<!--                                         * 1.25, 3), ')')) -->
<!-- chain1 = c(signif(alpha, 3), signif(beta, 3)) -->
<!-- chain2 = c(signif(1.5 * alpha, 3), signif(0.5 * beta, 3)) -->
<!-- chain3 = c(signif(0.5 * alpha, 3), signif(1.5 * beta, 3)) -->
<!-- diagnostics = c(signif(results$conv_output$gelman_diag[1], 3), -->
<!--                 signif(results$conv_output$gelman_diag[2], 3)) -->
<!-- convYN = c(results$conv_output$convYN[1], -->
<!--            results$conv_output$convYN[2]) -->
<!-- gamm_exist = tibble(parameter, prior, chain1, chain2, chain3, -->
<!--                     diagnostics, convYN) -->
<!-- gamm_exist %>% knitr::kable( -->
<!--   caption="Parameter information for the beta distribution  -->
<!--   UPL calculation for existing source chemical X, subcategory A.", -->
<!--              col.names=c('Parameter', -->
<!--                          'Prior', -->
<!--                          'Chain 1 initial value', -->
<!--                          'Chain 2 initial value', -->
<!--                          'Chain 3 initial value', -->
<!--                          'Gelman-Rubin diagnostic', -->
<!--                          'Converged?')) -->
<!-- ``` -->

## Adjustment for Below Detection Level Emissions Data 

To account for the effect of measurement imprecision associated with the data in the ICR databases (that include method detection level data), a protocol specified by the EPA was used. The procedure for determining a representative detection level (RDL) begins with identifying all of the available reported pollutant-specific method detection levels for the best performing units regardless of any subcategory (e.g., existing or new, fuel type, etc.). From that combined pool of data, we calculate the arithmetic mean value. By limiting the data set to those tests used to establish the floor or emissions limit (i.e., best performers), the EPA believes that the result is representative also of the best performing testing companies and laboratories. The EPA also believes that the outcome should minimize the effect of a test(s) with an inordinately high method detection level (e.g., the sample volume was too small, the laboratory technique was insufficiently sensitive, or the procedure for determining the detection level was other than that specified). The EPA then calls the resulting mean of the method detection levels the RDL, and the Agency considers it characteristic of accepted source emissions measurement performance. 

The second step in the process is to calculate three times the RDL to compare with the calculated floor or emissions limit. We use the multiplication factor of 3. For comparing to the floor, if three times the RDL were less than the calculated floor or emissions limit (e.g., calculated from the UPL), we would conclude that measurement variability was adequately addressed. The calculated floor or emissions limit would need no adjustment. If, on the other hand, the value equal to three times the RDL were greater than the UPL, we would conclude that the calculated floor or emissions limit does not account entirely for measurement variability. If indicated, we substituted the value equal to three times the RDL for the calculated MACT floor UPL emissions level.

*Add any RDL content specific to your source category*

```{r table1, echo = FALSE, out.width = "100%", appendix = TRUE}
col1=c('Subcategory 1','Subcategory 2','Subcategory 3',
       'Subcategory 4','Subcategory 5')
col2=c('HAP 1, HAP 2', 'HAP 1', 'HAP 1, HAP 2, HAP 3', 'None', 'None')
col3=c('HAP 1, HAP 2', 'HAP 1', 'HAP 1, HAP 2, HAP 3', 'None', 'None')
col4=c('HAP 1, HAP 2, X, Y, Z', 'HAP 1, X, Y, Z', 'HAP 1, HAP 2, HAP 3, 
       X, Y, Z', 'None', 'X, Y, Z')
table1=tibble(col1=col1,col2=col2,col3=col3,col4=col4)
table1%>%knitr::kable(
  caption="Here is an example of a table in case you need one",
             col.names=c('',
                         'Existing MACT floor (old units)',
                         'Existing MACT floor (new units)',
                         'New MACT floor'))
```

\newpage
## MACT Floor Data Set

*describe any data that was excluded here and why*

```{r cond1, child = 'App_report_children/conditional5.Rmd', eval = zero_warn}
```

Tables with the full data sets used for each existing source MACT floor analysis can be found listed in Appendix A. Tables with the full data sets used for each new source MACT floor analysis can be found listed in Appendix B.

<!-- Repeat the section below for each HAP or subcategory that you want a figure and UPL calculation for -->

The final data set for the *chemical X* existing source MACT floor for *subcategory A* included `r nrow(dat_top_exist)` test runs from `r length(unique(dat_top_exist$sources))` sources. The best performer for the new source calculation was `r best_source` with an average emission of `r signif(mean(dat_top_new$emissions), digits = 4)` from `r nrow(dat_top_new)` test runs. 

```{r plot1, echo=FALSE, eval=source_fig, appendix = TRUE, results=FALSE, fig.align = "center", fig.height=4, fig.cap="Observation density of emissions by source type. Source averages are indicated by vertical lines."}
ggplot(data = dat_top_exist, aes(emissions, fill = sources))+
          geom_density(alpha = 0.5, bounds = c(0, Inf))+
          ylab("Density")+xlab("Chemical X emissions (units)")+
          ggtitle("Subcategory A top sources test run densities")+
          multi_source_theme()+labs(fill = 'Top sources', 
                                    color = 'Top sources')+
          geom_vline(aes(xintercept = means, color = sources),
             data = dat_topmeans_exist, linewidth = 1)+
          guides(fill = guide_legend(nrow = 2, byrow = TRUE),
                color = guide_legend(nrow = 2, byrow = TRUE))+
          scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
          scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
          coord_cartesian(clip = 'off')+
          geom_rug(sides = 'b', aes(x = emissions, color = sources), 
                   data = dat_top_exist, alpha = 0.5, outside = TRUE)
```

<!-- If you used non-default options in obs_density() then you will need to alter the text below -->

Figures with density plots of emission test run data were created using `ggplot2` in `R`. The density curves were generated using `obs_density()` in `UPLforOAR` version `r packageVersion('UPLforOAR')`, using a `gamma` kernel type, a zero lower bound, a bandwidth of $\sigma n^{-2/5}$ where $\sigma$ is the emission value standard deviation and $n$ is the number of runs. For further details see the `R` code in Appendix C. 

\newpage
## MACT Floor Analysis Results

<!-- Repeat the section below for each HAP or subcategory that you want a figure and UPL calculation for -->

### Existing Source *Chemical X* in *Subcategory A*

The distribution that best represents the existing source data set for *chemical X* in *subcategory A* is the `r distribution_ex` distribution, with a corresponding UPL of `r signif(results$fit_table$UPL, digits = 4)` *(insert measurement units)*. The `r distribution_ex` distribution was the best fit distribution with a $SSE$ of `r signif(results$fit_table$SSE, digits = 4)`, `r results$fit_table$Obs_in_CI` out of `r nrow(dat_top_exist)` emission values within the 95\% CI of predictions, and the probability density function integrated to approximately `r signif(results$fit_table$pdf_integral, digits = 2)`.
```{r plot2, echo=FALSE, fig.dim = c(5,4), appendix = TRUE, fig.align = "center", fig.cap = "Observation density of chemical X emissions for the overall population of test runs for the existing source MACT floor for subcateogry A. The obseration data are indicated in black as points and a rug along the axis, with the observation density distribution as a black line. The fitted distribution that is the basis of the UPL estimate is colored orange. The average of the emission test runs is the vertical black line and the UPL result is the vertical orange line."}
ggplot()+
  geom_line(data = results$pred_pdf_dat, size = 0.75,
            aes(y = ydens, x = x_hat, color = 'a'))+
  geom_area(data = results$pred_pdf_dat, alpha = 0.25,
            aes(y = ydens, x = x_hat, fill = 'a'))+
  geom_point(aes(y = ydens, x = emissions), data = results$obs_pdf_dat,
             size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_errorbar(data=results$obs_pdf_dat,alpha=0.5,
                aes(ymin=low,ymax=up,x=emissions,color='b'))+
  geom_point(data=results$obs_pdf_dat,
             aes(x=emissions,y=med,fill='b'),
             size=3,alpha=0.5,shape=21,color='black')+
  geom_line(aes(y = pdf_hat, x = x_hat, color = 'b'),
               data = results$pred_pdf_dat, size = 0.75, linetype = 2)+
  geom_area(aes(y = pdf_hat, x = x_hat, fill = 'b'), alpha = 0.25,
            data = results$pred_pdf_dat)+
  ylab("Density")+xlab("Chemical X emissions (units)")+
  ggtitle("Subcategory A existing source\nMACT floor test run population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept = (mean(dat_top_exist$emissions)), 
                 color = 'a'), linewidth = 1, linetype = 1)+
  geom_vline(aes(xintercept = results$fit_table$UPL, color = 'b'), 
             linewidth = 1, linetype = 2)+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  coord_cartesian(clip = 'off')+
  labs(color = 'Distribution:', fill = 'Distribution:')+
  geom_rug(sides = 'b', aes(x = emissions), data = results$obs_pdf_dat,
           alpha = 0.5, outside = TRUE, color = 'black')+
  scale_color_manual(values = c('black', '#FF7F00'),
                     labels = c('Observations', distribution_ex))+
  scale_fill_manual(values = c('black', '#FF7F00'),
                    labels = c('Observations', distribution_ex))
```
The `r distribution_ex` distributions parameters converge well, passing both Gelman-Rubin diagnostic checks and with well-mixed MCMC chains (left figures below) and an approximately normal posterior distributions around the converged value (right figures below).

```{r conv_figs1, results = FALSE, appendix = TRUE, fig.height = 4, out.width = "100%", fig.align = "center", fig.cap = "Convergence figures for gamma parameters for existing source MACT floor analysis for chemical X in subcategory A."}
part1 = setup_likelihood(distribution = distribution_ex, data = dat_top_exist)
part2 = run_likelihood(model_input = part1, future_runs = params$future_runs, 
                       xvals = xvals, maxY = maxY)
part3 = converge_figs(distribution = distribution_ex, jags_model_run = part2)
part3
```

### New Source *Chemical X* in *Subcategory A*

The distribution that best represents the new source data set for *chemical X* in *subcategory A* is the `r distribution_new` distribution, with a corresponding UPL of `r signif(results2$fit_table$UPL, digits = 4)` *(insert measurement units)*. The `r distribution_new` distribution was the best fit distribution with a $SSE$ of `r signif(results2$fit_table$SSE, digits = 4)`, `r results2$fit_table$Obs_in_CI` out of `r nrow(dat_top_new)` emission values within the 95\% CI of predictions, and the probability density function integrated to approximately `r signif(results2$fit_table$pdf_integral, digits = 2)`.

```{r plot3, echo = FALSE, fig.dim = c(5, 4), appendix = TRUE, fig.align = "center", fig.cap = "Observation density of chemical X emissions for the overall population of test runs for the new source MACT floor for subcateogry A. The obseration data are indicated in black as points and a rug along the axis, with the observation density distribution as a black line. The fitted distribution that is the basis of the UPL estimate is colored blue The average of the emission test runs is the vertical black line and the UPL result is the vertical blue line."}
ggplot()+
  geom_line(data = results2$pred_pdf_dat, size = 0.75,
            aes(y = ydens, x = x_hat, color = 'a'))+
  geom_area(data = results2$pred_pdf_dat, alpha = 0.25,
            aes(y = ydens, x = x_hat, fill = 'a'))+
  geom_point(aes(y = ydens, x = emissions), data = results2$obs_pdf_dat,
             size = 3, alpha = 0.5, shape = 19, color = 'black')+
  geom_errorbar(data=results2$obs_pdf_dat,alpha=0.5,
                aes(ymin=low,ymax=up,x=emissions,color='b'))+
  geom_point(data=results2$obs_pdf_dat,
             aes(x=emissions,y=med,fill='b'),
             size=3,alpha=0.5,shape=21,color='black')+
  geom_line(aes(y = pdf_hat, x = x_hat, color = 'b'),
               data = results2$pred_pdf_dat, size = 0.75, linetype = 2)+
  geom_area(aes(y = pdf_hat, x = x_hat, fill = 'b'), alpha = 0.25,
            data = results2$pred_pdf_dat)+
  ylab("Density")+xlab("Chemical X emissions (units)")+
  ggtitle("Subcategory A new source\nMACT floor test run population")+
  pop_distr_theme()+
  geom_vline(aes(xintercept = (mean(dat_top_new$emissions)), 
                 color = 'a'), linewidth = 1, linetype = 1)+
  geom_vline(aes(xintercept = results2$fit_table$UPL, color = 'b'), 
             linewidth = 1, linetype = 2)+
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))+
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)))+
  coord_cartesian(clip = 'off')+
  labs(color = 'Distribution:', fill = 'Distribution:')+
  geom_rug(sides = 'b', aes(x = emissions), data = results2$obs_pdf_dat,
           alpha = 0.5, outside = TRUE, color = 'black')+
  scale_color_manual(values = c('black', '#377EB8'),
                     labels = c('Observations', distribution_new))+
  scale_fill_manual(values = c('black', '#377EB8'),
                    labels = c('Observations', distribution_new))
```

The `r distribution_new` distributions parameters converge well, passing both Gelman-Rubin diagnostic checks and with well-mixed MCMC chains (left figures below) and an approximately normal posterior distributions around the converged value (right figures below).
```{r conv_figs2, appendix = TRUE, results = FALSE, fig.height = 4, out.width = "100%", fig.align = "center", fig.cap = "Convergence figures for skew-normal parameters for new source MACT floor analysis for chemical X in subcategory A."}
part1 = setup_likelihood(distribution = distribution_new, data = dat_top_new)
part2 = run_likelihood(model_input = part1, future_runs = params$future_runs, 
                       xvals = xvals_new, maxY = maxY_new)
part3 = converge_figs(distribution = distribution_ex, jags_model_run = part2)
part3
```
### UPL Calculation Summary

<!-- If you have multiple HAP emissions or subcategories, consider combining into one table summarizing the UPL -->

The table below includes all UPL calculations for `r params$source` existing and new source MACT floor analyses using a significance of `r (params$significance)*100`\% and `r params$future_runs` future test runs.

```{r UPL_table, out.width = "80%", appendix = TRUE}
dat_UPL = tibble(HAP = c("Chemical X"), 
                 Subcategory = c("Type A"), 
                 Existing = c(signif(results$fit_table$UPL,4)), 
                 New = c(signif(results2$fit_table$UPL, 4)))
dat_UPL %>% knitr::kable(
  caption = "Upper Predictive Limit (UPL) results for existing and new source 
  MACT floors in (measurement units)", digits = 12, 
                         col.names = c("HAP", "Subcategory", 
                                       "Existing source UPL", 
                                       "New source UPL"))
```

\newpage
## References

Dragan Doric, et al. 2007. On measuring skewness and kurtosis. Springer Science + Buisness Media B. V. 2007.

Hobbs, N. Thompson and Mevin B. Hooten. 2015. "Bayesian Models: A statistical Primer for Ecologists". Princeton University Press.

Renault O. & O. Scaillet. On the way to recovery: A parametric bias free estimation of recovery rate densities. 2004. Journal of Banking and Finance. 28:12 p. 2915-2931.

\newpage
## Appendix A: Existing Source Data Tables

<!-- Add more tables for additional HAP or subcategories -->

`r params$source` existing source MACT floor analysis data set for HAP *chemical X* in *subcategory A*:

```{r dat_EG, echo = FALSE, appendix = TRUE}
dat_top_exist %>% knitr::kable(
  caption = "Top performering source emissions data used in existing source 
  UPL calculations.", digits = 12, col.names = names(dat_top_exist))
```

\newpage
## Appendix B: New source Data Tables

<!-- Add more tables for additional HAP or subcategories -->

`r params$source` new source MACT floor analysis data set for HAP *chemical X* in *subcategory A*:

```{r dat_NSPS, echo = FALSE, appendix = TRUE}
dat_top_new %>% knitr::kable(
  caption = "Best performering source emissions data used in new source 
  UPL calculations.", digits = 12, col.names = names(dat_top_new))
```

\newpage
## Appendix C: `R` Function Code

### Disclaimer

The United States Environmental Protection Agency (EPA) GitHub project code is provided on an "as is" basis and the user assumes responsibility for its use.  EPA has relinquished control of the information and no longer has responsibility to protect the integrity , confidentiality, or availability of the information.  Any reference to specific commercial products, processes, or services by service mark, trademark, manufacturer, or otherwise, does not constitute or imply their endorsement, recommendation or favoring by EPA.  The EPA seal and logo shall not be used in any manner to imply endorsement of any commercial product or activity by EPA or the United States Government.

### MACT_existing()
```{r function_docs1, results = TRUE, message = TRUE, echo = FALSE}
MACT_existing
```

### MACT_new()
```{r function_docs2, results = TRUE, message = TRUE, echo = FALSE}
MACT_new
```

### write_likelihood()
```{r function_docs3, results = TRUE, message = TRUE, echo = FALSE}
readLines("./../../R/write_likelihood.R")
```

### setup_likelihood()
```{r function_docs4, results = TRUE, message = TRUE, echo = FALSE}
setup_likelihood
```

### run_likelihood()
```{r function_docs5, results = TRUE, message = TRUE, echo = FALSE}
run_likelihood
```

### output_likelihood()
```{r function_docs6, results = TRUE, message = TRUE, echo = FALSE}
output_likelihood
```

### converge_likelihood()
```{r function_docs7, results = TRUE, message = TRUE, echo = FALSE}
converge_likelihood
```        

### converge_figs()
```{r function_docs8, results = TRUE, message = TRUE, echo = FALSE}
converge_figs
```        

### fit_likelihood()
```{r function_docs9, results = TRUE, message = TRUE, echo = FALSE}
fit_likelihood
```   

### Bayesian_UPL()
```{r function_docs10, results = TRUE, message = TRUE, echo = FALSE}
Bayesian_UPL
```  

### obs_density()
```{r function_docs11, results = TRUE, message = TRUE, echo = FALSE}
obs_density
```  

### JAGS Model Scripts
```{r function_docs12, results = TRUE, message = TRUE, echo = FALSE}
readLines("./../JAGS/Emission_beta_JAGS.R")
readLines("./../JAGS/Emission_gamma_JAGS.R")
readLines("./../JAGS/Emission_lnorm_JAGS.R")
readLines("./../JAGS/Emission_normal_JAGS.R")
readLines("./../JAGS/Emission_skewed_JAGS.R")
readLines("./../JAGS/Manual_emission_beta_JAGS.R")
readLines("./../JAGS/Manual_emission_gamma_JAGS.R")
readLines("./../JAGS/Manual_emission_lnorm_JAGS.R")
readLines("./../JAGS/Manual_emission_normal_JAGS.R")
readLines("./../JAGS/Manual_emission_skewed_JAGS.R")
```

\newpage
## Appendix D: System and package versions
```{r versions, results = TRUE, echo = FALSE}
sessionInfo()
```

\newpage
## Appendix E: Rmd Code Chunks

```{r ref.label = knitr::all_labels(appendix == TRUE), echo = TRUE, eval = FALSE}
```

